### 2025-05-12 paper/IEEE-conference-template-062824/main.tex
paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Quantum-Guided Autoencoding for Enhanced Neutral Atom Reservoir Computing in Medical Image Classification\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured for https://ieeexplore.ieee.org  and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{
\IEEEauthorblockN{Nuno Batista}
\IEEEauthorblockA{\textit{Department of Informatics Engineering} \\
\textit{Faculty of Sciences and Technology, University of Coimbra}\\
Coimbra, Portugal \\
\href{mailto:nunomarquesbatista@gmail.com}{nunomarquesbatista@gmail.com}}

\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}

ABSTRACT PLACEHOLDER

%============================================
% ABSTRACT
%============================================
\end{abstract}

\begin{IEEEkeywords}

Reservoir Computing, Quantum-Guided Autoencoding,
Neutral Atoms, Autoencoder, Dimensionality Reduction, 
Quantum Machine Learning, Hybrid Quantum-Classical Algorithms, 
Medical Image Classification

\end{IEEEkeywords}

%============================================
% INTRODUCTION
%============================================
\section{Introduction}

INTRODUCTION PLACEHOLDER

\subsection{Subsection Example}

SUBSECTION PLACEHOLDER

%============================================
% OVERVIEW OF QUANTUM RESERVOIR COMPUTING
%============================================
\section{Overview of Quantum Reservoir Computing}

OVERVIEW PLACEHOLDER

\subsection{Neutral Atoms for Reservoir Computing}

\subsection{Feature Encoding in an Atomic Register}

%============================================
% QUANTUM-GUIDED AUTOENCODING
%============================================
\section{Overview of Quantum-Guided Autoencoding}

\subsection{Autoencoders for Dimensionality Reduction}

\subsection{Quantum-Guided Autoencoding}

%============================================
% APPLICATION OF QUANTUM-GUIDED AUTOENCODING TO RESERVOIR COMPUTING
%============================================
\section{Application of Quantum-Guided Autoencoding to Reservoir Computing}

%============================================
% REFERENCES
%============================================
\section*{References}

\end{document}

### 2025-05-12 src/autoencoder.py
slight parameter changes

@@ -213,7 +213,8 @@ def train_autoencoder(data: np.ndarray, encoding_dim: int,
                      verbose: bool = True,
                      use_batch_norm: bool = True,
                      dropout: float = 0.1,
-                     weight_decay: float = 1e-5) -> Tuple[Autoencoder, float]:
+                     autoencoder_regularization: float = 1e-5) -> Tuple[Autoencoder, float]:
+                     #weight_decay: float = 1e-5
     
     """
     Train an autoencoder for dimensionality reduction with improved feature extraction.
@@ -240,8 +241,8 @@ def train_autoencoder(data: np.ndarray, encoding_dim: int,
         Whether to use batch normalization, by default True
     dropout : float, optional
         Dropout probability, by default 0.1
-    weight_decay : float, optional
-        Weight decay for regularization, by default 1e-5
+    autoencoder_regularization : float, optional
+        Regularization parameter for autoencoder, by default 1e-5
         
     Returns
     -------
@@ -283,7 +284,11 @@ def train_autoencoder(data: np.ndarray, encoding_dim: int,
     
     # Define loss and optimizer
     criterion = nn.MSELoss()
-    optimizer = optim.Adam(model.parameters(), lr=learning_rate, weight_decay=weight_decay)
+    optimizer = optim.Adam(
+            model.parameters(), 
+            lr=learning_rate, 
+            weight_decay=autoencoder_regularization
+        )
     
     # Learning rate scheduler (remove verbose parameter)
     scheduler = optim.lr_scheduler.ReduceLROnPlateau(optimizer, 'min', patience=5, factor=0.5)
@@ -433,10 +438,9 @@ class GuidedAutoencoder:
         Dimension of quantum embeddings
     hidden_dims : Optional[List[int]], optional
         Dimensions of hidden layers for autoencoder, by default None
-    alpha : float, optional
-        Weight for reconstruction loss, by default 0.7
-    beta : float, optional
-        Weight for classification loss, by default 0.3
+    guided_lambda : float, optional
+        Weight for classification loss (0-1), by default 0.3
+        Loss = (1-lambda)*reconstruction_loss + lambda*classification_loss
     use_batch_norm : bool, optional
         Whether to use batch normalization, by default True
     dropout : float, optional
@@ -448,8 +452,7 @@ class GuidedAutoencoder:
                  output_dim: int,
                  quantum_dim: int = None,
                  hidden_dims: Optional[List[int]] = None,
-                 alpha: float = 0.7, 
[…]

### 2025-05-12 src/feature_reduction.py

@@ -78,6 +78,7 @@ def apply_autoencoder(data: Dict[str, Any],
                     use_batch_norm: bool = True,
                     dropout: float = 0.1,
                     weight_decay: float = 1e-5,
+                    autoencoder_regularization: Optional[float] = None,
                     selected_indices: Optional[np.ndarray] = None,
                     selected_features: Optional[np.ndarray] = None) -> Tuple[np.ndarray, Autoencoder, float]:
     """
@@ -107,6 +108,8 @@ def apply_autoencoder(data: Dict[str, Any],
         Dropout probability, by default 0.1
     weight_decay : float, optional
         Weight decay for regularization, by default 1e-5
+    autoencoder_regularization : Optional[float], optional
+        Regularization parameter for autoencoder, by default None
     selected_indices : Optional[np.ndarray], optional
         Indices of selected samples, by default None
     selected_features : Optional[np.ndarray], optional
@@ -145,7 +148,8 @@ def apply_autoencoder(data: Dict[str, Any],
         verbose=verbose, 
         use_batch_norm=use_batch_norm, 
         dropout=dropout, 
-        weight_decay=weight_decay
+        #weight_decay=weight_decay,
+        autoencoder_regularization=autoencoder_regularization
     )
     
     # Encode data
@@ -304,8 +308,7 @@ def apply_guided_autoencoder(data: Dict[str, Any],
                             quantum_layer,
                             encoding_dim: int = 8,
                             hidden_dims: Optional[List[int]] = None,
-                            alpha: float = 0.7,
-                            beta: float = 0.3,
+                            guided_lambda: float = 0.3,
                             batch_size: int = 32,
                             epochs: int = 50,
                             learning_rate: float = 0.001,
@@ -316,6 +319,7 @@ def apply_guided_autoencoder(data: Dict[str, Any],
                             use_batch_norm: bool = True,
                             dropout: float = 0.1,
                             weight_decay: float = 1e-5,
+                            autoencoder_regularization: Optional[float] = None,
                             selected_indices: Optional[np.ndarray] = None,
                             selected_features: Optional[np.ndarray] = None,
                             selected_targets: Optional[np.ndarray] = None) -> Tuple[np.ndarray, GuidedAutoencoder, float]:
@@ -332,10 +336,9 @@ def apply_guided_autoencoder(data: Dict[str, Any],
[…]

### 2025-05-12 src/main.py

@@ -114,6 +114,10 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
     method_name = args.reduction_method.lower()
     reduction_name = method_name.upper()  # For display in result labels
     
+    # Clean memory between runs to avoid cache issues
+    import gc
+    gc.collect()
+    
     if method_name == "pca":
         # Apply PCA reduction
         print("Using PCA for feature reduction...")
@@ -152,7 +156,7 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
             verbose=not args.no_progress,
             use_batch_norm=True,
             dropout=0.1,
-            weight_decay=1e-5,
+            autoencoder_regularization=args.autoencoder_regularization,
             selected_features=train_features
         )
         
@@ -197,8 +201,7 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
             quantum_layer=quantum_layer,
             encoding_dim=dim_reduction,
             hidden_dims=args.autoencoder_hidden_dims,
-            alpha=args.guided_alpha,
-            beta=args.guided_beta,
+            guided_lambda=args.guided_lambda,
             batch_size=args.guided_batch_size,
             epochs=args.autoencoder_epochs,
             learning_rate=args.autoencoder_learning_rate,
@@ -208,10 +211,13 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
             verbose=not args.no_progress,
             use_batch_norm=True,  # Enable batch normalization
             dropout=0.1,  # Add dropout for regularization
-            weight_decay=1e-5,  # Add weight decay
+            autoencoder_regularization=args.autoencoder_regularization,  # Use autoencoder regularization
             selected_features=train_features,
             selected_targets=train_targets
         )
+        
+        # Explicitly clear cache after training
+        reduction_model.clear_cache()
         
         # Log the spectral range to help diagnose scaling issues
         print(f"Encoded data spectral range: {spectral}")
@@ -307,7 +313,7 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
         y_train=ys, 
         x_test=test_features, 
         y_test=test_targets, 
-        regularization=args.regularization, 
+        regularization=args.classifier_regularization,  # Use classifier regularization 
[…]

### 2025-05-12 src/parameter_sweep/run_experiments.py

@@ -11,6 +11,8 @@ from config_manager import ConfigManager
 
 # Define experiment profiles with parameter ranges
 EXPERIMENT_PROFILES = {
+    # NOT DONE
+    # SHOULD SHOW THAT USING MORE QUBITS IS BETTER
     "encoding_dimensions": {
         "description": "Sweep over encoding dimensions and methods",
         "param_grid": {
@@ -19,16 +21,30 @@ EXPERIMENT_PROFILES = {
             "n_shots": [500]
         }
     },
+    
+    "autoencoder_regularization": {
+        "description": "Sweep over autoencoder regularization parameters",
+        "param_grid": {
+            "reduction_method": ["guided_autoencoder"],
+            "autoencoder_regularization": [0.0, 0.00001, 0.0001, 0.001, 0.01],
+            "dim_reduction": [12],
+            "n_shots": [500]
+        }
+    },
+    
+    # NOT DONE
     "quantum_readouts": {
         "description": "Sweep over quantum readout types and time steps",
         "param_grid": {
             "readout_type": ["Z", "ZZ", "all"],
             "time_steps": [4, 8, 12, 16],
-            "n_shots": [500],
             "reduction_method": ["guided_autoencoder"],
             "dim_reduction": [12]
         }
     },
+    
+    # NOT DONE
+    # UNSURE
     "shot_noise": {
         "description": "Sweep over number of quantum shots",
         "param_grid": {
@@ -38,25 +54,42 @@ EXPERIMENT_PROFILES = {
             "dim_reduction": [12]
         }
     },
-    "guided_autoencoder": {
-        "description": "Sweep over guided autoencoder parameters",
+    
+    # RUNNING THIS ONE NOW
+    "guided_autoencoder_lambda": {
+        "description": "Sweep over guided autoencoder lambda parameter",
         "param_grid": {
             "reduction_method": ["guided_autoencoder"],
-            "guided_alpha": [0.3, 0.5, 0.7],
-            "guided_beta": [0.7, 0.5, 0.3],
+            "guided_lambda": [0.1, 0.3, 0.5, 0.7, 0.9],
+            "quantum_update_frequency": [1],
+            "dim_reduction": [12]
+        }
+    },
+    
+    # NOT DONE
+    # SHOULD SHOW THAT THE MORE YOU QUERY THE RESERVOIR THE BETTER
+    "guided_autoencoder_update_frequency": {
[…]

### 2025-05-15 src/autoencoder.py
reaedme, licence, deeper surrogate, fixed problems with softmax

@@ -33,7 +33,6 @@ class Autoencoder(nn.Module):
 
         # Default architecture if hidden_dims not provided
         if hidden_dims is None:
-            # TODO: MIGHT BE BETTER TO USE A FUNCTION OF THE INPUT DIMENSION
             hidden_dims = [
                 max(input_dim // 2, encoding_dim * 8),
                 max(input_dim // 4, encoding_dim * 4), 
@@ -43,7 +42,6 @@ class Autoencoder(nn.Module):
             # Remove layers that are smaller than encoding_dim
             hidden_dims = [dim for dim in hidden_dims if dim > encoding_dim]
 
-
         print("""
 
             **************************
@@ -51,7 +49,6 @@ class Autoencoder(nn.Module):
             **************************        
         
             """)
-
 
         # Create encoder layers
         encoder_layers = []
@@ -92,7 +89,6 @@ class Autoencoder(nn.Module):
             # Update previous dimension
             prev_dim = dim
         
-        
         # Final encoding layer
         encoder_layers.append(
             nn.Linear(
@@ -111,7 +107,6 @@ class Autoencoder(nn.Module):
             
         self.encoder = nn.Sequential(*encoder_layers)
         
-        
         print("""
 
             **************************
@@ -119,7 +114,6 @@ class Autoencoder(nn.Module):
             **************************    
         
             """)
-
 
         # Create decoder layers (reverse of encoder)
         decoder_layers = []
@@ -203,6 +197,205 @@ class Autoencoder(nn.Module):
         return self.encoder(x)
 
 
+def prepare_autoencoder_data(data: np.ndarray, batch_size: int, verbose: bool = True) -> Tuple[torch.Tensor, torch.utils.data.DataLoader]:
+    """
+    Prepare data for autoencoder training.
+    
+    Parameters
+    ----------
+    data : np.ndarray
+        Input data with shape (n_features, n_samples)
+    batch_size : int
+        Batch size for training
+    verbose : bool, optional
+        Whether to show warning messages, by default True
+        
+    Returns
+    -------
+    Tuple[torch.Tensor, torch.utils.data.DataLoader]
+        Input tensor and data loader
+    """
+    # Transpose data to have samples as first dimension (PyTorch convention)
+    X = torch.tensor(data.T, dtype=torch.float32)
+    
+    # Convert batch_size to Python native int to avoid PyTorch DataLoader errors
+    batch_size = int(batch_size)
+    
+    # Adjust batch size if it's larger than dataset size
[…]

### 2025-05-15 src/feature_reduction.py

@@ -409,7 +409,7 @@ def apply_guided_autoencoder(data: Dict[str, Any],
         verbose=verbose,
         use_batch_norm=use_batch_norm, 
         dropout=dropout, 
-        autoencoder_regularization=autoencoder_regularization
+        autoencoder_regularization=autoencoder_regularization,
     )
     
     # Encode data

### 2025-05-15 src/guided_autoencoder.py

@@ -9,7 +9,8 @@ from typing import Tuple, List, Dict, Any, Optional
 from autoencoder import Autoencoder
 from qrc_layer import DetuningLayer
 from quantum_surrogate import QuantumSurrogate, create_and_train_surrogate, train_surrogate
-    
+from models import LinearClassifier  # Import LinearClassifier from models.py
+
 class GuidedAutoencoder:
     """
     Quantum guided autoencoder that jointly optimizes reconstruction and quantum classification.
@@ -68,15 +69,10 @@ class GuidedAutoencoder:
         """Initialize classifier with the correct input dimension"""
         self.quantum_dim = quantum_dim
         
-        # TODO: MIGHT USE MORE THAN A SINGLE LINEAR LAYER
-        self.classifier = nn.Sequential(
-            nn.Linear(
-                    in_features=quantum_dim, 
-                    out_features=self.output_dim
-                ),
-            nn.Softmax(
-                    dim=1
-                )
+        # Use LinearClassifier from models.py
+        self.classifier = LinearClassifier(
+            input_dim=quantum_dim,
+            output_dim=self.output_dim,
         )
         return self.classifier
     
@@ -103,6 +99,375 @@ class GuidedAutoencoder:
         self.quantum_embeddings_cache = {}
 
 
+def prepare_guided_autoencoder_data(data: np.ndarray, labels: np.ndarray, 
+                                   batch_size: int, device: str, 
+                                   verbose: bool = True) -> Tuple[torch.Tensor, torch.Tensor, torch.utils.data.DataLoader]:
+    """
+    Prepare data for guided autoencoder training.
+    
+    Parameters
+    ----------
+    data : np.ndarray
+        Input data with shape (n_features, n_samples)
+    labels : np.ndarray
+        Label data
+    batch_size : int
+        Batch size for training
+    device : str
+        Device to use ('cpu' or 'cuda')
+    verbose : bool, optional
+        Whether to show warning messages, by default True
+        
+    Returns
+    -------
+    Tuple[torch.Tensor, torch.Tensor, torch.utils.data.DataLoader]
+        Input tensor, label tensor, and data loader
+    """
+    # Get number of samples and output classes
+    n_samples = data.shape[1]
+    output_dim = len(np.unique(labels))
+    
+    # Convert to PyTorch tensors
+    X = torch.tensor(data.T, dtype=torch.float32).to(device)
+    
[…]

### 2025-05-15 src/quantum_surrogate.py

@@ -5,6 +5,65 @@ import numpy as np
 from tqdm import tqdm
 from typing import Dict, Any, Optional, Tuple, List
 
+def get_proportional_hidden_dims(input_dim: int, output_dim: int, 
+                               scaling_factor: float = 1.0, 
+                               min_width: int = 256, 
+                               max_width: int = 8192) -> List[int]:
+    """
+    Calculate hidden dimensions proportional to input and output dimensions.
+    
+    Parameters
+    ----------
+    input_dim : int
+        Dimension of input features
+    output_dim : int
+        Dimension of quantum embeddings
+    scaling_factor : float, optional
+        Scale the network width, by default 1.0
+    min_width : int, optional
+        Minimum width for any layer, by default 256
+    max_width : int, optional
+        Maximum width for any layer, by default 8192
+        
+    Returns
+    -------
+    List[int]
+        List of hidden layer dimensions
+    """
+    # Base the network size on geometric mean of input and output dimensions
+    geo_mean = int(np.sqrt(input_dim * output_dim) * scaling_factor)
+    
+    # Determine network width based on dimensions
+    max_dim = max(input_dim, output_dim)
+    
+    # Scale network depth based on dimensionality
+    num_layers = max(3, min(6, int(np.log2(max_dim / 128))))
+    
+    # Determine maximum hidden layer width
+    max_hidden = min(max(geo_mean, min_width), max_width)
+    
+    # Create pyramid architecture that expands then contracts
+    hidden_dims = []
+    
+    # First expanding half
+    for i in range(num_layers // 2):
+        layer_width = min_width + int((max_hidden - min_width) * 
+                                     ((i + 1) / (num_layers // 2 + 1)))
+        hidden_dims.append(layer_width)
+    
+    # Middle layer(s) at maximum width
+    hidden_dims.append(max_hidden)
+    if num_layers % 2 == 0:
+        hidden_dims.append(max_hidden)
+    
+    # Contracting half
+    for i in range(num_layers // 2):
+        layer_width = max_hidden - int((max_hidden - min_width) * 
+                                     ((i + 1) / (num_layers // 2 + 1)))
+        hidden_dims.append(layer_width)
+    
+    return hidden_dims
+
 class QuantumSurrogate(nn.Module):
     """
     Neural network surrogate model that mimics the quantum layer.
@@ -27,17 +86,31 @@ class QuantumSurrogate(nn.Module):
     def __init__(self, 
[…]

### 2025-05-15 paper/IEEE-conference-template-062824/main.tex
more structured paper

@@
 \begin{abstract}
-
-ABSTRACT PLACEHOLDER
-
-%============================================
-% ABSTRACT
-%============================================
+This paper introduces a novel quantum-classical hybrid system for medical image classification using neutral atom quantum processors.
+We present the Quantum Guided Autoencoder with Reservoir Surrogate (QGARS), which addresses the gradient barrier problem in quantum-classical hybrid learning by introducing a differentiable surrogate model that enables end-to-end training.
+Our approach leverages quantum reservoir computing principles with Rydberg atom arrays while optimizing feature encoding through a specialized autoencoder that jointly minimizes reconstruction error and maximizes quantum classification performance.
+We evaluate our approach on polyp detection and classification tasks, demonstrating superior performance compared to traditional dimensionality reduction techniques and standard quantum reservoir computing implementations.
+We further present detailed ablation studies analyzing the impact of various quantum parameters, guided learning coefficients, and surrogate model architectures.
+Our results suggest that quantum guidance can significantly enhance feature encoding for quantum processing, pointing toward a practical pathway for quantum advantage in medical image analysis.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-
 Reservoir Computing, Quantum-Guided Autoencoding,
@@
 Medical Image Classification
-
 \end{IEEEkeywords}
@@
 \section{Introduction}
-
-INTRODUCTION PLACEHOLDER
-
-\subsection{Subsection Example}
-
-SUBSECTION PLACEHOLDER
+\subsection{Background and Motivation}
+\subsection{Challenges in Quantum-Classical Hybrid Systems}
+\subsection{Contributions of This Work}
 
 %============================================
-% OVERVIEW OF QUANTUM RESERVOIR COMPUTING
+% BACKGROUND 
 %============================================
-\section{Overview of Quantum Reservoir Computing}
-
-OVERVIEW PLACEHOLDER
-
-\subsection{Neutral Atoms for Reservoir Computing}
-
-\subsection{Feature Encoding in an Atomic Register}
+\section{Background}
+\subsection{Quantum Computing with Neutral Atoms}
+\subsection{Principles of Reservoir Computing}
+\subsection{Quantum Reservoir Computing}
+\subsection{Dimensionality Reduction for Image Data}
+\subsubsection{Principal Component Analysis}
+\subsubsection{Autoencoder Architectures}
 
 %============================================
-% QUANTUM-GUIDED AUTOENCODING
+% METHODOLOGY
 %============================================
-\section{Overview of Quantum-Guided Autoencoding}
-
-\subsection{Autoencoders for Dimensionality Reduction}
-
-\subsection{Quantum-Guided Autoencoding}
+\section{Methodology}
+\subsection{System Architecture Overview}
+\subsection{Quantum Guided Autoencoder}
+\subsubsection{Loss Function Design}
+\subsubsection{Balancing Reconstruction and Classification}
+\subsection{The Gradient Barrier Problem}
+\subsection{Surrogate Modeling for Quantum Layers}
+\subsubsection{Architecture and Training}
+\subsubsection{Gradient Flow Through Surrogate Models}
+\subsection{Rydberg Hamiltonian and Quantum Dynamics}
+\subsection{Data Encoding Schemes}
+\subsection{Quantum Readout Methods}
+\subsubsection{Single-atom Measurements}
+\subsubsection{Two-atom Correlations}
+\subsubsection{Three-atom Correlations}
 
 %============================================
-% APPLICATION OF QUANTUM-GUIDED AUTOENCODING TO RESERVOIR COMPUTING
+% EXPERIMENTAL SETUP
 %============================================
-\section{Application of Quantum-Guided Autoencoding to Reservoir Computing}
+\section{Experimental Setup}
+\subsection{Datasets}
+\subsection{Implementation Details}
+\subsubsection{Quantum Simulation Parameters}
+\subsubsection{Classical Network Architectures}
+\subsection{Comparison Methods}
+\subsection{Performance Metrics}
+\subsection{Parameter Sweep Strategy}
+
+%============================================
+% RESULTS
+%============================================
+\section{Results and Discussion}
+\subsection{Classification Performance Comparison}
+\subsection{Ablation Studies}
+\subsubsection{Impact of Guided Lambda Parameter}
+\subsubsection{Effect of Quantum Update Frequency}
+\subsubsection{Influence of Quantum Parameters}
+\subsection{Dimensionality Reduction Comparison}
+\subsection{Surrogate Model Fidelity Analysis}
+\subsection{Generalization to Unseen Data}
+
+%============================================
+% THEORETICAL ANALYSIS
+%============================================
+\section{Theoretical Analysis}
+\subsection{Information Encoding in Quantum Reservoirs}
+\subsection{Gradient Flow in Quantum-Classical Hybrid Systems}
+\subsection{Computational Complexity}
+\subsection{Quantum Resource Requirements}
+
+%============================================
+% LIMITATIONS AND FUTURE WORK
+%============================================
+\section{Limitations and Future Work}
+\subsection{Current Limitations}
+\subsection{Potential Extensions}
+\subsection{Hardware Implementation Considerations}
+
+%============================================
+% CONCLUSION
+%============================================
+\section{Conclusion}
 

### 2025-05-15 src/quantum_surrogate.py
surrogate scaling factor

@@ -97,7 +97,7 @@ class QuantumSurrogate(nn.Module):
             #hidden_dims = [256, 512, 256]
             # For high-dimensional quantum embeddings
         if output_dim > 1000:
-            scaling_factor = 1.2
+            scaling_factor = 2
         else:
             scaling_factor = 1.0
             
@@ -262,7 +262,7 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
     
     # For early stopping
     best_loss = float('inf')
-    patience = 10
+    patience = 20
     patience_counter = 0
     best_model_state = None
     
@@ -409,7 +409,7 @@ def create_and_train_surrogate(quantum_layer: Any,
         quantum_embeddings=quantum_embeddings,
         epochs=50,
         batch_size=32,
-        learning_rate=0.001,
+        learning_rate=0.003,
         device=device,
         verbose=verbose,
         n_shots=n_shots

### 2025-05-15 src/data/__init__.py
voodoo

@@ -1,0 +1,18 @@
+from .data_processing import (
+    load_dataset, show_sample_image, flatten_images,
+    one_hot_encode, select_random_samples
+)
+from .feature_reduction import (
+    apply_pca, apply_pca_to_test_data,
+    apply_autoencoder, apply_autoencoder_to_test_data,
+    apply_guided_autoencoder, apply_guided_autoencoder_to_test_data,
+    scale_to_detuning_range
+)
+
+__all__ = [
+    'load_dataset', 'show_sample_image', 'flatten_images', 'one_hot_encode', 
+    'select_random_samples', 'apply_pca', 'apply_pca_to_test_data',
+    'apply_autoencoder', 'apply_autoencoder_to_test_data',
+    'apply_guided_autoencoder', 'apply_guided_autoencoder_to_test_data',
+    'scale_to_detuning_range'
+]

### 2025-05-15 src/data/feature_reduction.py

@@ -2,10 +2,11 @@ import numpy as np
 from typing import Dict, Tuple, Optional, Any, Union, List, TYPE_CHECKING
 from sklearn.decomposition import PCA
 from tqdm import tqdm
-from data_processing import flatten_images
+from .data_processing import flatten_images  # Fix relative import
 
-from autoencoder import Autoencoder, train_autoencoder, encode_data
-from guided_autoencoder import GuidedAutoencoder, train_guided_autoencoder, encode_data_guided
+# Replace relative imports with absolute imports
+from models.autoencoder import Autoencoder, train_autoencoder, encode_data  # Fix import path
+from models.guided_autoencoder import GuidedAutoencoder, train_guided_autoencoder, encode_data_guided  # Fix import path
 
 
 def apply_pca(data: Dict[str, Any], 

### 2025-05-15 src/main.py

@@ -7,26 +7,27 @@ import matplotlib.pyplot as plt
 from typing import Dict, Any, Tuple, Optional
 import time
 
-# Fix the import path for the qrc_polyp_python module
+# Fix the import path for modules
 sys.path.append(os.path.dirname(os.path.abspath(__file__)))
 
-# Import custom modules
-from autoencoder import Autoencoder
-from guided_autoencoder import GuidedAutoencoder
+# Import custom modules from new structure
+from models.autoencoder import Autoencoder
+from models.guided_autoencoder import GuidedAutoencoder
 
-from data_processing import load_dataset, show_sample_image, flatten_images, one_hot_encode, select_random_samples
-from feature_reduction import (
+from data.data_processing import load_dataset, show_sample_image, flatten_images, one_hot_encode, select_random_samples
+from data.feature_reduction import (
     apply_pca, apply_pca_to_test_data, 
     apply_autoencoder, apply_autoencoder_to_test_data,
     apply_guided_autoencoder, apply_guided_autoencoder_to_test_data,
     scale_to_detuning_range
 )
-from qrc_layer import DetuningLayer
-from training import train
-from visualization import plot_training_results, print_results
+from models.qrc_layer import DetuningLayer
+from models.models import NeuralNetwork, LinearClassifier
+from models.training import train
+from visualization.visualization import plot_training_results, print_results
 
 # Update import to use the new function
-from cli_utils import get_args
+from utils.cli_utils import get_args
 import argparse
 
 def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarray, np.ndarray, np.ndarray, torch.nn.Module]]:
@@ -141,8 +142,8 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
         
     elif method_name == "autoencoder":
         # Use GPU if available and requested
-        device = 'cuda' if args.gpu and torch.cuda.is_available() else 'cpu'
-        if args.gpu and not torch.cuda.is_available():
+        device = 'cuda' if args.gpu_autoencoder and torch.cuda.is_available() else 'cpu'
+        if args.gpu_autoencoder and not torch.cuda.is_available():
             print("Warning: GPU requested but not available. Using CPU instead.")
         
         print(f"Using autoencoder for feature reduction (device: {device})...")
@@ -195,7 +196,8 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
[…]

### 2025-05-15 src/models/__init__.py

@@ -1,0 +1,14 @@
+# Use absolute imports to avoid circular dependencies
+from models.autoencoder import Autoencoder, train_autoencoder, encode_data
+from models.guided_autoencoder import GuidedAutoencoder, train_guided_autoencoder, encode_data_guided
+from models.quantum_surrogate import QuantumSurrogate, train_surrogate, create_and_train_surrogate
+from models.models import LinearClassifier, NeuralNetwork, QRCModel
+from models.training import train
+
+__all__ = [
+    'Autoencoder', 'train_autoencoder', 'encode_data',
+    'GuidedAutoencoder', 'train_guided_autoencoder', 'encode_data_guided',
+    'QuantumSurrogate', 'train_surrogate', 'create_and_train_surrogate',
+    'LinearClassifier', 'NeuralNetwork', 'QRCModel',
+    'train'
+]

### 2025-05-15 src/models/guided_autoencoder.py

@@ -6,10 +6,10 @@ import numpy as np
 from tqdm import tqdm
 from typing import Tuple, List, Dict, Any, Optional
 
-from autoencoder import Autoencoder
-from qrc_layer import DetuningLayer
-from quantum_surrogate import QuantumSurrogate, create_and_train_surrogate, train_surrogate
-from models import LinearClassifier  # Import LinearClassifier from models.py
+from .autoencoder import Autoencoder  # Fix relative import
+from .qrc_layer import DetuningLayer  # Fix import path
+from .quantum_surrogate import QuantumSurrogate, create_and_train_surrogate, train_surrogate  # Fix relative import
+from .models import LinearClassifier  # Fix relative import
 
 class GuidedAutoencoder:
     """

### 2025-05-15 src/models/qrc_layer.py

@@ -2,11 +2,11 @@ import numpy as np
 from bloqade.analog.ir.location import Chain
 from tqdm import tqdm
 from typing import Dict, Any, Union, Optional, List, Callable
-from quantum_task import get_embeddings_emulation
+from quantum_backends import get_backend  # Direct import from backends package
 
 class DetuningLayer:
     """
-    Bloqade implementation of the quantum dynamics simulation.
+    High-level interface to quantum simulation backends.
     
     Parameters
     ----------
@@ -26,6 +26,10 @@ class DetuningLayer:
         Type of readout measurements ("Z", "ZZ", or "all")
     encoding_scale : float, optional
         Scale for encoding features as detunings
+    backend : str, optional
+        Quantum simulation backend ('bloqade' or 'qutip')
+    use_gpu : bool, optional
+        Whether to use GPU acceleration with QuTiP
     """
     def __init__(self, 
                  geometry: str = 'chain', 
@@ -36,6 +40,8 @@ class DetuningLayer:
                  n_steps: int = 12,
                  readout_type: str = "all",
                  encoding_scale: float = 9.0,
+                 backend: str = "bloqade",
+                 use_gpu: bool = False,
                  print_params: bool = True):
         
         # Create chain geometry (only supported option)
@@ -52,12 +58,18 @@ class DetuningLayer:
         self.qrc_params = {
             "atom_number": n_atoms_int,
             "geometry_spec": atom_geometry,
+            "lattice_spacing": lattice_spacing_float,
             "encoding_scale": float(encoding_scale),
             "rabi_frequency": float(rabi_freq),
             "total_time": float(t_end),
             "time_steps": int(n_steps),
             "readouts": readout_type,
+            "backend": backend,
+            "use_gpu": use_gpu
         }
+        
+        # Initialize the backend
+        self.backend = get_backend(backend)
         
         if print_params:
             print(f"""
@@ -73,6 +85,8 @@ class DetuningLayer:
                     *    Number of time steps: {int(n_steps)}
                     *    Readout type: {readout_type}
                     *    Encoding scale: {float(encoding_scale)}
+                    *    Backend: {backend}
+                    *    Use GPU: {use_gpu}
                     *    
                     *******************************************
                   """)
@@ -97,27 +111,47 @@ class DetuningLayer:
         """
         if len(x.shape) == 1:
             # Single sample
-            return get_embeddings_emulation(x.reshape(-1, 1), self.qrc_params, 1, n_shots)
[…]

### 2025-05-15 src/models/training.py

@@ -5,7 +5,7 @@ import torch.optim as optim
 from torch.utils.data import DataLoader, TensorDataset
 from tqdm import tqdm
 from typing import Tuple, List, Dict, Any, Optional, Union
-from models import LinearClassifier, NeuralNetwork
+from .models import LinearClassifier, NeuralNetwork
 
 def train(x_train: np.ndarray, y_train: np.ndarray, 
           x_test: np.ndarray, y_test: np.ndarray, 

### 2025-05-15 src/parameter_sweep/__init__.py

@@ -1,1 +1,8 @@
 from .parameter_sweep import ParameterSweep
+from .results_manager import save_experiment_results, load_experiment_results
+from .reduction_methods_comparison import plot_qrc_accuracy_comparison
+
+__all__ = [
+    'ParameterSweep', 'save_experiment_results', 
+    'load_experiment_results', 'plot_qrc_accuracy_comparison'
+]

### 2025-05-15 src/quantum_backends/__init__.py

@@ -1,0 +1,4 @@
+from .backend_factory import get_backend
+from .backend_interface import QuantumBackend, process_results
+
+__all__ = ['get_backend', 'QuantumBackend', 'process_results']

### 2025-05-15 src/quantum_backends/backend_factory.py

@@ -1,0 +1,25 @@
+from typing import Dict, Any, Optional
+from .backend_interface import QuantumBackend
+
+def get_backend(backend_name: str) -> QuantumBackend:
+    """
+    Factory function to get the appropriate backend.
+    
+    Parameters
+    ----------
+    backend_name : str
+        Name of the backend ('bloqade' or 'qutip')
+    
+    Returns
+    -------
+    QuantumBackend
+        An instance of the requested backend
+    """
+    if backend_name.lower() == 'bloqade':
+        from .bloqade.simulator import BloqadeBackend
+        return BloqadeBackend()
+    elif backend_name.lower() == 'qutip':
+        from .qutip.simulator import QuTiPBackend
+        return QuTiPBackend()
+    else:
+        raise ValueError(f"Unknown backend: {backend_name}. Available backends: 'bloqade', 'qutip'")

### 2025-05-15 src/quantum_backends/backend_interface.py

@@ -1,0 +1,100 @@
+from abc import ABC, abstractmethod
+from typing import Dict, Any, List, Tuple, Optional, Union
+import numpy as np
+
+class QuantumBackend(ABC):
+    """Abstract base class defining the interface for quantum backends."""
+    
+    @abstractmethod
+    def build_task(self, qrc_parameters: Dict[str, Any], detunings: np.ndarray) -> Any:
+        """
+        Build a quantum task for the backend.
+        
+        Parameters
+        ----------
+        qrc_parameters : Dict[str, Any]
+            Dictionary with QRC parameters
+        detunings : np.ndarray
+            Array of detunings for each atom
+        
+        Returns
+        -------
+        Any
+            A backend-specific task ready to be executed
+        """
+        pass
+    
+    @abstractmethod
+    def get_embeddings(self, xs: np.ndarray, qrc_params: Dict[str, Any], 
+                      num_examples: int, n_shots: int = 1000) -> np.ndarray:
+        """
+        Get quantum embeddings for input data.
+        
+        Parameters
+        ----------
+        xs : np.ndarray
+            Input data
+        qrc_params : Dict[str, Any]
+            Dictionary with QRC parameters
+        num_examples : int
+            Number of examples to process
+        n_shots : int, optional
+            Number of shots for simulation, by default 1000
+        
+        Returns
+        -------
+        np.ndarray
+            Array with quantum embeddings
+        """
+        pass
+
+# Helper functions shared across backends
+def process_results(QRC_parameters: Dict[str, Any], report: Any) -> np.ndarray:
+    """
+    Process the results from a quantum task.
+    
+    Parameters
+    ----------
+    QRC_parameters : Dict[str, Any]
+        Dictionary with QRC parameters
+    report : Any
+        Report from the quantum task
+    
+    Returns
+    -------
+    np.ndarray
+        Array of expectation values
+    """
+    # Initialize array for embedding vector
+    embedding = []
+    atom_number = QRC_parameters["atom_number"]
+    readout_type = QRC_parameters.get("readouts", "ZZ")
+    time_steps = QRC_parameters["time_steps"]
+    
+    # Process bitstrings for each time step
+    for t in range(time_steps):
+        # Convert bit values (0,1) to spin values (-1,+1)
+        spin_values = -1.0 + 2.0 * (report.bitstrings()[t])
+        n_shots = spin_values.shape[0]
+        
+        # Process based on readout type
+        
[…]

### 2025-05-15 src/quantum_backends/bloqade/__init__.py

@@ -1,0 +1,3 @@
+from .simulator import BloqadeBackend
+
+__all__ = ['BloqadeBackend']

### 2025-05-15 src/quantum_backends/bloqade/simulator.py

@@ -1,0 +1,103 @@
+import numpy as np
+from bloqade.analog.ir.location import Chain
+from typing import Dict, Any
+from tqdm import tqdm
+from ..backend_interface import QuantumBackend, process_results
+
+class BloqadeBackend(QuantumBackend):
+    """Bloqade backend implementation."""
+    
+    def build_task(self, QRC_parameters: Dict[str, Any], detunings: np.ndarray):
+        """
+        Build a quantum task using Bloqade.
+        
+        Parameters
+        ----------
+        QRC_parameters : Dict[str, Any]
+            Dictionary with QRC parameters
+        detunings : np.ndarray
+            Array of detunings for each atom
+        
+        Returns
+        -------
+            A Bloqade Program ready to be executed
+        """
+        # Get parameters
+        atom_geometry = QRC_parameters["geometry_spec"]
+        rabi_frequency = float(QRC_parameters["rabi_frequency"])  # Convert to standard float
+        total_time = float(QRC_parameters["total_time"])  # Convert to standard float
+        encoding_scale = float(QRC_parameters["encoding_scale"])  # Convert to standard float
+        
+        # Time between consecutive probes
+        delta_t = total_time / int(QRC_parameters["time_steps"])  # Convert to standard int
+        
+        # Create the bloqade program using the flexible API
+
+        # Initialize the program with the atom geometry
+        program = atom_geometry
+            
+        # We can customize the rabi amplitude and detuning profiles (constant for now)
+        # Add Rabi amplitude based on the constant profile
+        program = program.rydberg.rabi.amplitude.uniform.constant(
+            duration="run_time",
+            value=rabi_frequency
+        )
+            
+        # Add detuning based on the constant profile
+        program = program.detuning.uniform.constant(
+            duration="run_time",
+            value=encoding_scale/2
+        ).scale(list(detunings)).constant(
+            duration="run_time",
+            value=-encoding_scale
+        )
+        
+        # Batch assign to probe the quantum system at multiple timesteps
+        program_job = program.batch_assign(
+            run_time=np.arange(1, int(QRC_parameters["time_steps"])+1, 1) * delta_t
+        )
+        
+        return program_job
+    
+    def get_embeddings(self, xs: np.ndarray, qrc_params: Dict[str, Any], 
+                       num_examples: int, n_shots: int = 1000) -> np.ndarray:
+        """
+        Function to get the embeddings from the Bloqade quantum task.
+        
+        Parameters
[…]

### 2025-05-15 src/quantum_backends/qutip/__init__.py

@@ -1,0 +1,3 @@
+from .simulator import QuTiPBackend
+
+__all__ = ['QuTiPBackend']

### 2025-05-15 src/quantum_backends/qutip/simulator.py

@@ -1,0 +1,409 @@
+import numpy as np
+import qutip as qt
+from typing import Dict, Any, List, Tuple, Optional
+from tqdm import tqdm
+from ..backend_interface import QuantumBackend, process_results
+
+class QuTiPBackend(QuantumBackend):
+    """QuTiP backend implementation."""
+    
+    def create_rydberg_hamiltonian(
+        self,
+        n_atoms: int, 
+        lattice_spacing: float,
+        rabi_frequency: float,
+        detunings: np.ndarray,
+        encoding_scale: float,
+        use_gpu: bool = False
+    ) -> qt.Qobj:
+        """
+        Create a Rydberg atom Hamiltonian using QuTiP.
+        
+        Parameters
+        ----------
+        n_atoms : int
+            Number of atoms in the chain
+        lattice_spacing : float
+            Spacing between atoms in μm
+        rabi_frequency : float
+            Rabi frequency (constant)
+        detunings : np.ndarray
+            Array of detunings for each atom
+        encoding_scale : float
+            Scale for encoding features as detunings
+        use_gpu : bool, optional
+            Whether to use GPU for Hamiltonian construction, by default False
+            
+        Returns
+        -------
+        qt.Qobj
+            QuTiP Hamiltonian operator
+        """
+        # Try to enable CuPy data layer if requested
+        data_type = "dense"
+        cupyd_available = False
+        
+        if use_gpu:
+            try:
+                # Import qutip_cupy to register the CuPyDense data type
+                import qutip_cupy
+                try:
+                    # Test it by creating a simple object
+                    test_obj = qt.qeye(2).to("cupyd")
+                    cupyd_available = True
+                    data_type = "cupyd"
+                    print("QuTiP GPU acceleration enabled via CuPy data layer")
+                except (RuntimeError, ImportError, OSError) as gpu_err:
+                    print(f"CuPy data layer initialization failed: {str(gpu_err)}")
+                    print("Falling back to CPU mode")
+                    data_type = "dense"
+            except ImportError:
+                print("qutip_cupy package not available. Using CPU instead.")
+                data_type = "dense"
+        
+        # Define operators for each atom
+        operators = []
+        for i in range(n_atoms):
+            # Create sigma_z operators for each atom (ProjectionOperator in Rydberg system)
+            sz_i = qt.tensor([qt.sigmaz() if j == i else qt.qeye(2) for j in range(n_atoms)])
+            if data_type == "cupyd" and cupyd_available:
+                try:
[…]

### 2025-05-15 src/quantum_task.py

@@ -1,193 +1,0 @@
-import numpy as np
-from bloqade.analog.ir.location import Chain
-from typing import Dict, Any
-from tqdm import tqdm
-
-def build_task(QRC_parameters: Dict[str, Any], detunings: np.ndarray):
-    """
-    Build a quantum task using Bloqade.
-    
-    Parameters
-    ----------
-    QRC_parameters : Dict[str, Any]
-        Dictionary with QRC parameters
-    detunings : np.ndarray
-        Array of detunings for each atom
-    
-    Returns
-    -------
-        A Bloqade Program ready to be executed
-    """
-    # Get parameters
-    atom_geometry = QRC_parameters["geometry_spec"]
-    rabi_frequency = float(QRC_parameters["rabi_frequency"])  # Convert to standard float
-    total_time = float(QRC_parameters["total_time"])  # Convert to standard float
-    encoding_scale = float(QRC_parameters["encoding_scale"])  # Convert to standard float
-    
-    # Time between consecutive probes
-    delta_t = total_time / int(QRC_parameters["time_steps"])  # Convert to standard int
-    
-    # Create the bloqade program using the flexible API
-
-    # Initialize the program with the atom geometry
-    program = atom_geometry
-        
-    # We can customize the rabi amplitude and detuning profiles (constant for now)
-    # Add Rabi amplitude based on the constant profile
-    program = program.rydberg.rabi.amplitude.uniform.constant(
-        duration="run_time",
-        value=rabi_frequency
-    )
-        
-    # Add detuning based on the constant profile
-    program = program.detuning.uniform.constant(
-        duration="run_time",
-        value=encoding_scale/2
-    ).scale(list(detunings)).constant(
-        duration="run_time",
-        value=-encoding_scale
-    )
-    
-    
-    # Batch assign to probe the quantum system at multiple timesteps
-    program_job = program.batch_assign(
-        run_time=np.arange(1, int(QRC_parameters["time_steps"])+1, 1) * delta_t
-    )
-    
-    return program_job
-
-def process_results(QRC_parameters: Dict[str, Any], report: Any) -> np.ndarray:
-    """
-    Process the results from a quantum task.
-    
-    Parameters
-    ----------
-    QRC_parameters : Dict[str, Any]
-        Dictionary with QRC parameters
-    report : Any
-        Report from the quantum task
-    
-    Returns
-    -------
-    np.ndarray
-        Array of expectation values
-    """
-    # Initialize array for embedding vector
-    embedding = []
[…]

### 2025-05-15 src/utils/__init__.py

@@ -1,0 +1,7 @@
+from .config_manager import ConfigManager, get_config_args, create_default_config_if_missing
+from .cli_utils import get_args, parse_args
+
+__all__ = [
+    'ConfigManager', 'get_config_args', 'create_default_config_if_missing',
+    'get_args', 'parse_args', 'train'
+]

### 2025-05-15 src/utils/cli_utils.py

@@ -6,6 +6,7 @@ import sys
 AVAILABLE_GEOMETRIES = ["chain"] 
 AVAILABLE_READOUT_TYPES = ["Z", "ZZ", "all"]
 AVAILABLE_REDUCTION_METHODS = ["pca", "autoencoder", "guided_autoencoder"]
+AVAILABLE_BACKENDS = ["bloqade", "qutip"]  # Add the backends option
 
 def parse_args() -> argparse.Namespace:
     """Parse command line arguments"""
@@ -50,7 +51,8 @@ def parse_args() -> argparse.Namespace:
                        help="Hidden dimensions for autoencoder (e.g. --autoencoder-hidden-dims 256 128)")
     parser.add_argument("--autoencoder-regularization", type=float, default=1e-5,
                        help="Regularization strength for autoencoder training (weight decay)")
-    parser.add_argument("--gpu", action="store_true", help="Use GPU for autoencoder training if available")
+    parser.add_argument("--gpu-autoencoder", action="store_true", 
+                       help="Use GPU for autoencoder training if available")
     
     # Quantum parameters
     parser.add_argument("--geometry", type=str, choices=AVAILABLE_GEOMETRIES, default="chain",
@@ -71,6 +73,10 @@ def parse_args() -> argparse.Namespace:
                        help="Maximum detuning value")
     parser.add_argument("--encoding-scale", type=float, default=9.0,
                        help="Scale for encoding features as detunings")
+    parser.add_argument("--backend", type=str, choices=AVAILABLE_BACKENDS, default="bloqade",
+                       help="Quantum simulation backend to use (bloqade or qutip)")
+    parser.add_argument("--gpu-quantum-evolution", action="store_true", default=False,
+                       help="Use GPU acceleration with QuTiP (requires qutip-gpu)")
     
     # Training parameters
     parser.add_argument("--classifier-regularization", type=float, default=0.0005,
@@ -105,7 +111,7 @@ def get_args() -> argparse.Namespace:
     # Check if any arguments were provided (beyond script name)
     if len(sys.argv) == 1:
         # No command line args, use config file
-        from config_manager import get_config_args
+        from utils.config_manager import get_config_args  # Fix import
         return get_config_args()
     else:
         # Command line args provided, use them

### 2025-05-15 src/utils/config_manager.py

@@ -120,7 +120,7 @@ class ConfigManager:
             "autoencoder_learning_rate": 0.001,
             "autoencoder_hidden_dims": None,
             "autoencoder_regularization": 1e-5,
-            "gpu": False,
+            "gpu_autoencoder": False,  # Updated parameter name
             "geometry": "chain",
             "lattice_spacing": 10.0,
             "rabi_freq": 2*np.pi,
@@ -130,6 +130,8 @@ class ConfigManager:
             "n_shots": 1000,
             "detuning_max": 6.0,
             "encoding_scale": 9.0,
+            "backend": "bloqade",  # Default to Bloqade
+            "gpu_quantum_evolution": False,  # Updated parameter name
             "classifier_regularization": 0.0005,
             "nepochs": 100,
             "batchsize": 1000,

### 2025-05-15 src/visualization/__init__.py

@@ -1,0 +1,3 @@
+from .visualization import plot_training_results, print_results
+
+__all__ = ['plot_training_results', 'print_results']

### 2025-05-15 src/quantum_surrogate.py
gpu fix

@@ -240,14 +240,15 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
     if isinstance(input_data, np.ndarray):
         input_tensor = torch.tensor(input_data.T, dtype=torch.float32).to(device)
     else:
-        input_tensor = input_data.detach().to(device)  # Detach to ensure we don't track unnecessary gradients
-    
+        input_tensor = input_data.detach().cpu()#.to(device)  # Detach to ensure we don't track unnecessary gradients
+        
+    quantum_embeddings_cpu = quantum_embeddings.cpu() if hasattr(quantum_embeddings, 'is_cuda') else quantum_embeddings
     # Determine batch size based on dataset size
     num_samples = input_tensor.shape[0]
     actual_batch_size = min(batch_size, num_samples)
     
     # Create simple dataset and dataloader with pinned memory for faster data transfer
-    dataset = torch.utils.data.TensorDataset(input_tensor, quantum_embeddings)
+    dataset = torch.utils.data.TensorDataset(input_tensor, quantum_embeddings_cpu)
     dataloader = torch.utils.data.DataLoader(
         dataset, 
         batch_size=actual_batch_size, 

### 2025-05-15 src/feature_reduction.py
better stat tracking + surrogate loss

@@ -323,7 +323,7 @@ def apply_guided_autoencoder(data: Dict[str, Any],
                             autoencoder_regularization: Optional[float] = None,
                             selected_indices: Optional[np.ndarray] = None,
                             selected_features: Optional[np.ndarray] = None,
-                            selected_targets: Optional[np.ndarray] = None) -> Tuple[np.ndarray, GuidedAutoencoder, float]:
+                            selected_targets: Optional[np.ndarray] = None) -> Tuple[np.ndarray, GuidedAutoencoder, float, Dict[str, List[float]]]:
     """
     Apply guided autoencoder to reduce image dimensions with quantum guidance.
     
@@ -371,10 +371,11 @@ def apply_guided_autoencoder(data: Dict[str, Any],
         
     Returns
     -------
-    Tuple[np.ndarray, GuidedAutoencoder, float]
+    Tuple[np.ndarray, GuidedAutoencoder, float, Dict[str, List[float]]]
         - xs: Encoded features for the selected examples
         - model: Trained guided autoencoder model
         - spectral: Max absolute value of the encoded features (for scaling)
+        - loss_history: Dictionary containing loss history
     """
     
     # Use provided features if available, otherwise use all data
@@ -393,7 +394,7 @@ def apply_guided_autoencoder(data: Dict[str, Any],
     
     # Train guided autoencoder
     print("Training guided autoencoder with quantum feedback...")
-    model, spectral = train_guided_autoencoder(
+    model, spectral, loss_history = train_guided_autoencoder(
         data=data_flat, 
         labels=targets, 
         quantum_layer=quantum_layer, 
@@ -425,7 +426,7 @@ def apply_guided_autoencoder(data: Dict[str, Any],
     # No need to select a subset now, as we already have the selected samples
     xs = encoded_data
     
-    return xs, model, spectral
+    return xs, model, spectral, loss_history
 
 
 def apply_guided_autoencoder_to_test_data(data: Dict[str, Any], 

### 2025-05-15 src/guided_autoencoder.py

@@ -247,7 +247,7 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
                              device: str,
                              batch_size_qe: int = 50,
                              detuning_max: float = 6.0,
-                             verbose: bool = True) -> QuantumSurrogate:
+                             verbose: bool = True) -> Tuple[QuantumSurrogate, List[float]]:
     """
     Update quantum embeddings and train surrogate model.
     
@@ -272,8 +272,8 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
         
     Returns
     -------
-    QuantumSurrogate
-        Trained surrogate model
+    Tuple[QuantumSurrogate, List[float]]
+        Trained surrogate model and surrogate loss history
     """
     model.clear_cache()
     n_samples = X.shape[0]
@@ -334,9 +334,11 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
         if verbose:
             tqdm.write("Training surrogate model with current quantum embeddings...")
         
+        surrogate_loss_history = []
+        
         if model.surrogate is None:
             # Create and train new surrogate
-            surrogate = create_and_train_surrogate(
+            surrogate, surrogate_loss_history = create_and_train_surrogate(
                 quantum_layer=quantum_layer,
                 encodings=all_encodings_tensor,
                 quantum_embeddings=all_quantum_embs_tensor,
@@ -346,7 +348,7 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
             )
         else:
             # Update existing surrogate
-            surrogate = train_surrogate(
+            surrogate, surrogate_loss_history = train_surrogate(
                 surrogate_model=model.surrogate,
                 quantum_layer=quantum_layer,
                 input_data=all_encodings_tensor,
@@ -363,7 +365,7 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
         model.initialize_surrogate(surrogate)
         
     model.autoencoder.train()
-    return surrogate
+    return surrogate, surrogate_loss_history
 
 def train_guided_epoch(model: GuidedAutoencoder,
                       dataloader: torch.utils.data.DataLoader,
@@ -488,7 +490,7 @@ def train_guided_autoencoder(
     detuning_max: float = 6.0,
     recon_scale: float = 100.0,
     class_scale: float = 1.0
-) -> Tuple[GuidedAutoencoder, float]:
+) -> Tuple[GuidedAutoencoder, float, Dict[str, List[float]]]:
     """
[…]

### 2025-05-15 src/main.py

@@ -113,6 +113,9 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
     # Determine reduction dimension
     dim_reduction = args.dim_reduction
     
+    # Prepare a container for guided autoencoder losses
+    guided_autoencoder_losses = None
+    
     # Perform feature reduction based on selected method
     method_name = args.reduction_method.lower()
     reduction_name = method_name.upper()  # For display in result labels
@@ -199,7 +202,7 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
         )
         
         # Apply guided autoencoder reduction with improved parameters
-        xs_raw, reduction_model, spectral = apply_guided_autoencoder(
+        xs_raw, reduction_model, spectral, guided_autoencoder_losses = apply_guided_autoencoder(
             data_train,
             quantum_layer=quantum_layer,
             encoding_dim=dim_reduction,
@@ -385,6 +388,12 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
     print_results(results)
     if not args.no_plot:
         plot_training_results(results)
+    
+    # Save statistics if running as main script (not as part of parameter sweep)
+    if not hasattr(args, '_parameter_sweep') or not args._parameter_sweep:
+        from statistics_tracking import save_all_statistics
+        output_dir = save_all_statistics(results, guided_autoencoder_losses)
+        print(f"Saved run statistics to {output_dir}")
     
     return results
 

### 2025-05-15 src/parameter_sweep/parameter_sweep.py

@@ -161,13 +161,35 @@ class ParameterSweep:
         # Convert config to argparse Namespace
         args = ConfigManager.config_to_args(config)
         
+        # Mark this as a parameter sweep run to handle statistics properly
+        setattr(args, '_parameter_sweep', True)
+        
+        # Create experiment directory
+        exp_dir = os.path.join(sweep_dir, exp_id)
+        os.makedirs(exp_dir, exist_ok=True)
+        
         # Run the experiment
         try:
             self.experiment_counter += 1
             print(f"\nExperiment {self.experiment_counter}: Running with {exp_id}\n")
             
+            # Import statistics tracking here to avoid circular imports
+            sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), '..')))
+            from statistics_tracking import save_all_statistics
+            
             # Run main function with args
             results = main(args)
+            
+            # Save statistics in the experiment directory
+            guided_losses = None
+            
+            # Check if guided autoencoder data is available in the global scope
+            import __main__
+            if hasattr(__main__, 'guided_autoencoder_losses') and __main__.guided_autoencoder_losses is not None:
+                guided_losses = __main__.guided_autoencoder_losses
+            
+            # Save statistics
+            save_all_statistics(results, guided_losses, exp_dir)
             
             # Extract metrics
             metrics = self._extract_metrics(results)
@@ -178,10 +200,6 @@ class ParameterSweep:
                 if key not in ['autoencoder_hidden_dims', 'target_size']:
                     metrics[key] = value
                     
-            # Save individual experiment results
-            exp_dir = os.path.join(sweep_dir, exp_id)
-            os.makedirs(exp_dir, exist_ok=True)
-            
             # Save configuration
             with open(os.path.join(exp_dir, "config.json"), "w") as f:
                 config_serializable = {k: str(v) if isinstance(v, np.ndarray) else v 

### 2025-05-15 src/quantum_surrogate.py

@@ -165,7 +165,7 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
                    learning_rate: float = 0.001,
                    device: str = 'cpu',
                    verbose: bool = True,
-                   n_shots: int = 1000) -> QuantumSurrogate:
+                   n_shots: int = 1000) -> Tuple[QuantumSurrogate, List[float]]:
     """
     Train the surrogate model to mimic the quantum layer.
     
@@ -194,8 +194,8 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
         
     Returns
     -------
-    QuantumSurrogate
-        Trained surrogate model
+    Tuple[QuantumSurrogate, List[float]]
+        Trained surrogate model and loss history
     """
     # Ensure model is on the correct device and in training mode
     surrogate_model = surrogate_model.to(device)
@@ -204,6 +204,9 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
     # Set up optimizer and loss function
     optimizer = optim.Adam(surrogate_model.parameters(), lr=learning_rate)
     criterion = nn.MSELoss()
+    
+    # Track loss history
+    loss_history = []
     
     # Convert input data to numpy for quantum layer if needed
     input_numpy = None
@@ -230,7 +233,6 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
         quantum_embeddings = torch.tensor(quantum_embs.T, dtype=torch.float32)
     
     # Make sure quantum embeddings are on the correct device
-    # quantum_embeddings = torch.tensor(quantum_embeddings, dtype=torch.float32).to(device)
     if isinstance(quantum_embeddings, torch.Tensor):
         quantum_embeddings = quantum_embeddings.clone().detach().to(device).requires_grad_(True)
     else:
@@ -240,7 +242,7 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
     if isinstance(input_data, np.ndarray):
         input_tensor = torch.tensor(input_data.T, dtype=torch.float32).to(device)
     else:
-        input_tensor = input_data.detach().cpu()#.to(device)  # Detach to ensure we don't track unnecessary gradients
+        input_tensor = input_data.detach().cpu()  # Detach to ensure we don't track unnecessary gradients
         
     quantum_embeddings_cpu = quantum_embeddings.cpu() if hasattr(quantum_embeddings, 'is_cuda') else quantum_embeddings
     # Determine batch size based on dataset size
[…]

### 2025-05-15 src/visualization.py

@@ -1,7 +1,8 @@
 import matplotlib.pyplot as plt
-from typing import Dict, Tuple, List, Any
+from typing import Dict, Tuple, List, Any, Optional
 
-def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]]) -> None:
+def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]], 
+                          save_path: Optional[str] = None) -> None:
     """
     Plot training results for multiple models.
     
@@ -9,6 +10,8 @@ def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float]
     ----------
     results_dict : Dict[str, Tuple[List[float], List[float], List[float], Any]]
         Dictionary mapping model names to (losses, accs_train, accs_test, model) tuples
+    save_path : Optional[str], optional
+        Path to save the figure, by default None (only display)
     """
     plt.figure(figsize=(12, 6))
     
@@ -34,6 +37,11 @@ def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float]
     plt.title("Training Loss")
     
     plt.tight_layout()
+    
+    # Save figure if path provided
+    if save_path:
+        plt.savefig(save_path, dpi=300, bbox_inches='tight')
+    
     plt.show()
 
 def print_results(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]]) -> None:

### 2025-05-15 src/quantum_backends/qutip/simulator.py
a

@@ -7,6 +7,9 @@ from ..backend_interface import QuantumBackend, process_results
 class QuTiPBackend(QuantumBackend):
     """QuTiP backend implementation."""
     
+    # Add a class variable to track GPU availability
+    _gpu_memory_insufficient = False
+    
     def create_rydberg_hamiltonian(
         self,
         n_atoms: int, 
@@ -39,6 +42,11 @@ class QuTiPBackend(QuantumBackend):
         qt.Qobj
             QuTiP Hamiltonian operator
         """
+        # If we've already detected insufficient GPU memory, don't try to use GPU
+        if QuTiPBackend._gpu_memory_insufficient:
+            use_gpu = False
+            print("Using CPU mode (GPU previously failed due to memory limitations)")
+        
         # Try to enable CuPy data layer if requested
         data_type = "dense"
         cupyd_available = False
@@ -57,9 +65,11 @@ class QuTiPBackend(QuantumBackend):
                     print(f"CuPy data layer initialization failed: {str(gpu_err)}")
                     print("Falling back to CPU mode")
                     data_type = "dense"
+                    QuTiPBackend._gpu_memory_insufficient = True
             except ImportError:
                 print("qutip_cupy package not available. Using CPU instead.")
                 data_type = "dense"
+                QuTiPBackend._gpu_memory_insufficient = True
         
         # Define operators for each atom
         operators = []
@@ -74,6 +84,7 @@ class QuTiPBackend(QuantumBackend):
                     print("Falling back to CPU mode for all operators")
                     data_type = "dense"
                     cupyd_available = False
+                    QuTiPBackend._gpu_memory_insufficient = True
             operators.append(sz_i)
         
         # Create sigma_x operators for each atom (drive terms)
@@ -83,9 +94,11 @@ class QuTiPBackend(QuantumBackend):
             if data_type == "cupyd" and cupyd_available:
                 try:
                     sx_i = sx_i.to("cupyd")  # Convert to GPU-backed Qobj
-                except Exception:
-                    # We already reported error above, just skip conversion
-                    pass
+                except Exception as e:
+                    print(f"Warning: Failed to convert sx operator to GPU: {str(e)}")
+                    data_type = "dense"
+                    cupyd_available = False
+                    QuTiPBackend._gpu_memory_insufficient = True
             sx_operators.append(sx_i)
         
         # Create the Hamiltonian
[…]

### 2025-05-19 src/guided_autoencoder.py
more data analysis

@@ -488,7 +488,7 @@ def train_guided_autoencoder(
     detuning_max: float = 6.0,
     recon_scale: float = 100.0,
     class_scale: float = 1.0
-) -> Tuple[GuidedAutoencoder, float]:
+) -> Tuple[GuidedAutoencoder, float, Dict[str, List[float]]]:
     """
     Train a guided autoencoder jointly with quantum embeddings.
     
@@ -536,8 +536,8 @@ def train_guided_autoencoder(
         
     Returns
     -------
-    Tuple[GuidedAutoencoder, float]
-        Trained guided autoencoder and maximum absolute value for scaling
+    Tuple[GuidedAutoencoder, float, Dict[str, List[float]]]
+        Trained guided autoencoder, maximum absolute value for scaling, and loss history dictionary
     """
     
     # Prepare data
@@ -600,6 +600,13 @@ def train_guided_autoencoder(
     patience = 10
     best_model_state = None
     
+    # Track loss history
+    loss_history = {
+        "recon_loss": [],
+        "class_loss": [],
+        "total_loss": []
+    }
+    
     # Training loop
     iterator = range(epochs)
     if verbose:
@@ -637,6 +644,11 @@ def train_guided_autoencoder(
             recon_scale=recon_scale,
             class_scale=class_scale
         )
+        
+        # Store losses in history
+        loss_history["recon_loss"].append(avg_recon_loss)
+        loss_history["class_loss"].append(avg_class_loss)
+        loss_history["total_loss"].append(avg_total_loss)
         
         # Update learning rate scheduler
         prev_lr = current_lr
@@ -692,7 +704,7 @@ def train_guided_autoencoder(
     if spectral == 0:  # Avoid division by zero
         spectral = 1.0
     
-    return model, spectral
+    return model, spectral, loss_history
 
 
 def encode_data_guided(model: GuidedAutoencoder, data: np.ndarray, device: str = 'cpu', 

### 2025-05-19 src/main.py

@@ -4,7 +4,7 @@ import numpy as np
 import random
 import torch
 import matplotlib.pyplot as plt
-from typing import Dict, Any, Tuple, Optional
+from typing import Dict, Any, Tuple, Optional, List
 import time
 
 # Fix the import path for the qrc_polyp_python module
@@ -29,7 +29,7 @@ from visualization import plot_training_results, print_results
 from cli_utils import get_args
 import argparse
 
-def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarray, np.ndarray, np.ndarray, torch.nn.Module]]:
+def main(args: Optional[argparse.Namespace] = None) -> Tuple[Dict[str, Tuple[np.ndarray, np.ndarray, np.ndarray, torch.nn.Module]], Optional[Dict[str, List[float]]]]:
     """
     Main function to run the quantum reservoir computing pipeline.
     
@@ -40,8 +40,8 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
     
     Returns
     -------
-    Dict[str, Tuple[np.ndarray, np.ndarray, np.ndarray, torch.nn.Module]]
-        Dictionary of results for each model
+    Tuple[Dict[str, Tuple[np.ndarray, np.ndarray, np.ndarray, torch.nn.Module]], Optional[Dict[str, List[float]]]]
+        Dictionary of results for each model and guided autoencoder losses if available
     """
     
     # Parse arguments if not provided
@@ -392,10 +392,14 @@ def main(args: Optional[argparse.Namespace] = None) -> Dict[str, Tuple[np.ndarra
     # Save statistics if running as main script (not as part of parameter sweep)
     if not hasattr(args, '_parameter_sweep') or not args._parameter_sweep:
         from statistics_tracking import save_all_statistics
-        output_dir = save_all_statistics(results, guided_autoencoder_losses)
+        output_dir = save_all_statistics(
+            results_dict=results, 
+            guided_losses=guided_autoencoder_losses,
+            args=args  # Pass the configuration arguments
+        )
         print(f"Saved run statistics to {output_dir}")
     
-    return results
+    return results, guided_autoencoder_losses
 
 if __name__ == "__main__":
     args = get_args()

### 2025-05-19 src/parameter_sweep/data_analysis_scripts/autoencoder_regularization_analysis.py

@@ -1,0 +1,138 @@
+import os
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+from typing import Optional
+
+def load_sweep_results(filepath: str) -> pd.DataFrame:
+    """Load parameter sweep results from CSV file."""
+    if not os.path.exists(filepath):
+        raise FileNotFoundError(f"Results file not found: {filepath}")
+    
+    try:
+        df = pd.read_csv(filepath)
+        print(f"Loaded {len(df)} experiment results from {filepath}")
+    except Exception as e:
+        print(f"Error loading results: {e}")
+        df = pd.DataFrame()
+    
+    return df
+
+def plot_autoencoder_regularization_effect(df: pd.DataFrame, output_path: Optional[str] = None):
+    """Create a simple plot showing regularization vs accuracy."""
+    # Set up aesthetics
+    sns.set_style("whitegrid")
+    
+    # Sort by regularization value
+    df_sorted = df.sort_values(by='autoencoder_regularization')
+    
+    # Convert accuracy to percentage
+    y_values = df_sorted['QRC_final_test_acc'].values * 100
+    x_values = df_sorted['autoencoder_regularization'].values
+    
+    # Create a figure with appropriate size
+    plt.figure(figsize=(10, 6))
+    
+    # Plot line connecting points
+    plt.plot(x_values, y_values, '-', linewidth=2, color='#1f77b4')
+    
+    # Plot individual points
+    plt.scatter(x_values, y_values, s=80, color='#1f77b4', zorder=5)
+    
+    # Add value labels
+    for i, (x, y) in enumerate(zip(x_values, y_values)):
+        plt.annotate(f"{y:.1f}%", (x, y), xytext=(0, 10), 
+                    textcoords="offset points", ha='center')
+    
+    # Format x-axis to show values properly
+    if 0 in x_values:
+        # Use a semi-log scale with special handling for zero
+        plt.xscale('symlog', linthresh=1e-6)  # symlog handles zero values
+        
+        # Custom x-ticks for all values including zero
+        plt.xticks(x_values)
+        
+        # Set custom formatter for x-axis labels
+        from matplotlib.ticker import ScalarFormatter, FormatStrFormatter
+        formatter = ScalarFormatter()
+        formatter.set_scientific(False)
+        plt.gca().xaxis.set_major_formatter(formatter)
[…]

### 2025-05-19 src/parameter_sweep/parameter_sweep.py

@@ -12,6 +12,7 @@ sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), '..')))
 
 from main import main
 from config_manager import ConfigManager
+from statistics_tracking import save_all_statistics  # Add this import
 
 class ParameterSweep:
     """
@@ -161,23 +162,22 @@ class ParameterSweep:
         # Convert config to argparse Namespace
         args = ConfigManager.config_to_args(config)
         
+        # Mark this as a parameter sweep run to prevent duplicate saving
+        setattr(args, '_parameter_sweep', True)
+        
         # Run the experiment
         try:
             self.experiment_counter += 1
             print(f"\nExperiment {self.experiment_counter}: Running with {exp_id}\n")
             
-            # Run main function with args
-            results = main(args)
+            # Create experiment directory
+            exp_dir = os.path.join(sweep_dir, exp_id)
+            os.makedirs(exp_dir, exist_ok=True)
             
-            # Save statistics in the experiment directory
-            guided_losses = None
+            # Run main function with args - now unpacking both return values
+            results, guided_losses = main(args)
             
-            # Check if guided autoencoder data is available in the global scope
-            import __main__
-            if hasattr(__main__, 'guided_autoencoder_losses') and __main__.guided_autoencoder_losses is not None:
-                guided_losses = __main__.guided_autoencoder_losses
-            
-            # Save statistics
+            # Save statistics in the experiment directory - no need to check __main__
             save_all_statistics(results, guided_losses, exp_dir)
             
             # Extract metrics
@@ -189,10 +189,6 @@ class ParameterSweep:
                 if key not in ['autoencoder_hidden_dims', 'target_size']:
                     metrics[key] = value
                     
-            # Save individual experiment results
-            exp_dir = os.path.join(sweep_dir, exp_id)
-            os.makedirs(exp_dir, exist_ok=True)
-            
             # Save configuration
             with open(os.path.join(exp_dir, "config.json"), "w") as f:
                 config_serializable = {k: str(v) if isinstance(v, np.ndarray) else v 

### 2025-05-19 src/parameter_sweep/reduction_methods_comparison.py

@@ -1,84 +1,0 @@
-import pandas as pd
-import matplotlib.pyplot as plt
-import numpy as np
-import os
-import glob
-
-def plot_qrc_accuracy_comparison(csv_path):
-    """
-    Plot QRC accuracy comparison for different dimension reduction methods.
-    
-    Parameters:
-    -----------
-    csv_path : str
-        Path to the all_results.csv file
-    """
-    # Load the data
-    df = pd.read_csv(csv_path)
-    
-    # Filter successful experiments
-    df = df[df['status'] == 'success']
-    
-    # Create figure
-    plt.figure(figsize=(10, 6))
-    
-    # Define colors and markers for each method
-    styles = {
-        'pca': {'color': 'blue', 'marker': 'o', 'label': 'PCA'},
-        'autoencoder': {'color': 'green', 'marker': 's', 'label': 'Autoencoder'},
-        'guided_autoencoder': {'color': 'red', 'marker': '^', 'label': 'Guided Autoencoder'}
-    }
-    
-    # For each reduction method, plot a line
-    for method, style in styles.items():
-        # Filter data for this method
-        method_data = df[df['reduction_method'] == method]
-        
-        if not method_data.empty:
-            # Sort by dimension
-            method_data = method_data.sort_values('dim_reduction')
-            
-            # Plot the line
-            plt.plot(method_data['dim_reduction'], method_data['QRC_final_test_acc'], 
-                     marker=style['marker'], color=style['color'], 
-                     label=style['label'], linewidth=2, markersize=8)
-    
-    # Add labels and title
-    plt.xlabel('Dimension Reduction', fontsize=12)
-    plt.ylabel('QRC Accuracy', fontsize=12)
-    plt.title('QRC Accuracy vs Dimension Reduction by Method', fontsize=14)
-    plt.legend(fontsize=10)
-    plt.grid(True, linestyle='--', alpha=0.7)
-    
-    # Set axes limits
-    plt.ylim(0, 1.0)
-    plt.xticks(np.arange(4, 14, 2))  # 4, 6, 8, 10, 12
-    
-    # Add a note about PCA limitations
-    plt.annotate('Note: PCA failed for dimensions > 4\ndue to sample size limitations', 
-                 xy=(0.5, 0.02), xycoords='axes fraction', 
-                 bbox=dict(boxstyle="round,pad=0.3", fc="lightyellow", ec="orange", alpha=0.8),
-                 ha='center', fontsize=10)
-    
[…]

### 2025-05-19 src/statistics_tracking.py

@@ -116,16 +116,17 @@ def save_guided_autoencoder_losses(losses: Dict[str, List[float]], output_dir: s
     Parameters
     ----------
     losses : Dict[str, List[float]]
-        Dictionary with keys 'total', 'reconstruction', 'classification' mapping to loss histories
+        Dictionary with keys 'total_loss', 'recon_loss', 'class_loss' mapping to loss histories
     output_dir : str
         Directory to save the plot
     """
     # Plot main losses (reconstruction, classification, total)
     plt.figure(figsize=(10, 6))
     
-    plt.plot(losses['total'], label='Total Loss', color='blue', linewidth=2)
-    plt.plot(losses['reconstruction'], label='Reconstruction Loss', color='green', linewidth=1.5, linestyle='--')
-    plt.plot(losses['classification'], label='Classification Loss', color='red', linewidth=1.5, linestyle=':')
+    # Use the correct key names from the loss_history dictionary
+    plt.plot(losses['total_loss'], label='Total Loss', color='blue', linewidth=2)
+    plt.plot(losses['recon_loss'], label='Reconstruction Loss', color='green', linewidth=1.5, linestyle='--')
+    plt.plot(losses['class_loss'], label='Classification Loss', color='red', linewidth=1.5, linestyle=':')
     
     plt.xlabel("Epoch")
     plt.ylabel("Loss")
@@ -266,9 +267,45 @@ def save_metrics_json(metrics: Dict[str, Any], output_dir: str) -> None:
     
     logging.info(f"Saved metrics to {metrics_path}")
 
+def save_config_file(args: Any, output_dir: str) -> None:
+    """
+    Save the configuration as a JSON file.
+    
+    Parameters
+    ----------
+    args : Any
+        Command line arguments or configuration object
+    output_dir : str
+        Directory to save the configuration
+    """
+    # Convert args to dictionary if it's not already
+    if hasattr(args, '__dict__'):
+        config = vars(args)
+    else:
+        config = args
+    
+    # Convert numpy arrays and other non-serializable types
+    config_serializable = {}
+    for k, v in config.items():
+        if isinstance(v, np.ndarray):
+            config_serializable[k] = v.tolist()
+        elif isinstance(v, (int, float, str, bool, list, dict, tuple)) or v is None:
+            config_serializable[k] = v
+        else:
+            # Convert other types to string representation
+            config_serializable[k] = str(v)
+    
+    # Save as JSON
+    config_path = os.path.join(output_dir, "config.json")
[…]

### 2025-05-19 paper/main.tex
QGARS pipeline latex

--- 2025-05-15 paper/IEEE-conference-template-062824/main.tex
@@
 \usepackage{hyperref}
+
+% Add TikZ package and necessary libraries
+\usepackage{tikz}
+\usetikzlibrary{arrows.meta,positioning,fit,backgrounds,shapes.geometric}
+
+% Add circuitikz package for circuit diagrams
+\usepackage{circuitikz}
+
+% Add float control parameters to help with figure placement
+\renewcommand{\floatpagefraction}{0.8}
+\renewcommand{\topfraction}{0.8}
+\renewcommand{\bottomfraction}{0.8}
+\renewcommand{\textfraction}{0.1}
+\setcounter{totalnumber}{50}
+\setcounter{topnumber}{50}
+\setcounter{bottomnumber}{50}
+
+% Force LaTeX to place figures earlier
+\renewcommand{\dblfloatpagefraction}{0.7}
+\renewcommand{\dbltopfraction}{0.8}
 
@@
 \section{Introduction}
+
 \subsection{Background and Motivation}
+
+In this paper, we introduce the Quantum Guided Autoencoder with Reservoir Surrogate (QGARS) architecture, illustrated in Fig.~\ref{fig:system_architecture}.
+This approach combines classical deep learning with quantum reservoir computing to enable efficient medical image classification.
+
 \subsection{Challenges in Quantum-Classical Hybrid Systems}

### 2025-05-20 src/guided_autoencoder.py
experimental results and data analysis

@@ -617,7 +617,7 @@ def train_guided_autoencoder(
     
     for epoch in iterator:
         # Update quantum embeddings and surrogate periodically
-        if epoch % quantum_update_frequency == 0:
+        if (epoch % quantum_update_frequency == 0) and (epoch > 0):
             if verbose:
                 tqdm.write(f"Epoch {epoch+1}: Updating quantum embeddings and surrogate model...")
             

### 2025-05-20 src/parameter_sweep/run_experiments.py

@@ -24,7 +24,7 @@ EXPERIMENT_PROFILES = {
         }
     },
     
-    # RUNNING THIS ONE NOW
+    # DONE, 0.7 yielded the best results
     "guided_autoencoder_lambda": {
         "description": "Sweep over guided autoencoder lambda parameter",
         "param_grid": {
@@ -49,14 +49,14 @@ EXPERIMENT_PROFILES = {
         }
     },
     
-    # WILL RUN ON RTX3095
+    # RUNNING ON RTX3095
     # SHOULD SHOW THAT THE MORE YOU QUERY THE RESERVOIR THE BETTER
     "guided_autoencoder_update_frequency": {
         "description": "Sweep over guided autoencoder update frequency",
         "param_grid": {
             "reduction_method": ["guided_autoencoder"],
             "guided_lambda": [0.7], # FIX WITH THE ONE THAT WORKED BEST IN THE PREVIOUS SWEEP
-            "quantum_update_frequency": [1, 3, 5, 10, 25, 50],
+            "quantum_update_frequency": [1, 3, 5, 7, 10, 15, 25, 0],
             "dim_reduction": [12], # PEAK PERFORMANCE
             "autoencoder_regularization": [1e-5] # FIXED FROM THE PREVIOUS STEPS
         }

### 2025-05-20 src/config_manager.py
fixes

@@ -106,13 +106,13 @@ class ConfigManager:
         return {
             "dataset_type": "cvc_clinic_db_patches",
             "data_dir": None,
-            "target_size": [128, 128],
+            "target_size": [28, 28],
             "split_ratio": 0.8,
             "reduction_method": "autoencoder",
             "dim_reduction": 12,
             "num_examples": 10000,
             "num_test_examples": 400,
-            "guided_lambda": 0.3,
+            "guided_lambda": 0.7,
             "quantum_update_frequency": 1,
             "guided_batch_size": 32,
             "autoencoder_epochs": 50,

### 2025-05-20 src/guided_autoencoder.py

@@ -617,7 +617,7 @@ def train_guided_autoencoder(
     
     for epoch in iterator:
         # Update quantum embeddings and surrogate periodically
-        if (epoch % quantum_update_frequency == 0) and (epoch > 0):
+        if (epoch % quantum_update_frequency == 0):
             if verbose:
                 tqdm.write(f"Epoch {epoch+1}: Updating quantum embeddings and surrogate model...")
             

### 2025-05-20 src/main.py

@@ -120,6 +120,10 @@ def main(args: Optional[argparse.Namespace] = None) -> Tuple[Dict[str, Tuple[np.
     method_name = args.reduction_method.lower()
     reduction_name = method_name.upper()  # For display in result labels
     
+    # If there is no lambda or queries to the reservoir, it's just an autoencoder
+    if args.reduction_method == "guided_autoencoder" and (args.guided_lambda == 0 or args.quantum_update_frequency == 0):
+        method_name = "autoencoder" 
+    
     # Clean memory between runs to avoid cache issues
     import gc
     gc.collect()

### 2025-05-20 paper/main.tex
clean slate

@@
 \begin{abstract}
-This paper introduces a novel quantum-classical hybrid system for medical image classification using neutral atom quantum processors.
-We present the Quantum Guided Autoencoder with Reservoir Surrogate (QGARS), which addresses the gradient barrier problem in quantum-classical hybrid learning by introducing a differentiable surrogate model that enables end-to-end training.
-Our approach leverages quantum reservoir computing principles with Rydberg atom arrays while optimizing feature encoding through a specialized autoencoder that jointly minimizes reconstruction error and maximizes quantum classification performance.
-We evaluate our approach on polyp detection and classification tasks, demonstrating superior performance compared to traditional dimensionality reduction techniques and standard quantum reservoir computing implementations.
-We further present detailed ablation studies analyzing the impact of various quantum parameters, guided learning coefficients, and surrogate model architectures.
-Our results suggest that quantum guidance can significantly enhance feature encoding for quantum processing, pointing toward a practical pathway for quantum advantage in medical image analysis.
 \end{abstract}
@@
 \subsection{Background and Motivation}
-
-In this paper, we introduce the Quantum Guided Autoencoder with Reservoir Surrogate (QGARS) architecture, illustrated in Fig.~\ref{fig:system_architecture}.
-This approach combines classical deep learning with quantum reservoir computing to enable efficient medical image classification.
 

### 2025-05-26 paper/main.tex
advancements

@@
 \usepackage{cite}
+
 \usepackage{amsmath,amssymb,amsfonts}
@@
 should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
 }
@@
 \begin{abstract}
+We introduce a quantum-classical hybrid 
+architecture for medical image classification 
+based on neutral atom quantum processors. 
+This approach is designed to address the challenges of 
+medical imaging, with a particular focus on tasks such 
+as polyp detection and classification.
+By integrating an autoencoder guided by a quantum 
+reservoir, the pipeline learns compact and discriminative 
+representations of image data that are also well-suited for quantum reservoir 
+computing. To overcome the non-differentiability of quantum 
+measurements, we circumvent this `gradient barrier' 
+by incorporating a differentiable surrogate model that simulates the 
+behaviour of the quantum layer, enabling end-to-end backpropagation. 
+The guided training process jointly optimizes for both image 
+reconstruction and classification accuracy, ensuring that the latent 
+representations are both meaningful and effective for quantum processing. 
+In our implementation, image data is encoded as atom detuning parameters 
+in a Rydberg Hamiltonian, and quantum embeddings are obtained through expectation 
+values. These embeddings are then passed to a linear classifier, 
+enabling faster training and inference compared to deep classical networks.
+Our experiments show that this method outperforms traditional 
+approaches using PCA or unguided autoencoders. We also conduct 
+ablation studies to evaluate the impact of quantum and training 
+parameters, demonstrating the robustness and flexibility of the 
+proposed pipeline for real-world medical imaging applications, 
+even in the NISQ era.
 \end{abstract}
@@
 Quantum Machine Learning, Hybrid Quantum-Classical Algorithms, 
-Medical Image Classification
+Medical Image Classification, Quantum Surrogate Models
 \end{IEEEkeywords}
@@
 \subsection{Background and Motivation}
+Advances in medical imaging have significantly improved 
+disease diagnosis and treatment planning. For conditions 
+like colorectal cancer, early detection of polyps through 
+colonoscopy image analysis is critical for reducing mortality \cite{estevaGuideDeepLearning2019a}. 
+Deep learning techniques, especially autoencoders, are widely 
+used to extract compressed, informative features from 
+high-dimensional images for classification and segmentation 
+tasks \cite{bengioLearningDeepArchitecturesa}. However, classical neural networks may struggle 
+to capture intricate correlations in complex medical data.
 
-\subsection{Challenges in Quantum-Classical Hybrid Systems}
+Quantum computing offers novel opportunities 
+for machine learning, particularly through quantum reservoir 
+computing (QRC), where a physical quantum system processes 
+classical inputs into high-dimensional nonlinear embeddings 
+\cite{tanakaRecentAdvancesPhysical2019,fujiiHarnessingDisorderedQuantum2017}. Recent works show that analog quantum systems, 
+such as neutral-atom platforms, can serve as untrained 
+reservoirs with rich dynamics for temporal and pattern 
+recognition tasks \cite{domingoOptimalQuantumReservoir2022,kornjavcaLargescaleQuantumReservoir2024}. 
+In hybrid approaches, a classical encoder compresses 
+image data, and a quantum reservoir expands 
+the encoded features into a higher-dimensional space, 
+potentially boosting classification performance.
+
+A major challenge in such hybrid quantum-classical 
+models is the non-differentiability of quantum measurements, 
+which obstructs gradient-based optimization. Additionally, 
+tuning quantum parameters can suffer from barren plateaus, where 
+gradients vanish in high-dimensional Hilbert spaces \cite{mccleanBarrenPlateausQuantum2018}. 
+To address this, we introduce a classical neural surrogate 
+that emulates the quantum reservoir's input-output behavior. 
+This surrogate enables end-to-end training via backpropagation, 
+while the quantum system remains fixed and non-trainable.
+
 \subsection{Contributions of This Work}
+We propose a quantum-guided autoencoder architecture that 
+integrates a classical image encoder with a neutral-atom
+quantum reservoir.
+
+A classical surrogate network of the reservoir itself enables 
+gradient flow through the whole model during training.
+
+The model is evaluated and compared to classical benchmarks on three 
+different datasets: 
+\begin{enumerate}
+    \item A synthetic dataset of polyp images, generated to simulate realistic medical imaging scenarios.
+    \item Real image patches extracted from the CVC-ClinicDB dataset, a well-known benchmark for polyp detection.
+    \item A reduced version of the MNIST dataset, containing only the digits 0 and 1, suitable for binary classification tasks.
+\end{enumerate}
+
+Our results illustrate the viability of QRC for real-world medical 
+tasks and offer a scalable path to hybrid quantum-classical learning,
+even in the noisy intermediate-scale quantum (NISQ) era.
 
@@
 \section{Background}
+
+\subsection{Principles of Reservoir Computing}
+Reservoir computing is a computational framework 
+derived from recurrent neural networks (RNNs). It 
+involves a fixed, high-dimensional dynamical system—the 
+reservoir—that projects input data into a rich feature 
+space. Only the output layer is trained, simplifying 
+the learning process and reducing computational 
+overhead. This approach is particularly effective 
+for time-series prediction and pattern recognition tasks.
+
+Mathematically, let \( u(t) \in \mathbb{R}^m \) be the input at time \( t \),
+\( x(t) \in \mathbb{R}^n \) the reservoir state, and
+\( y(t) \in \mathbb{R}^k \) the output. The reservoir dynamics
+and output are given by:
+\begin{equation}
+    x(t) = f(W_{in} u(t) + W_{res} x(t-1))
+\end{equation}
+\begin{equation}
+    y(t) = W_{out} x(t)
+\end{equation}
+
+%Where ff is a nonlinear activation function, WinWin​ and WW are fixed input and reservoir weight matrices, and WoutWout​ is the trained output weight matrix.
+
+Where \( f \) is a nonlinear activation function,
+\( W_{in} \) and \( W_{res} \) are fixed input and reservoir weight matrices, 
+and \( W_{out} \) is the trained output weight matrix.
+
+% import figure from reservoir_architecture.tex
+\input{reservoir_architecture.tex}
+
+\subsection{Quantum Reservoir Computing}
+Quantum Reservoir Computing (QRC) extends the reservoir 
+computing paradigm into the quantum domain. By leveraging quantum systems' 
+inherent properties, such as superposition and entanglement, QRC aims 
+to enhance computational capabilities. Implementations using quantum 
+oscillators have shown promise in solving complex learning tasks, 
+offering advantages over classical counterparts. Notably, large-scale 
+experiments utilizing neutral-atom analog quantum computers have demonstrated 
+the scalability and effectiveness of QRC in various machine learning applications 
+\cite{kornjavcaLargescaleQuantumReservoir2024}
+
+In QRC, classical input data \( u(t) \) is encoded into quantum states \( |\psi(t)\rangle \),
+which evolve under a fixed Hamiltonian \( H \):
+\begin{equation}
+    |\psi(t+1)\rangle = U |\psi(t)\rangle = e^{-iH\Delta t} |\psi(t)\rangle,
+\end{equation}
+where \( U \) is the unitary evolution operator. Measurements of observables \( \hat{O} \) yield outputs:
+\begin{equation}
+    y(t) = \langle \psi(t) | \hat{O} | \psi(t) \rangle.
+\end{equation}
+The output weights are trained classically, while the quantum reservoir remains fixed.
+
 \subsection{Quantum Computing with Neutral Atoms}
-\subsection{Principles of Reservoir Computing}
-\subsection{Quantum Reservoir Computing}
+
+% TALK ABOUT HAMILTONIAN
+
 \subsection{Dimensionality Reduction for Image Data}
@@
 \subsubsection{Autoencoder Architectures}
+\subsection{Quantum-Guided Autoencoding}
 
@@
 %============================================
-\section*{References}
+\bibliographystyle{IEEEtran}
+\bibliography{references}
 

### 2025-05-27 paper/autoencoder_architecture.tex
advancements

\begin{figure}[!t]
    \centering
    \resizebox{\columnwidth}{!}{%
        \begin{circuitikz}
            \tikzstyle{every node}=[font=\LARGE]
            \draw[ fill={rgb,255:red,246; green,245; blue,244} ] (4.25,11.75) rectangle (5.5,10.5);
            \draw[ fill={rgb,255:red,246; green,245; blue,244} ] (4.75,11.25) rectangle (6,10);
            \node[font=\LARGE] at (5,12.25) {x};
            \draw[->, >=Stealth] (6,11) -- (8,11);
            \draw(8,12.25) rectangle  node {\Huge $f_{\theta}(x)$} (10.5,9.75);
            \node[font=\large] at (9.5,11.25) {};
            \node[font=\large] at (9,10.5) {};
            \node[font=\large] at (10.5,11.25) {};
            \draw[short] (13.75,11.5) -- (10.5,12.25);
            \draw[short] (10.5,9.75) -- (13.75,10.5);
            \draw(13.75,11.75) rectangle  node {\LARGE z} (14.5,10);
            \draw(17.75,12.25) rectangle  node {\Huge $g_{\phi}(z)$} (20.25,9.75);
            \draw[short] (17.75,12.25) -- (14.5,11.5);
            \draw[short] (14.5,10.5) -- (17.75,9.75);
            \node[font=\Large] at (14,13.75) {Autoencoder};
            \node[font=\large] at (10,12.75) {Encoder Network};
            \node[font=\large] at (18.25,12.75) {Decoder Network};
            \draw[->, >=Stealth] (20.25,11) -- (22.25,11);
            \draw[ fill={rgb,255:red,154; green,153; blue,150} ] (22.75,11.75) rectangle (24,10.5);
            \draw[ fill={rgb,255:red,154; green,153; blue,150} ] (22.25,11.25) rectangle (23.5,10);
            \node[font=\LARGE] at (23,12.25) {$\hat{x}$};
            \draw[<->, >=Stealth, dashed] (5.25,11.75) .. controls (7.5,16.25) and (21.75,15.75) .. (23,11.75)node[pos=0.5, fill=white]{$\mathcal{L}$};
            \node[font=\huge] at (17.75,51.5) {};
            \node[font=\Large] at (14,12.25) {Latent};
        \end{circuitikz}
    }%

\caption{
    Autoencoder architecture. The encoder network $f_{\theta}(x)$ 
    compresses the input $x$ into a latent representation $z$, while the decoder 
    network $g_{\phi}(z)$ reconstructs the input from the latent representation, 
    producing $\hat{x}$. The reconstruction loss $\mathcal{L}$ measures the 
    compresses the input $x$ into a latent representation $z$, while the decoder 
    network $g_{\phi}(z)$ reconstructs the input from the latent representation, 
    producing $\hat{x}$.
}
\label{fig:autoencoder_architecture}
\end{figure}

### 2025-05-27 paper/main.tex
advancements

@@
 \usepackage{hyperref}
+\usepackage{physics}
 
@@
 
-\title{Quantum-Guided Autoencoding for Enhanced Neutral Atom Reservoir Computing in Medical Image Classification\\
+\title{Medical Imaging Classification with Cold-Atom Reservoir Computing using Auto-Encoders and Surrogate-Driven Training\\
 {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured for https://ieeexplore.ieee.org  and
@@
 like colorectal cancer, early detection of polyps through 
-colonoscopy image analysis is critical for reducing mortality \cite{estevaGuideDeepLearning2019a}. 
+colonoscopy image analysis is critical for reducing mortality~\cite{estevaGuideDeepLearning2019a}. 
 Deep learning techniques, especially autoencoders, are widely 
@@
 high-dimensional images for classification and segmentation 
-tasks \cite{bengioLearningDeepArchitecturesa}. However, classical neural networks may struggle 
+tasks~\cite{bengioLearningDeepArchitecturesa}. However, classical neural networks may struggle 
 to capture intricate correlations in complex medical data.
@@
 computing (QRC), where a physical quantum system processes 
-classical inputs into high-dimensional nonlinear embeddings 
-\cite{tanakaRecentAdvancesPhysical2019,fujiiHarnessingDisorderedQuantum2017}. Recent works show that analog quantum systems, 
+classical inputs into high-dimensional nonlinear embeddings~\cite{tanakaRecentAdvancesPhysical2019,fujiiHarnessingDisorderedQuantum2017}. Recent works show that analog quantum systems, 
 such as neutral-atom platforms, can serve as untrained 
@@
 
-% import figure from reservoir_architecture.tex
+A diagram of a typical reservoir computing architecture is shown in `Fig.-\ref{fig:reservoir_architecture}'.
+
 \input{reservoir_architecture.tex}
@@
 experiments utilizing neutral-atom analog quantum computers have demonstrated 
-the scalability and effectiveness of QRC in various machine learning applications 
-\cite{kornjavcaLargescaleQuantumReservoir2024}
+the scalability and effectiveness of QRC in various machine learning applications~\cite{kornjavcaLargescaleQuantumReservoir2024}
 
@@
 
-\subsection{Quantum Computing with Neutral Atoms}
+\subsection{Reservoir Computing with Neutral Atoms}
+As quantum Reservoir Computing (QRC) leverages the complex 
+dynamics of quantum systems to process information, 
+extending the classical reservoir computing 
+paradigm into the quantum domain. Neutral atom platforms, 
+particularly those utilizing Rydberg states, 
+have emerged as promising candidates for implementing 
+QRC due to their scalability and controllable interactions.
 
-% TALK ABOUT HAMILTONIAN
+In the work by M. Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}, 
+a large-scale, gradient-free QRC algorithm was 
+developed and experimentally implemented on a 
+neutral-atom analog quantum computer. 
+This system achieved competitive performance across 
+various machine learning tasks, 
+including classification and time-series prediction, 
+demonstrating effective learning with increasing system 
+sizes up to 108 qubits.
+
+The dynamics of such neutral atom systems can be 
+described by the Rydberg Hamiltonian, which captures 
+the essential physics of laser-driven interactions 
+among atoms in Rydberg states. Following Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}
+The Hamiltonian for a system of neutral-atoms is given by:
+
+\begin{equation}
+    \begin{aligned}
+        H(t) &= \dfrac{\Omega(t)}{2}\sum_j \left(\ket{g_j}\bra{r_j}+\ket{r_j}\bra{g_j}\right) \\
+             &{{\quad}} + \sum_{j<k}V_{jk}n_jn_k - \sum_j \left[\Delta_{\mathrm{g}}(t) + \alpha_j\Delta_{\mathrm{l}}(t)\right] n_j,
+    \end{aligned}
+\end{equation}
+
+where \( \Omega(t) \) is the global Rabi drive amplitude between a 
+ground state \( |g_j\rangle \) and a highly-excited Rydberg state 
+\( |r_j\rangle \) of an atom,
+\( n_j = |r_j\rangle \langle r_j| \), 
+\( V_{jk} = C/\|r_j - r_k\|^6 \) describes the van der Waals interactions
+between atoms, and the detuning is split into a global term 
+\( \Delta_{\mathrm{g}}(t) \)
+and a site-dependent term \( \Delta_{\mathrm{l}}(t) \), with site 
+modulation \( \alpha_j \in [0, 1] \).
+
+By initializing the system in a specific state and allowing 
+it to evolve under this Hamiltonian, the resulting quantum 
+state encodes information about the input data. Measurements 
+of observables on this state yield outputs that can be used 
+for tasks such as classification or prediction, with only the 
+final readout layer requiring training.
 
 \subsection{Dimensionality Reduction for Image Data}
+Dimensionality reduction is a crucial preprocessing step in 
+machine learning and data analysis, aiming to reduce the 
+number of input variables in a dataset while preserving as 
+much information as possible. This process enhances 
+computational efficiency, mitigates the curse of 
+dimensionality, and facilitates data visualization
+
 \subsubsection{Principal Component Analysis}
+Principal Component Analysis (PCA)~\cite{shlensTutorialPrincipalComponent2014} is a linear dimensionality 
+reduction technique that transforms a set of correlated 
+variables into a set of uncorrelated variables called 
+principal components. The goal is to capture the maximum 
+variance in the data with the fewest number of components.
+
+PCA is effective for datasets where the principal components 
+align with the directions of maximum variance, but it may not 
+capture complex, nonlinear relationships in the data~\cite{jolliffePrincipalComponentAnalysis2016}.
+
 \subsubsection{Autoencoder Architectures}
-\subsection{Quantum-Guided Autoencoding}
+Autoencoders are a class of artificial neural networks 
+designed to learn efficient codings of input data in an 
+unsupervised manner. They consist of two main parts: an 
+encoder that compresses the input into a latent-space 
+representation, and a decoder that reconstructs the input 
+from this representation.
+
+Given an input \( x \in \mathbb{R}^d \), the encoder maps 
+\( x \) to a latent representation \( z \in \mathbb{R}^k \) 
+(where \( k < d \)):
+
+\begin{equation}
+    z = f_\theta(x)
+\end{equation}
+
+The decoder then reconstructs the input:
+\begin{equation}
+    \hat{x} = g_\phi(z)
+\end{equation}
+The network is trained to minimize the reconstruction loss:
+\begin{equation}
+    \mathcal{L}(x, \hat{x}) = \| x - \hat{x} \|^2
+\end{equation}
+
+This typical architecture of an autoencoder is illustrated in 
+`Fig.-\ref{fig:autoencoder_architecture}'.
+\input{autoencoder_architecture.tex}
+
+Autoencoders can capture complex, nonlinear 
+relationships in the data, making them suitable for tasks 
+like image compression, denoising, and anomaly 
+detection~\cite{hintonReducingDimensionalityData2006, estevaGuideDeepLearning2019a}.
+
+\subsubsection{Quantum-Guided Autoencoding}
+Quantum-Guided Autoencoders integrate quantum 
+computing principles into the autoencoder framework 
+to leverage quantum advantages in processing and 
+representing data. These models aim to perform 
+dimensionality reduction and classification within a 
+single architecture, enhancing performance on complex 
+datasets.
+
+In the Quantum-Guided Autoencoder (QGAE) model, a 
+classical encoder first reduced the dimensionality
+of the input data. The compressed data is then processed 
+by a parametrized quantum circuit, which acts as the decoder
+and classifier. The quantum circuit transforms the 
+input state \( |\psi_{\text{in}}\rangle \) into an
+output state \( |\psi_{\text{out}}\rangle \) using a unitary operation
+\begin{equation}
+    |\psi_{\text{out}}\rangle = U(\theta) |\psi_{\text{in}}\rangle
+\end{equation}
+Measurements on \( |\psi_{\text{out}}\rangle \) yield the final
+classification result. The parameters \( \theta \) are
+optimized to minimize a loss function that combines
+reconstruction error and classification accuracy.
+
+This approach has demonstrated improved performance 
+over traditional methods in tasks such as identifying 
+the Higgs boson in particle collision data, 
+showcasing the potential of quantum-guided models in 
+handling high-dimensional, complex datasets~\cite{belisGuidedQuantumCompression2024}.
 
@@
 \subsection{System Architecture Overview}
+Our proposed pipeline integrates classical and quantum 
+components to perform dimensionality reduction and
+classification. 
+A classical autoencoder compresses
+high-dimensional image data into a lower-dimensional
+latent space. 
+This compressed representation is 
+encoded into a Rydberg Atom chain, which serves as 
+the input to a quantum reservoir. 
+The reservoir will then produce quantum embeddings
+by measuring quantum observables, which capture the
+dynamics of the quantum system in response to the input data.
+
+These quantum embeddings will be used to train a
+surrogate model that aims to 
+approximates the quantum reservoir's behaviour.
+
+The surrogate model's outputs will then pass 
+through a linear classifier, which is trained
+to map the approximated quantume embeddings to class labels.
+
+The autoencoder is trained to minimize in interpolation
+between the data reconstruction's loss function as well 
+as the classification task loss function.
+
+Our approach, allows the model to learn representations that are not only
+compact and informative for the reconstruction of the input data,
+but also well-suited for quantum reservoir computing.
+
+When autoencoder model is trained,
+we can use the trained encoder to compress the train
+data into the latent space, and then pass the 
+compressed data through the real quantum reservoir.
+The obtained embeddings are then used to train a linear
+classifier that maps the quantum embeddings to class labels.
+
+The overall architecture is illustrated in `Fig.-\ref{fig:qgars_pipeline}'.
+
+% input pdf image
+\begin{figure*}[!b]
+    \centering
+    \resizebox{0.9\textwidth}{!}{
+        \includegraphics{qgars_pipeline_cropped.pdf}
+    }
+    \label{fig:qgars_pipeline}
+    \caption{Overview of the Quantum-Guided Autoencoder Reservoir Computing System.}
+\end{figure*}
+
 \subsection{Quantum Guided Autoencoder}
+The Quantum Guided Autoencoder (QGA) combines classical 
+preprocessing with quantum processing to leverage 
+the strengths of both paradigms. The classical encoder 
+reduces the dimensionality of the input data, facilitating 
+efficient quantum processing. The quantum circuit 
+then processes the compressed data, aiming to reconstruct 
+the original input and perform classification simultaneously.
+
 \subsubsection{Loss Function Design}
+% Loss Function Design
+
+% The QGA is trained to minimize a composite loss function that balances reconstruction fidelity and classification accuracy:
+% L=λrec⋅Lrec+λcls⋅Lcls,
+% L=λrec​⋅Lrec​+λcls​⋅Lcls​,
+
+% where:
+
+%     Lrec=∥x−x^∥2Lrec​=∥x−x^∥2 is the reconstruction loss, measuring the mean squared error between the input xx and its reconstruction x^x^.
+
+%     Lcls=−∑iyilog⁡(y^i)Lcls​=−∑i​yi​log(y^​i​) is the classification loss, computed as the cross-entropy between the true labels yiyi​ and the predicted probabilities y^iy^​i​.
+
+%     λrecλrec​ and λclsλcls​ are hyperparameters that control the trade-off between reconstruction and classification objectives [2].
+
+The QGA is trained to minimize a composite loss 
+function that balances reconstruction fidelity and 
+classification accuracy:
+
+\begin{equation}
+    \mathcal{L} = (1-\lambda) \cdot \mathcal{L}_R + \lambda \cdot \mathcal{L}_{C},
+\end{equation}
+where:
+\begin{itemize}
+    \item \( \mathcal{L}_R = \| x - \hat{x} \|^2 \) is the reconstruction loss, measuring the mean squared error between the input \( x \) and its reconstruction \( \hat{x} \).
+    \item \( \mathcal{L}_C = -\sum_i y_i \log(\hat{y}_i) \) is the classification loss, computed as the cross-entropy between the true labels \( y_i \) and the predicted probabilities \( \hat{y}_i \).
+    \item \( 0 < \lambda < 1 \) is a hyperparameter that controls the trade-off between reconstruction and classification objectives.
+\end{itemize}
+
 \subsubsection{Balancing Reconstruction and Classification}
+
 \subsection{The Gradient Barrier Problem}

### 2025-05-27 paper/main.tex
advancements

@@
 \usepackage{physics}
+\usepackage{float}
 
@@
 \title{Medical Imaging Classification with Cold-Atom Reservoir Computing using Auto-Encoders and Surrogate-Driven Training\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured for https://ieeexplore.ieee.org  and
-should not be used}
+% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured for https://ieeexplore.ieee.org  and
+% should not be used}
 }
@@
     \end{aligned}
+    \label{eq:rydberg_hamiltonian}
 \end{equation}
@@
 `Fig.-\ref{fig:autoencoder_architecture}'.
+
 \input{autoencoder_architecture.tex}
@@
 Our proposed pipeline integrates classical and quantum 
-components to perform dimensionality reduction and
-classification. 
-A classical autoencoder compresses
-high-dimensional image data into a lower-dimensional
-latent space. 
-This compressed representation is 
-encoded into a Rydberg Atom chain, which serves as 
-the input to a quantum reservoir. 
-The reservoir will then produce quantum embeddings
-by measuring quantum observables, which capture the
-dynamics of the quantum system in response to the input data.
+components for dimensionality reduction and classification. 
+A classical autoencoder compresses high-dimensional image 
+data into a lower-dimensional latent space. This latent 
+representation is encoded into a Rydberg atom chain, 
+serving as input to a quantum reservoir. The reservoir's 
+dynamics, governed by the Rydberg Hamiltonian, produce 
+quantum embeddings through measurements of quantum observables.
 
-These quantum embeddings will be used to train a
-surrogate model that aims to 
-approximates the quantum reservoir's behaviour.
+To facilitate end-to-end training, we employ a differentiable 
+surrogate model that approximates the quantum reservoir's 
+behavior, enabling gradient flow from the classification 
+loss back to the autoencoder. The surrogate's outputs feed 
+into a linear classifier trained to map quantum embeddings to class labels.
 
-The surrogate model's outputs will then pass 
-through a linear classifier, which is trained
-to map the approximated quantume embeddings to class labels.
+The autoencoder is trained to minimize a combined loss 
+function that interpolates between reconstruction loss and 
+classification loss, promoting representations that are both 
+informative for input reconstruction and effective for quantum 
+reservoir computing.
 
-The autoencoder is trained to minimize in interpolation
-between the data reconstruction's loss function as well 
-as the classification task loss function.
-
-Our approach, allows the model to learn representations that are not only
-compact and informative for the reconstruction of the input data,
-but also well-suited for quantum reservoir computing.
-
-When autoencoder model is trained,
-we can use the trained encoder to compress the train
-data into the latent space, and then pass the 
-compressed data through the real quantum reservoir.
-The obtained embeddings are then used to train a linear
-classifier that maps the quantum embeddings to class labels.
+Once trained, the encoder compresses data into the latent 
+space, which is then processed by the actual quantum reservoir. 
+The resulting quantum embeddings are used to train a linear 
+classifier for final predictions.
 
@@
 
-% input pdf image
-\begin{figure*}[!b]
+\begin{figure*}[!h]
     \centering
@@
 \subsubsection{Loss Function Design}
-% Loss Function Design
-
-% The QGA is trained to minimize a composite loss function that balances reconstruction fidelity and classification accuracy:
-% L=λrec⋅Lrec+λcls⋅Lcls,
-% L=λrec​⋅Lrec​+λcls​⋅Lcls​,
-
-% where:
-
-%     Lrec=∥x−x^∥2Lrec​=∥x−x^∥2 is the reconstruction loss, measuring the mean squared error between the input xx and its reconstruction x^x^.
-
-%     Lcls=−∑iyilog⁡(y^i)Lcls​=−∑i​yi​log(y^​i​) is the classification loss, computed as the cross-entropy between the true labels yiyi​ and the predicted probabilities y^iy^​i​.
-
-%     λrecλrec​ and λclsλcls​ are hyperparameters that control the trade-off between reconstruction and classification objectives [2].
-
 The QGA is trained to minimize a composite loss 
@@
 
-\subsubsection{Balancing Reconstruction and Classification}
+%\subsubsection{Balancing Reconstruction and Classification}
 
 \subsection{The Gradient Barrier Problem}
+Hybrid quantum-classical models, such as the Quantum Guided Autoencoder 
+(QGAE), aim to leverage quantum computational advantages within 
+classical machine learning frameworks. However, a significant 
+challenge arises due to the non-differentiable nature of quantum 
+operations, commonly referred to as the `gradient barrier'.
+
+In classical neural networks, training relies on backpropagation, 
+which requires the computation of gradients through all components 
+of the network. In hybrid models, the inclusion of quantum 
+layers introduces operations that are inherently non-differentiable:
+
+\begin{itemize}
+    \item \textbf{Quantum State Preparation}: 
+    Encoding classical data into 
+    quantum states often involves operations 
+    that are not smoothly differentiable with 
+    respect to the input data.
+
+    \item \textbf{Quantum Measurements}:
+    Observing quantum states collapses them, 
+    introducing stochasticity and breaking the 
+    deterministic gradient flow required 
+    for backpropagation.
+\end{itemize}
+
+These aspects hinder the direct application of 
+gradient-based optimization methods across the 
+quantum-classical boundary, obstructing end-to-end 
+training of hybrid models.
+
 \subsection{Surrogate Modeling for Quantum Layers}
-\subsubsection{Architecture and Training}
+To circumvent the gradient barrier, surrogate models 
+have been proposed. These are classical, differentiable 
+models trained to approximate the input-output behavior 
+of quantum circuits. By replacing the non-differentiable 
+quantum components with their surrogate counterparts during 
+training, gradients can be propagated through the entire network, 
+enabling end-to-end optimization.
+
+Recent studies have demonstrated the efficacy of 
+surrogate models in mitigating the effects of barren plateaus 
+and facilitating the training of hybrid quantum-classical models. 
+These approaches allow for the practical implementation of 
+quantum-enhanced machine learning models by leveraging classical 
+optimization techniques while still capturing quantum computational 
+advantages~\cite{xieQuantumSurrogateDrivenImage2025b}
+
+\subsubsection{Architecture}
+In our pipeline, the surrogate model is implemented as a 
+feedforward neural network designed to learn the mapping 
+from the autoencoder's latent space to the quantum 
+embeddings produced by the quantum reservoir.
+
+Its architecture comprises multiple fully connected 
+layers with nonlinear activation functions, facilitating 
+the approximation of the complex, non-linear 
+transformations inherent in quantum dynamics.
+
+This approach draws inspiration from recent studies 
+demonstrating the effectiveness of surrogate models 
+in approximating quantum circuit behaviors, thereby 
+facilitating efficient training of hybrid 
+models.~\cite{xieQuantumSurrogateDrivenImage2025b,schreiberClassicalSurrogatesQuantum2023}
+
+\subsubsection{Training Procedure}
+\begin{itemize}
+    \item Data Input: 
+    Feed a batch of latent data \( z \) from the autoencoder 
+    into the real quantum reservoir, recording the resulting quantum 
+    embeddings.
+
+    \item Surrogate Training:
+    Train the surrogate model using the recorded quantum
+    embeddings as targets, minimizing the loss function:
+    \begin{equation}
+        \mathcal{L}_{\text{surrogate}} = \| \hat{y} - y \|^2,
+    \end{equation}
+    where \( \hat{y} \) are the surrogate model's predictions
+    and \( y \) are the actual quantum embeddings.
+
+    \item Regular Updates:
+    Periodically update the surrogate model during training
+    to insure it accurately reflects the quantum reservoir's
+    behavior. This can be done after a fixed number of epochs
+    or when the surrogate's performance on a validation set
+    reaches a certain threshold.
+\end{itemize}
+
 \subsubsection{Gradient Flow Through Surrogate Models}
+By integrating the surrogate model into the training pipeline, 
+we establish a continuous computational graph from the 
+output layer back to the input layer, enabling the use of 
+gradient-based optimization techniques. The surrogate model 
+serves as a differentiable proxy for the quantum reservoir, 
+allowing the classification loss to influence the 
+autoencoder's parameters effectively.
+
+As a results, we not only achieve end-to-end training of the
+hybrid model but also makes it computationally efficient,
+as we reduce the need for frequent quantum measurements.
+
 \subsection{Rydberg Hamiltonian and Quantum Dynamics}
-\subsection{Data Encoding Schemes}
-\subsection{Quantum Readout Methods}
-\subsubsection{Single-atom Measurements}
-\subsubsection{Two-atom Correlations}
-\subsubsection{Three-atom Correlations}
+The quantum reservoir in our architecture is implemented
+using a neutral-atom system, specifically utilizing
+Rydberg atoms to create a programmable quantum reservoir.
+The dynamics of the system are governed by the Rydberg Hamiltonian,
+which describes the interactions between atoms in Rydberg states~\ref{eq:rydberg_hamiltonian}.
+
+\subsubsection{Data Encoding Schemes}
+The input data is encoded into the quantum system by mapping
+the latent representations from the autoencoder to the
+detuning parameters of the Rydberg Hamiltonian.
+This encoding process involves adjusting the detuning
+parameters \( \Delta_{\mathrm{l}}(t) \) for each atom in the chain,
+allowing the system to represent the input data in a way that
+is compatible with the quantum dynamics of the reservoir.
+
+\subsubsection{Quantum Readout Methods}
+The quantum reservoir's output is obtained by probing, 
+in successive time steps, the state of the atoms 
+after the systems evolution up to a certain time \( t \).
+
+The readout is performed by measuring specific observables,
+such as:
+\begin{itemize}
+    \item \textbf{Single-atom Measurements}: 
+    \( \langle Z_i \rangle \) for each atom \( i \), where \( Z_i \) is the Pauli-Z operator.
+    \item \textbf{Two-atom Correlations}: 
+    \( \langle Z_i Z_j \rangle \) for pairs of atoms \( (i, j) \), capturing correlations between their states.
+    \item \textbf{Three-atom Correlations}:
+    \( \langle Z_i Z_j Z_k \rangle \) for triplets of atoms \( (i, j, k) \), providing higher-order correlations.
+\end{itemize}
 
@@
 %============================================
-% THEORETICAL ANALYSIS
-%============================================
-\section{Theoretical Analysis}
-\subsection{Information Encoding in Quantum Reservoirs}
-\subsection{Gradient Flow in Quantum-Classical Hybrid Systems}
-\subsection{Computational Complexity}
-\subsection{Quantum Resource Requirements}
-
-%============================================
 % LIMITATIONS AND FUTURE WORK

### 2025-05-27 paper/main.tex
advancements

@@
     \mathcal{L} = (1-\lambda) \cdot \mathcal{L}_R + \lambda \cdot \mathcal{L}_{C},
+    \label{eq:composed_loss_function}
 \end{equation}
@@
 \subsection{Datasets}
-\subsection{Implementation Details}
-\subsubsection{Quantum Simulation Parameters}
+We evaluate our proposed architecture on three primary datasets:
+\begin{enumerate}
+    \item \textbf{Synthetic Polyp Dataset:} 
+    A synthetic dataset of polyp images generated to simulate 
+    realistic medical imaging scenarios.
+
+    \item \textbf{CVC-ClinicDB Dataset:} 
+    Real image patches extracted from the CVC-ClinicDB 
+    dataset, a well-known benchmark for polyp detection. 
+    This dataset provides a diverse set of polyp images 
+    with varying characteristics, allowing for robust 
+    evaluation of classification performance.
+
+    \item \textbf{MNIST Dataset:} 
+    A reduced version of the MNIST dataset, 
+    containing only the digits 0 and 1, suitable 
+    for binary classification tasks. This dataset 
+    serves as a controlled environment to assess 
+    the model's performance on simpler classification tasks.
+\end{enumerate}
+
+\subsection{Feature Reduction Methods}
+We implemented and compared three distinct feature reduction approaches:
+\subsubsection{Principal Component Analysis (PCA)}
+PCA was employed as a baseline linear 
+dimensionality reduction technique. 
+We extracted the top-k principal components 
+(typically 4-12) from the flattened image data, 
+which were then used as inputs to both classical and 
+quantum-enhanced classifiers.
+
+\subsubsection{Autoencoder}
+We implemented a neural network-based autoencoder with the following architecture:
+
+\begin{itemize}
+    \item An encoder network that compresses
+    the input to a lower-dimensional latent space.
+
+    \item A decoder network that reconstructs 
+    the original input from the latent representation
+
+    \item Batch normalization and dropout layers 
+    to improve training stability and prevent overfitting.
+\end{itemize}
+
+The autoencoder was trained to minimize reconstruction 
+loss using the Adam optimizer with learning rates 
+between 0.001—0.01 and weight decay for regularization.
+
+\subsubsection{Quantum-Guided Autoencoder}
+We introduced a quantum-guided autoencoder 
+that incorporates feedback from the quantum 
+reservoir during training. This approach combines:
+
+\begin{itemize}
+    \item A standard autoencoder architecture
+    \item A linear classifier that maps quantum embeddings to class labels.
+    \item A classification loss computed on quantum embeddings
+    \item A weighting parameter $\lambda$ to control the trade-off between 
+    reconstruction and classification objectives~\ref{eq:composed_loss_function}.
+    \item A differentiable surrogate model that approximates the quantum reservoir's behavior,
+    enabling end-to-end backpropagation through the entire pipeline. The frequency 
+    of quantum updates is controlled by a hyperparameter, allowing for
+    flexible integration of quantum dynamics into the training process.
+\end{itemize}
+
+\subsubsection{Quantum Reservoir Computing Layer}
+Our quantum reservoir computing implementation uses 
+a Rydberg atom simulator with the following parameterizable
+components:
+
+\begin{itemize}
+    \item \textbf{Atom Chain Length:} 
+    The number of atoms in the chain, which can be adjusted 
+    to control the reservoir's capacity and complexity.
+
+    \item \textbf{Rabi Frequency:} 
+    The global Rabi drive amplitude, which influences the 
+    strength of interactions between atoms.
+
+    \item \textbf{Detuning Parameters:} 
+    Site-dependent detuning parameters that encode the input data.
+
+    \item \textbf{Measurement Strategy:} 
+    The specific observables measured to obtain quantum embeddings.
+
+    \item \textbf{Encoding Scale:}
+    The scale of the input data encoding, which can be adjusted
+    to optimize the quantum reservoir's response to the input data.
+
+    \item \textbf{Time steps:}
+    The number of time steps for which the quantum reservoir evolves,
+    allowing for the capture of temporal dynamics in the data.
+\end{itemize}
+
 \subsubsection{Classical Network Architectures}
+After the traning process of the quantum-guided autoencoder,
+the following quantum embeddings are passed to a linear classifier.
+The classifier is a simple feedforward neural network with the following architecture:
+\begin{itemize}
+    \item A single hidden layer with a ReLU activation function.
+    \item Cross-entropy loss for training, optimized using the Adam optimizer.
+\end{itemize}
+
 \subsection{Comparison Methods}
+We benchmarked our model against some
+classical methods, including:
+\begin{itemize}
+    \item \textbf{PCA + Linear Classifier:} 
+    A linear classifier trained on the PCA-reduced features.
+
+    \item \textbf{Autoencoder + Linear Classifier:} 
+    A linear classifier trained on the latent representations from the autoencoder.
+
+    \item \textbf{PCA + Neural Network:}
+    A fully connected neural network trained on the PCA-reduced features.
+
+    \item \textbf{Autoencoder + Neural Network:}
+    A fully connected neural network trained on the latent representations from the autoencoder.
+\end{itemize}
+
 \subsection{Performance Metrics}
+We evaluated the performance of our model using the following metrics:
+\begin{itemize}
+    \item \textbf{Classification Accuracy:} 
+    The percentage of correctly classified instances in the test set.
+
+    \item \textbf{F1 Score:} 
+    The harmonic mean of precision and recall, providing a balanced measure of classification performance.
+
+    \item \textbf{Confusion Matrix:} 
+    A matrix summarizing the classification results, showing true positives, false positives, true negatives, and false negatives.
+
+\end{itemize}
+
 \subsection{Parameter Sweep Strategy}
+To optimize the performance of our quantum-guided autoencoder, 
+we conducted a systematic parameter sweep over the following hyperparameters:
+\begin{itemize}
+    \item \textbf{Guided Lambda Parameter (\( \lambda \))}: 
+    This parameter controls the trade-off between reconstruction and classification objectives in the loss function~\ref{eq:composed_loss_function}.
+We tested values ranging from 0.1 to 0.9 in increments of 0.1.
+
+    \item \textbf{Quantum Update Frequency:} 
+    The frequency at which the quantum reservoir is updated during training. We experimented with update frequencies of 1, 5, and 10 epochs.
+
+    \item \textbf{Quantum Parameters:} 
+    Including the number of atoms in the chain, Rabi frequency, and detuning parameters. We varied these parameters to assess their impact on classification performance.
+    \item \textbf{Batch Size:}
+    \item \textbf{Learning Rate:}
+\end{itemize}
 

### 2025-05-27 paper/main.tex
autoencoder formatting issues
Updates from Overleaf

@@
 \documentclass[conference]{IEEEtran}
+\usepackage[utf8]{inputenc}
 \IEEEoverridecommandlockouts
@@
 \usepackage{float}
+\usepackage{inputenc}
 
@@
 
-\author{
-\IEEEauthorblockN{Nuno Batista}
-\IEEEauthorblockA{\textit{Department of Informatics Engineering} \\
-\textit{Faculty of Sciences and Technology, University of Coimbra}\\
-Coimbra, Portugal \\
-\href{mailto:nunomarquesbatista@gmail.com}{nunomarquesbatista@gmail.com}}
+\author{\IEEEauthorblockN{Anonymous Authors}}
+% \author{
+% \IEEEauthorblockN{Nuno Batista}
+% \IEEEauthorblockA{\textit{Department of Informatics Engineering} \\
+% \textit{Faculty of Sciences and Technology, University of Coimbra}\\
+% Coimbra, Portugal \\
+% \href{mailto:nunomarquesbatista@gmail.com}{nunomarquesbatista@gmail.com}}
 
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-}
+% \and
+% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% \textit{name of organization (of Aff.)}\\
+% City, Country \\
+% email address or ORCID}
+% }
 
 \maketitle
+
+%ADD SPACE TO BREATHE
 
@@
 like colorectal cancer, early detection of polyps through 
-colonoscopy image analysis is critical for reducing mortality~\cite{estevaGuideDeepLearning2019a}. 
+colonoscopy image analysis is critical for reducing mortality~\cite{estevaGuideDeepLearning2019}. 
 Deep learning techniques, especially autoencoders, are widely 
@@
 high-dimensional images for classification and segmentation 
-tasks~\cite{bengioLearningDeepArchitecturesa}. However, classical neural networks may struggle 
+tasks~\cite{bengioLearningDeepArchitectures}. However, classical neural networks may struggle 
 to capture intricate correlations in complex medical data.
@@
 
-The model is evaluated and compared to classical benchmarks on three 
-different datasets: 
-\begin{enumerate}
-    \item A synthetic dataset of polyp images, generated to simulate realistic medical imaging scenarios.
-    \item Real image patches extracted from the CVC-ClinicDB dataset, a well-known benchmark for polyp detection.
-    \item A reduced version of the MNIST dataset, containing only the digits 0 and 1, suitable for binary classification tasks.
-\end{enumerate}
-
 Our results illustrate the viability of QRC for real-world medical 
@@
 %============================================
-\section{Background}
+\section{Background and Related Work}
 
@@
 
-A diagram of a typical reservoir computing architecture is shown in `Fig.-\ref{fig:reservoir_architecture}'.
+A diagram of a typical Reservoir Computing architecture is shown in `Fig.-\ref{fig:reservoir_architecture}'.
 
@@
 \subsection{Reservoir Computing with Neutral Atoms}
-As quantum Reservoir Computing (QRC) leverages the complex 
+As Quantum Reservoir Computing (QRC) leverages the complex 
 dynamics of quantum systems to process information, 
-extending the classical reservoir computing 
+extending the Classical Reservoir Computing 
 paradigm into the quantum domain. Neutral atom platforms, 
@@
 among atoms in Rydberg states. Following Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}
-The Hamiltonian for a system of neutral-atoms is given by:
+the Hamiltonian for a system of neutral-atoms is given by:
 
@@
 \end{equation}
-
 where \( \Omega(t) \) is the global Rabi drive amplitude between a 
@@
 like image compression, denoising, and anomaly 
-detection~\cite{hintonReducingDimensionalityData2006, estevaGuideDeepLearning2019a}.
+detection~\cite{hintonReducingDimensionalityData2006, estevaGuideDeepLearning2019}.
 
@@
 \subsection{System Architecture Overview}
+
 Our proposed pipeline integrates classical and quantum 
@@
 quantum embeddings through measurements of quantum observables.
+The overall architecture is illustrated in `Fig.-\ref{fig:qgars_pipeline}'.
 
@@
 classifier for final predictions.
-
-The overall architecture is illustrated in `Fig.-\ref{fig:qgars_pipeline}'.
 
@@
 quantum-classical boundary, obstructing end-to-end 
-training of hybrid models.
+training of hybrid models. The surrogate model 
+solution, proposed in the next section, addresses 
+this problem.
 
@@
 The readout is performed by measuring specific observables,
-such as:
-\begin{itemize}
-    \item \textbf{Single-atom Measurements}: 
-    \( \langle Z_i \rangle \) for each atom \( i \), where \( Z_i \) is the Pauli-Z operator.
-    \item \textbf{Two-atom Correlations}: 
-    \( \langle Z_i Z_j \rangle \) for pairs of atoms \( (i, j) \), capturing correlations between their states.
-    \item \textbf{Three-atom Correlations}:
-    \( \langle Z_i Z_j Z_k \rangle \) for triplets of atoms \( (i, j, k) \), providing higher-order correlations.
-\end{itemize}
+such as $\langle Z_i \rangle$, $\langle Z_i \, Z_j\rangle$, and $\langle Z_i \, Z_j \, Z_k \rangle$, which measure respectively one-, two- and three-qubit correlations.
+% \begin{itemize}
+%     \item \textbf{Single-atom Measurements}: 
+%     \( \langle Z_i \rangle \) for each atom \( i \), where \( Z_i \) is the Pauli-Z operator.
+%     \item \textbf{Two-atom Correlations}: 
+%     \( \langle Z_i Z_j \rangle \) for pairs of atoms \( (i, j) \), capturing correlations between their states.
+%     \item \textbf{Three-atom Correlations}:
+%     \( \langle Z_i Z_j Z_k \rangle \) for triplets of atoms \( (i, j, k) \), providing higher-order correlations.
+% \end{itemize}
 
@@
 loss using the Adam optimizer with learning rates 
-between 0.001—0.01 and weight decay for regularization.
+between $0.001$--$0.01$ and weight decay for regularization.
 
@@
 We evaluated the performance of our model using the following metrics:
-\begin{itemize}
-    \item \textbf{Classification Accuracy:} 
-    The percentage of correctly classified instances in the test set.
+Classification Accuracy, F1 Score, Confusion Matrix.
 
-    \item \textbf{F1 Score:} 
-    The harmonic mean of precision and recall, providing a balanced measure of classification performance.
+% \begin{itemize}
+%     \item \textbf{Classification Accuracy} 
+%     %The percentage of correctly classified instances in the test set.
 
-    \item \textbf{Confusion Matrix:} 
-    A matrix summarizing the classification results, showing true positives, false positives, true negatives, and false negatives.
+%     \item \textbf{F1 Score} 
+%     %The harmonic mean of precision and recall, providing a balanced measure of %classification performance.
 
-\end{itemize}
+%     \item \textbf{Confusion Matrix} 
+%    % A matrix summarizing the classification results, showing true positives, %false positives, true negatives, and false negatives.
+
+% \end{itemize}
 
@@
 
+\newpage
+\newpage
+\tableofcontents
+
 \end{document}

### 2025-05-27 parameter_sweep/data_analysis_scripts/guided_lambda_analysis.py
a

@@ -18,6 +18,50 @@ def load_sweep_results(filepath: str) -> pd.DataFrame:
     
     return df
 
+# def plot_guided_lambda_effect(df: pd.DataFrame, output_path: Optional[str] = None):
+    # """Create a plot showing guided_lambda vs QRC test accuracy."""
+    # # Set up aesthetics
+    # sns.set_style("whitegrid")
+    
+    # # Sort by guided_lambda value
+    # df_sorted = df.sort_values(by='guided_lambda')
+    
+    # # Convert accuracy to percentage
+    # y_values = df_sorted['QRC_final_test_acc'].values * 100
+    # x_values = df_sorted['guided_lambda'].values
+    
+    # # Create a figure with appropriate size
+    # plt.figure(figsize=(10, 6))
+    
+    # # Plot line connecting points
+    # plt.plot(x_values, y_values, '-', linewidth=2, color='#1f77b4')
+    
+    # # Plot individual points
+    # plt.scatter(x_values, y_values, s=80, color='#1f77b4', zorder=5)
+    
+    # # Add value labels
+    # for i, (x, y) in enumerate(zip(x_values, y_values)):
+    #     plt.annotate(f"{y:.1f}%", (x, y), xytext=(0, 10), 
+    #                 textcoords="offset points", ha='center')
+    
+    # # Labels and title
+    # plt.xlabel('Guided Lambda', fontsize=12)
+    # plt.ylabel('QRC Test Accuracy (%)', fontsize=12)
+    # plt.title("Effect of Guided Lambda on QRC Accuracy", fontsize=14)
+    
+    # # Add grid for better readability
+    # plt.grid(True, linestyle='--', alpha=0.7)
+    
+    # plt.tight_layout()
+    
+    # # Save figure if output path is provided
+    # if output_path:
+    #     os.makedirs(os.path.dirname(output_path), exist_ok=True)
+    #     plt.savefig(output_path, dpi=300, bbox_inches='tight')
+    #     print(f"Figure saved to {output_path}")
+    
+    # plt.show()
+
 def plot_guided_lambda_effect(df: pd.DataFrame, output_path: Optional[str] = None):
     """Create a plot showing guided_lambda vs QRC test accuracy."""
     # Set up aesthetics
@@ -44,8 +88,11 @@ def plot_guided_lambda_effect(df: pd.DataFrame, output_path: Optional[str] = Non
         plt.annotate(f"{y:.1f}%", (x, y), xytext=(0, 10), 
[…]

### 2025-05-27 parameter_sweep/data_analysis_scripts/quantum_update_frequency_analysis.py

@@ -1,0 +1,163 @@
+import os
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+from typing import Optional
+import argparse
+
+def load_sweep_results(filepath: str) -> pd.DataFrame:
+    """Load parameter sweep results from CSV file."""
+    if not os.path.exists(filepath):
+        raise FileNotFoundError(f"Results file not found: {filepath}")
+    
+    try:
+        df = pd.read_csv(filepath)
+        print(f"Loaded {len(df)} experiment results from {filepath}")
+    except Exception as e:
+        print(f"Error loading results: {e}")
+        df = pd.DataFrame()
+    
+    return df
+
+def plot_quantum_update_frequency_effect(df: pd.DataFrame, output_path: Optional[str] = None):
+    """Create a plot showing quantum update frequency vs QRC accuracies."""
+    # Set up aesthetics
+    sns.set_style("whitegrid")
+    
+    # Sort by update frequency value
+    df_sorted = df.sort_values(by='quantum_update_frequency')
+    
+    # Convert accuracies to percentage
+    train_acc = df_sorted['QRC_final_train_acc'].values * 100
+    test_acc = df_sorted['QRC_final_test_acc'].values * 100
+    x_values = df_sorted['quantum_update_frequency'].values
+    
+    # Create a figure with appropriate size
+    plt.figure(figsize=(10, 6))
+    
+    # Plot test accuracy (solid line)
+    plt.plot(x_values, test_acc, '-', linewidth=2, color='#1f77b4', label='Test Accuracy')
+    plt.scatter(x_values, test_acc, s=80, color='#1f77b4', zorder=5)
+    
+    # Plot train accuracy (dashed line)
+    plt.plot(x_values, train_acc, '--', linewidth=2, color='#ff7f0e', label='Training Accuracy')
+    plt.scatter(x_values, train_acc, s=80, color='#ff7f0e', zorder=5, marker='s')
+    
+    # Add value labels for test accuracy
+    for i, (x, y) in enumerate(zip(x_values, test_acc)):
+        plt.annotate(f"{y:.1f}%", (x, y), xytext=(0, 10), 
+                    textcoords="offset points", ha='center')
+    
+    # Special handling for x=0 (no quantum update)
+    if 0 in x_values:
+        plt.axvline(x=0, color='gray', linestyle=':', alpha=0.5)
+        plt.text(0, plt.ylim()[0] + 5, 'No Quantum\nUpdate', 
[…]

### 2025-05-27 parameter_sweep/data_analysis_scripts/time_steps_evolution_time_analysis.py

@@ -1,0 +1,194 @@
+import os
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+import json
+import argparse
+from typing import Optional, Tuple
+
+def load_data(results_file: str, param_grid_file: Optional[str] = None) -> Tuple[pd.DataFrame, dict]:
+    """
+    Load experiment results and parameter grid information.
+    
+    Parameters
+    ----------
+    results_file : str
+        Path to the CSV file with experiment results
+    param_grid_file : Optional[str]
+        Path to the parameter grid JSON file. If None, inferred from results_file.
+        
+    Returns
+    -------
+    Tuple[pd.DataFrame, dict]
+        DataFrame with results and parameter grid dictionary
+    """
+    # Load results CSV
+    if not os.path.exists(results_file):
+        raise FileNotFoundError(f"Results file not found: {results_file}")
+    
+    df = pd.read_csv(results_file)
+    print(f"Loaded {len(df)} experiment results from {results_file}")
+    
+    # Filter only successful runs
+    df = df[df['status'] == 'success']
+    print(f"After filtering, {len(df)} successful experiments remain")
+    
+    # Load parameter grid if provided, otherwise infer from results directory
+    if param_grid_file is None:
+        results_dir = os.path.dirname(results_file)
+        param_grid_file = os.path.join(results_dir, "param_grid.json")
+    
+    if os.path.exists(param_grid_file):
+        with open(param_grid_file, 'r') as f:
+            param_grid = json.load(f)
+        print(f"Loaded parameter grid from {param_grid_file}")
+    else:
+        # If param_grid.json doesn't exist, create a simple grid from unique values in results
+        param_grid = {
+            'time_steps': sorted(df['time_steps'].unique().tolist()),
+            'evolution_time': sorted(df['evolution_time'].unique().tolist())
+        }
+        print("Parameter grid file not found, inferred from results")
+    
+    return df, param_grid
+
+def create_heatmap(df: pd.DataFrame, param_grid: dict, output_path: Optional[str] = None):
+    """
+    Create a heatmap of QRC test accuracy for different time_steps and evolution_time values.
+    
+    Parameters
+    ----------
+    df : pd.DataFrame
+        DataFrame with experiment results
+    param_grid : dict
+        Parameter grid dictionary
[…]

### 2025-05-27 parameter_sweep/run_experiments.py

@@ -30,7 +30,12 @@ EXPERIMENT_PROFILES = {
         "param_grid": {
             "reduction_method": ["guided_autoencoder"],
            # "guided_lambda": [1, 0.95, 0.9, 0.8, 0.7, 0.5, 0.3, 0.1, 0.05, 0], # SWEEP IN LOG SCALE!
-            "guided_lambda": [0.0000001, 0.00001, 0.0001, 0.001, 0.01, 0.1, 0.0, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1],
+           
+           
+           # ADD SYMETRIC
+            "guided_lambda": [0.0000001, 0.000001, 0.00001, 0.0001, 0.001, 0.01, 0.1, 0.0,
+                              0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 
+                              1-0.01, 1-0.001, 1-0.0001, 1-0.00001, 1-0.000001, 1],
             "quantum_update_frequency": [5],
             "dim_reduction": [12]
         }
@@ -50,8 +55,7 @@ EXPERIMENT_PROFILES = {
         }
     },
     
-    # RUNNING ON RTX3095
-    # SHOULD SHOW THAT THE MORE YOU QUERY THE RESERVOIR THE BETTER
+    # DONE 
     "guided_autoencoder_update_frequency": {
         "description": "Sweep over guided autoencoder update frequency",
         "param_grid": {
@@ -85,14 +89,15 @@ EXPERIMENT_PROFILES = {
         }
     },
     
-    # NOT DONE
-    # UNSURE
+    # RUNNING ON CRAY-4
     "shot_noise": {
         "description": "Sweep over number of quantum shots",
         "param_grid": {
             "n_shots": [100, 500, 1000, 2000],
             "readout_type": ["all"],
             "reduction_method": ["guided_autoencoder"],
+            "quantum_update_frequency": [5],
+            "guided_lambda": [0.7], 
             "dim_reduction": [12]
         }
     },

### 2025-05-27 paper/main.tex
draft

@@
 We introduce a quantum-classical hybrid
-architecture for medical image classification 
-based on neutral atom quantum processors. 
+architecture for medical image classification based on neutral-atom quantum processors. 
 This approach is designed to address the challenges of
@@
 %============================================
-% INTRODUCTION
+% hilbert
 %============================================
@@
 computing (QRC), where a physical quantum system processes 
-classical inputs into high-dimensional nonlinear embeddings~\cite{tanakaRecentAdvancesPhysical2019,fujiiHarnessingDisorderedQuantum2017}. Recent works show that analog quantum systems, 
-such as neutral-atom platforms, can serve as untrained 
-reservoirs with rich dynamics for temporal and pattern 
-recognition tasks \cite{domingoOptimalQuantumReservoir2022,kornjavcaLargescaleQuantumReservoir2024}. 
-In hybrid approaches, a classical encoder compresses 
-image data, and a quantum reservoir expands 
-the encoded features into a higher-dimensional space, 
-potentially boosting classification performance.
+classical inputs into high-dimensional nonlinear embeddings~\cite{tanakaRecentAdvancesPhysical2019,fujiiHarnessingDisorderedQuantum2017}.
+Recent works show that analog quantum systems, such as neutral-atom platforms, can serve as untrained reservoirs with rich dynamics for temporal and pattern recognition tasks~\cite{domingoOptimalQuantumReservoir2022,kornjavcaLargescaleQuantumReservoir2024}.
 
-A major challenge in such hybrid quantum-classical 
-models is the non-differentiability of quantum measurements, 
-which obstructs gradient-based optimization. Additionally, 
-tuning quantum parameters can suffer from barren plateaus, where 
-gradients vanish in high-dimensional Hilbert spaces \cite{mccleanBarrenPlateausQuantum2018}. 
-To address this, we introduce a classical neural surrogate 
-that emulates the quantum reservoir's input-output behavior. 
-This surrogate enables end-to-end training via backpropagation, 
-while the quantum system remains fixed and non-trainable.
+In hybrid approaches, a classical encoder compresses image data, and a quantum reservoir expands the encoded features into a higher-dimensional Hilbert space, potentially boosting classification performance, as these expanded features might capture high-order correlations and nonlinearities that classical methods cannot tractably represent.
+
+A major challenge in such hybrid quantum-classical models is the non-differentiability of quantum measurements, which obstructs gradient-based optimization.
+Additionally, tuning quantum parameters can suffer from barren plateaus, where gradients vanish in high-dimensional Hilbert spaces~\cite{mccleanBarrenPlateausQuantum2018}.
+To address this, we introduce a classical neural surrogate that emulates the quantum reservoir's input-output behavior.
+This surrogate enables end-to-end training via backpropagation, while the quantum system remains fixed and non-trainable and is only used to train the surrogate model and in downstream classifiers.
 
 \subsection{Contributions of This Work}
-We propose a quantum-guided autoencoder architecture that 
-integrates a classical image encoder with a neutral-atom
-quantum reservoir.
+We propose a quantum-guided autoencoder architecture that integrates a classical image encoder with a neutral-atom quantum reservoir.
 
@@
 
-\input{reservoir_architecture.tex}
+\input{images/diagrams/reservoir_architecture.tex}
 
 \subsection{Quantum Reservoir Computing}
+
+\subsubsection{Main Differences}
 Quantum Reservoir Computing (QRC) extends the reservoir 
@@
 
-\subsection{Reservoir Computing with Neutral Atoms}
-As Quantum Reservoir Computing (QRC) leverages the complex 
-dynamics of quantum systems to process information, 
-extending the Classical Reservoir Computing 
-paradigm into the quantum domain. Neutral atom platforms, 
-particularly those utilizing Rydberg states, 
-have emerged as promising candidates for implementing 
-QRC due to their scalability and controllable interactions.
+\subsubsection{Quantum Reservoir Computing with Neutral Atoms}
+Neutral atom platforms, particularly those utilizing Rydberg states, have emerged as promising candidates for implementing QRC due to their scalability and controllable interactions.
 
-In the work by M. Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}, 
-a large-scale, gradient-free QRC algorithm was 
-developed and experimentally implemented on a 
-neutral-atom analog quantum computer. 
-This system achieved competitive performance across 
-various machine learning tasks, 
-including classification and time-series prediction, 
-demonstrating effective learning with increasing system 
-sizes up to 108 qubits.
+In the work by M.
+Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}, a large-scale, gradient-free QRC algorithm was developed and experimentally implemented on a neutral-atom analog quantum computer.
+This system achieved competitive performance across various machine learning tasks, including classification and time-series prediction, demonstrating effective learning with increasing system sizes up to $108$ qubits.
 
-The dynamics of such neutral atom systems can be 
-described by the Rydberg Hamiltonian, which captures 
-the essential physics of laser-driven interactions 
-among atoms in Rydberg states. Following Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}
-the Hamiltonian for a system of neutral-atoms is given by:
+The dynamics of such neutral atom systems can be described by the Rydberg Hamiltonian, which captures 
+the essential physics of laser-driven interactions among atoms in Rydberg states.
+Following Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024} the Hamiltonian for a system of neutral-atoms is given by:
 
@@
 
-\input{autoencoder_architecture.tex}
+\input{images/diagrams/autoencoder_architecture.tex}
 
@@
 \subsection{System Architecture Overview}
-
-Our proposed pipeline integrates classical and quantum 
-components for dimensionality reduction and classification. 
-A classical autoencoder compresses high-dimensional image 
-data into a lower-dimensional latent space. This latent 
-representation is encoded into a Rydberg atom chain, 
-serving as input to a quantum reservoir. The reservoir's 
-dynamics, governed by the Rydberg Hamiltonian, produce 
-quantum embeddings through measurements of quantum observables.
+Our proposed pipeline is a Quantum-Guided Autoencoder with a 
+Reservoir Surrogate (QGARS). It is a hybrid architecture that 
+synergizes classical autoencoding techniques with quantum 
+reservoir processing to achieve efficient learning of features
+that work well with the reservoir computing layer whose embeddings
+will be used to perferm a classification task.
+The system is designed to overcome the inherent non-differentiability of quantum operations, which traditionally impede gradient-based training methods.
 The overall architecture is illustrated in `Fig.-\ref{fig:qgars_pipeline}'.
 
-To facilitate end-to-end training, we employ a differentiable 
-surrogate model that approximates the quantum reservoir's 
-behavior, enabling gradient flow from the classification 
-loss back to the autoencoder. The surrogate's outputs feed 
-into a linear classifier trained to map quantum embeddings to class labels.
+In a higher level, the \textbf{architectural components} are the following:
+\begin{itemize}
+    \item \textbf{Classical Autoencoder:}
+        \begin{itemize}
+            \item \textbf{Encoder:}
+            The encoder network \(\varepsilon_\omega(x)\)
+            transforms high-dimensional input data \(\mathbf{x} \in \mathbb{R}^{n}\) into 
+            a lower-dimensional latent representation \(\mathbf{z} \in \mathbb{R}^m\), 
+            where \(m < n\).
+            \item \textbf{Decoder:}
+            The decoder network \(\mathcal{D}_\rho(z)\)
+            Reconstructs the input data from the latent representation, producing
+            \(\mathbf{\hat{x}} \in \mathbb{R}\).
+        \end{itemize}
 
-The autoencoder is trained to minimize a combined loss 
-function that interpolates between reconstruction loss and 
-classification loss, promoting representations that are both 
-informative for input reconstruction and effective for quantum 
-reservoir computing.
+    \item \textbf{Quantum Reservoir Layer:}
+        \begin{itemize}
+            \item \textbf{Parameter Mapping:}
+            \(g(\mathbf{z})\) maps the latent representation \(\mathbf{z}\) to the local detuning frequencies
+            \(\Delta_{\mathrm{i}}\) of the Rydberg Atoms.
+            \item \textbf{Quantum Evolution:}
+            Evolves the system under a Hamiltonian \(H(\Delta)\), resulting in a quantum 
+            state \(|\psi(\Delta)\rangle\).
+            \item \textbf{Measurement:}
+            Performs measurements on the quantum state to obtain embeddings \(\mathbf{q} \in \mathbb{R}^k\), which serve as training targets for the Surrogate Model or as 
+            inputs for the classification task later on.
+        \end{itemize}
 
-Once trained, the encoder compresses data into the latent 
-space, which is then processed by the actual quantum reservoir. 
-The resulting quantum embeddings are used to train a linear 
-classifier for final predictions.
+    \item \textbf{Surrogate Model:}
+    A differentiable classical neural network \( f_{\text{sur}}(\mathbf{z}; \theta_{\text{sur}}) \)
+    where \(\theta_{\text{sur}}\), which are it's parameters, are tuned to approximate the 
+    mapping from \(\mathbf{z}\) to \(\mathbf{q}\), effectively emulating
+    the quantum layer's behaviour.
 
-\begin{figure*}[!h]
+    \item \textbf{Linear Classifier:}
+    A classical neural network that maps the quantum embeddings \(\mathbf{q}\) to 
+    predicted class labels \(\hat{y}\).
+
+\end{itemize}
+The following sections will give a more detailed explanation of each 
+component.
+
+\begin{figure*}[!ht]
     \centering
     \resizebox{0.9\textwidth}{!}{
-        \includegraphics{qgars_pipeline_cropped.pdf}
+        \includegraphics{images/diagrams/qgars_pipeline_cropped.pdf}
     }
@@
 preprocessing with quantum processing to leverage 
-the strengths of both paradigms. The classical encoder 
-reduces the dimensionality of the input data, facilitating 
-efficient quantum processing. The quantum circuit 
-then processes the compressed data, aiming to reconstruct 
-the original input and perform classification simultaneously.
+the strengths of both paradigms.
+
+Ideally, the latent space representation created by the encoder \(\varepsilon_\omega(x)\)
+would be passed through two sections of the model:
+\begin{itemize}
+    \item \textbf{The Decoder Network \(\mathcal{D}_\rho(z)\):}
+    that tries to reconstruct the original input data.
+    \item \textbf{The Quantum Reservoir:}
+    that extracts highly dimensional quantum embeddings and
+    passes them to a linear classifier to perform a classification
+    task and map the embeddings to the image labels.
+\end{itemize}
+However, some problems emerge when using a real quantum reservoir directly during the autoencoder training stage, which will be discussed in a later section.
 
@@
 \begin{itemize}
-    \item \( \mathcal{L}_R = \| x - \hat{x} \|^2 \) is the reconstruction loss, measuring the mean squared error between the input \( x \) and its reconstruction \( \hat{x} \).
-    \item \( \mathcal{L}_C = -\sum_i y_i \log(\hat{y}_i) \) is the classification loss, computed as the cross-entropy between the true labels \( y_i \) and the predicted probabilities \( \hat{y}_i \).
+    \item \( \mathcal{L}_R = \| \mathbf{x} - \hat{\mathbf{x}} \|^2 \) is the reconstruction loss, measuring the mean squared error between the input \( \mathbf{x} \) and its reconstruction \( \mathbf{\hat{x}} \).
+    \item \( \mathcal{L}_C = -\sum_i y_i \log(\hat{y}_i) \) is the classification loss, computed as the cross-entropy between the true labels \( y_i \) and the predicted probabilities \( \hat{y}_i \), which are obtained through a linear mapping of the
+    approximated embeddings \(\mathbf{q}'\).
     \item \( 0 < \lambda < 1 \) is a hyperparameter that controls the trade-off between reconstruction and classification objectives.
 \end{itemize}    
-
-%\subsubsection{Balancing Reconstruction and Classification}
-
 \subsection{The Gradient Barrier Problem}
@@
 (QGAE), aim to leverage quantum computational advantages within 
-classical machine learning frameworks. However, a significant 
+classical machine learning frameworks. However, a significant g
 challenge arises due to the non-differentiable nature of quantum 
@@
 
-\subsubsection{Architecture}
-In our pipeline, the surrogate model is implemented as a 
-feedforward neural network designed to learn the mapping 
+\subsubsection{Surrogate Model Architecture}
+In our pipeline, the surrogate model \( f_{\text{sur}}(\mathbf{z}; \theta_{\text{sur}}) \) 
+is implemented as a feedforward neural network 
+designed to learn the mapping 
 from the autoencoder's latent space to the quantum 
 embeddings produced by the quantum reservoir. 
-
+\begin{equation}
+    f_{\text{sur}}(\mathbf{z}; \theta_{\text{sur}}) = \mathbf{q}' \approx \mathbf{q} = \mathrm{Measure}(|\psi(g(\mathbf{z})\rangle)
+\end{equation}
 Its architecture comprises multiple fully connected 
@@
     \item Data Input: 
-    Feed a batch of latent data \( z \) from the autoencoder 
+    Feed a batch of latent data \( \mathbf{z} \) from the autoencoder 
     into the real quantum reservoir, recording the resulting quantum 
@@
     \item Surrogate Training:
-    Train the surrogate model using the recorded quantum
-    embeddings as targets, minimizing the loss function:
+    Train the surrogate model  by minimizing the loss function:
     \begin{equation}
-        \mathcal{L}_{\text{surrogate}} = \| \hat{y} - y \|^2,
+        \mathcal{L}_{\text{sur}} = \| q' - q \|^2,
     \end{equation}
-    where \( \hat{y} \) are the surrogate model's predictions
-    and \( y \) are the actual quantum embeddings.
 
@@
     Periodically update the surrogate model during training
-    to insure it accurately reflects the quantum reservoir's
+    to ensure it accurately reflects the quantum reservoir's
     behavior. This can be done after a fixed number of epochs
@@
 
-As a results, we not only achieve end-to-end training of the
+As a result, we not only achieve end-to-end training of the
 hybrid model but also makes it computationally efficient,
@@
 after the systems evolution up to a certain time \( t \).
-
 The readout is performed by measuring specific observables,
@@
 
-\subsubsection{Quantum Reservoir Computing Layer}
+\subsection{Quantum Reservoir Computing Layer}
 Our quantum reservoir computing implementation uses 
@@
 
-\subsubsection{Classical Network Architectures}
-After the traning process of the quantum-guided autoencoder,
+\subsection{QRC Embeddings Linear Classifier}
+After the traning process of the QGARS model,
 the following quantum embeddings are passed to a linear classifier.
@@
 \begin{itemize}
-    \item \textbf{PCA + Linear Classifier:} 
-    A linear classifier trained on the PCA-reduced features.
+    \item \textbf{PCA/Autoencoder + Linear Classifier:} 
+    A linear classifier trained on the PCA-reduced features or the latent representations
+    from the autoencoder.
 
-    \item \textbf{Autoencoder + Linear Classifier:} 
-    A linear classifier trained on the latent representations from the autoencoder.
+    \item \textbf{PCA/Autoencoder + Neural Network:}
+    A fully connected neural network trained on the features from the 
+    same reduction methods as the linear classifier. 
+\end{itemize}
 
-    \item \textbf{PCA + Neural Network:}
-    A fully connected neural network trained on the PCA-reduced features.
-
-    \item \textbf{Autoencoder + Neural Network:}
-    A fully connected neural network trained on the latent representations from the autoencoder.
-\end{itemize}
+We also benchmarked our QGARS encoding strategy
+against the performance of the Linear Mapping of
+QRC embeddings extracted from the latent representation 
+of data using other feature reduction methods, such as
+PCA and the Classical Autoencoder.
 
@@
     \item \textbf{Guided Lambda Parameter (\( \lambda \))}: 
-    This parameter controls the trade-off between reconstruction and classification objectives in the loss function~\ref{eq:composed_loss_function}.
-We tested values ranging from 0.1 to 0.9 in increments of 0.1.
+    To analyse the effect that the trade-off between reconstruction and classification objectives in the loss function~\ref{eq:composed_loss_function} affects
+    the QRC
 
@@
     \item \textbf{Quantum Parameters:} 
-    Including the number of atoms in the chain, Rabi frequency, and detuning parameters. We varied these parameters to assess their impact on classification performance.
-    \item \textbf{Batch Size:}
-    \item \textbf{Learning Rate:}
+    Rabi frequency, and detuning parameters. We varied these parameters to assess their impact on classification performance.
+
+    \item \textbf{Number of Features:}
+    The number of features is the same as the number of qubits (atoms), 
+    as each feature encoded in each atom.
+
 \end{itemize}
+The results from these experiments will be discussed in the next section.
 
@@
 \section{Results and Discussion}
+
 \subsection{Classification Performance Comparison}
+We first compare classification accuracy of the Linear Classifier on the QRC embeddings across PCA,
+vanilla autoencoder, and our QGARS pipeline on each dataset. 
+Table~\ref{tab:qrc_reduction_accuracy} summarizes mean accuracy and standard deviation over 
+five runs, using 12 features.
+
+\begin{table}[!htb]
+  \caption{Classification accuracy (\%) on test sets}
+  \label{tab:qrc_reduction_accuracy}
+  \centering
+  \begin{tabular}{lccc}
+    \hline
+    Method & Synthetic Polyp & CVC-ClinicDB & MNIST(0/1) \\
+    \hline
+    PCA + QRC + Linear      &  00 ± 0.0  &  00 ± 0.0  &  00 ± 0.0 \\
+    AE + QRC + Linear       &  00 ± 0.0  &  00 ± 0.0  &  00 ± 0.0 \\
+    QGARS + QRC + Linear    &  \textbf{00 ± 0.0} &  \textbf{00 ± 0.0} &  \textbf{00 ± 0.0} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\subsection{Dimensionality Reduction Comparison}
+
+To evaluate how well the quantum reservoir embeddings separate classes, we project the high-dimensional embeddings down to $2$ dimensions via t-distributed Stochastic Neighbour Embedding (t-SNE).
+Figure~\ref{fig:embeddings_tsne} shows side-by-side plots for PCA → QRC, AE → QRC, and QGARS → QRC pipelines.
+The QGARS embeddings exhibit the tightest and most distinct clusters, indicating superior discriminative structure.
+
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[width=0.32\textwidth]{pca_qrc_tsne.pdf}
+  \includegraphics[width=0.32\textwidth]{ae_qrc_tsne.pdf}
+  \includegraphics[width=0.32\textwidth]{qgars_qrc_tsne.pdf}
+  \caption{t-SNE projections of the 4 700-D QRC embeddings for (left) PCA + QRC, (center) AE + QRC, (right) QGARS + QRC.}
+  \label{fig:embeddings_tsne}
+\end{figure*}
+
+Quantitatively, we compute the Silhouette Score on the original 4 700-D embeddings (Table~\ref{tab:silhouette}).
+The QGARS pipeline achieves the highest score, confirming better inter-class separation:
+
+\begin{table}[ht]
+  \caption{Silhouette Scores for QRC embeddings}
+  \label{tab:silhouette}
+  \centering
+  \begin{tabular}{lcc}
+    \hline
+    Pipeline         & Silhouette Score \\
+    \hline
+    PCA + QRC        & 0.0 \\
+    AE + QRC         & 0.0 \\
+    QGARS + QRC      & \textbf{0.0} \\
+    \hline
+  \end{tabular}
+\end{table}
+
 \subsection{Ablation Studies}
+
 \subsubsection{Impact of Guided Lambda Parameter}
+We vary the guidance weighting $\lambda$ in Eq.~\ref{eq:composed_loss_function} to assess its influence on performance. Figure~\ref{fig:lambda_sweep} plots test accuracy vs.\ $\lambda$.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1\linewidth]{paper/images/results/generated_polyp_dataset/guided_lambda_log_effect.pdf}
+  \caption{Accuracy as a function of the reconstruction/classification trade-off parameter $\lambda$.}
+  \label{fig:lambda_sweep}
+\end{figure}
+
+\noindent
+We find optimal performance around $\lambda=XXX$; too little guidance ($\lambda<XXX$) yields under-trained classifiers, while too much ($\lambda>XXX$) degrades reconstruction quality.
+
 \subsubsection{Effect of Quantum Update Frequency}
+
 \subsubsection{Influence of Quantum Parameters}
-\subsection{Dimensionality Reduction Comparison}
+
 \subsection{Surrogate Model Fidelity Analysis}
-\subsection{Generalization to Unseen Data}
+We assess how well the surrogate approximates the quantum reservoir by 
+computing the MSE between true quantum embeddings and surrogate outputs over the test set (Figure~\ref{fig:surrogate_fidelity}). Low MSE indicates high fidelity.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.8\linewidth]{surrogate_mse.pdf}
+  \caption{Surrogate vs.\ quantum embedding MSE over training epochs.}
+  \label{fig:surrogate_fidelity}
+\end{figure}
+
+\subsection{Comparison with Classical Methods}
+To benchmark the benefit of our QGARS pipeline, 
+we compare against purely classical feature-reduction 
+approaches followed by standard classifiers. 
+Table~\ref{tab:classical_comp} reports test accuracy on each dataset 
+for the different methods.
+
+\begin{table}[ht]
+  \caption{Test set classification accuracy (\%) for classical baselines}
+  \label{tab:classical_comp}
+  \centering
+  \begin{tabular}{lccc}
+    \hline
+    Method            & Synthetic Polyp & CVC-ClinicDB & MNIST (0/1) \\
+    \hline
+    PCA + Linear      &  0.0 ± 0.0     &  0.0 ± 0.0   &  0.0 ± 0.0    \\
+    AE + Linear       &  0.0 ± 0.0     &  0.0 ± 0.0   &  0.0 ± 0.0    \\
+    AE + NN           &  0.0 ± 0.0     &  0.0 ± 0.0   &  0.0 ± 0.0    \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\noindent
+In contrast, the QGARS pipeline achieves 0.0\%, 0.0\%, and 0.0\% on the same datasets (see Table~\ref{tab:qrc_reduction_accuracy}).
+
+\subsubsection{Training Loss Dynamics}
+
+Figure~\ref{fig:loss_curves} shows the training loss curves for the classical baselines and QGARS on the Synthetic Polyp dataset.
+We plot the reconstruction loss for PCA/AE methods combined with classification loss for the AE+NN and QGARS pipelines.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.9\linewidth]{loss_comparison.pdf}
+  \caption{Normalized training loss vs.\ epoch for PCA+Linear, AE+Linear, AE+NN, and QGARS on the Synthetic Polyp dataset.}
+  \label{fig:loss_curves}
+\end{figure}
+
+\noindent
+Although the loss function of the QGARS method shows
+signs of instability, is not only converges faster but also reaches a 
+lower combined loss
 
@@
 \subsection{Current Limitations}
+While QGARS demonstrates strong performance, several limitations remain:
+\begin{itemize}
+    \item \textbf{Surrogate Fidelity:}
+    Imperfect surrogate approximation can introduce bias in the gradients and limit end‐to‐end 
+    performance, especially if the latent distribution shifts significantly during training.
+    \item \textbf{Classical overhead:}  
+    Frequent surrogate retraining and batch quantum simulations introduce 
+    nontrivial classical compute overhead, potentially offsetting some of the 
+    quantum advantage.
+\end{itemize}
+
 \subsection{Potential Extensions}
+Several avenues can further enhance QGARS:
+\begin{itemize}
+    \item \textbf{Different Hamiltonians:}
+    Explore alternative interaction Hamiltonians (e.g., XY, Heisenberg) to better tailor reservoir dynamics, as experimented by A. D. Lorenzis et al~\cite{lorenzisHarnessingQuantumExtreme2025}.
+
+    \item \textbf{Convolutional Autoencoders:}
+    Replace the fully connected autoencoder with a convolutional AE to learn spatially localized features, making the model input‐size agnostic and enabling richer local structure for the reservoir.
+
+    \item \textbf{Reservoir encoding schemes:}
+    Investigate alternate encodings—such as position encoding via atom placement, global versus local pulse shaping, or multi‐tone driving—to diversify the reservoir’s nonlinear response.
+
+    \item \textbf{\(\lambda\)-Scheduling \& Curriculum Learning:}  
+    Implement a curriculum that gradually increases the guidance weight \(\lambda\) from 0 to 1, allowing the autoencoder to first stabilize reconstruction before introducing quantum guidance.
+Hybrid ramping schedules may accelerate convergence.
+
+    \item \textbf{Noise‐Aware Training:}  
+    Incorporate realistic noise models (decoherence, measurement error) into surrogate training so that the surrogate learns hardware‐like quantum embeddings, improving robustness upon deployment in NISQ devices.
+
+\end{itemize}
+
 \subsection{Hardware Implementation Considerations}
+Bringing QGARS to physical hardware entails additional challenges:
+\begin{itemize}
+    \item \textbf{Calibration and drift:}  
+        Rydberg atom platforms require careful calibration of laser frequencies and intensities; drift over time may degrade reservoir fidelity.
+    \item \textbf{Shot noise and sampling cost:}  
+        Finite‐shot measurements introduce statistical noise; optimizing shot budgets versus surrogate fidelity is critical.
+    \item \textbf{Scalability:}  
+        Extending to hundreds of atoms demands sophisticated trap control and cross‐talk mitigation; resource scheduling on shared QPUs may limit throughput.
+    \item \textbf{Integration with classical pipelines:}  
+        Efficient data transfer between classical and quantum subsystems (e.g.\ via FPGA or GPU–QPU interconnects) and asynchronous surrogate updates will be essential for low‐latency operation.
+\end{itemize}
 
@@
 \section{Conclusion}
+
+We have presented QGARS, a hybrid quantum-classical architecture that integrates a classical autoencoder with a neutral-atom quantum reservoir and a differentiable surrogate model to enable end‐to‐end training.
+By jointly optimizing reconstruction and classification losses, QGARS learns compact and discriminative features tailored for quantum reservoir computing.
+Our experiments on synthetic polyp, CVC-ClinicDB, and binary MNIST datasets demonstrate that QGARS outperforms classical PCA‐ and autoencoder‐based baselines.
+Ablation studies highlight the importance of the guidance weight \(\lambda\), surrogate update frequency, and reservoir parameters.
+While challenges remain—particularly regarding high‐dimensional embeddings, surrogate fidelity, and classical overhead, our proposed extensions and hardware considerations chart a clear path toward robust, scalable quantum‐enhanced medical imaging pipelines in the NISQ era.
 
@@
 
-\newpage
-\newpage
-\tableofcontents
-
 \end{document}

### 2025-05-27 parameter_sweep/data_analysis_scripts/autoencoder_regularization_analysis.py
latent visualization

@@ -89,7 +89,7 @@ def plot_autoencoder_regularization_effect(df: pd.DataFrame, output_path: Option
     # Labels and title
     plt.xlabel('Autoencoder Regularization (Weight Decay)', fontsize=12)
     plt.ylabel('QRC Test Accuracy (%)', fontsize=12)
-    plt.title("Effect of Autoencoder Regularization on QRC Accuracy for the Generated Polyp Dataset", fontsize=14)
+    #plt.title("Effect of Autoencoder Regularization on QRC Accuracy for the Generated Polyp Dataset", fontsize=14)
     
     # Add grid for better readability
     plt.grid(True, which='both', linestyle='--', alpha=0.7)
@@ -102,7 +102,7 @@ def plot_autoencoder_regularization_effect(df: pd.DataFrame, output_path: Option
     # Save figure if output path is provided
     if output_path:
         os.makedirs(os.path.dirname(output_path), exist_ok=True)
-        plt.savefig(output_path, dpi=300, bbox_inches='tight')
+        plt.savefig(output_path, format='pdf', bbox_inches='tight')
         print(f"Figure saved to {output_path}")
     
     plt.show()
@@ -121,7 +121,7 @@ def main():
     print(df[['autoencoder_regularization', 'QRC_final_test_acc']].describe())
     
     # Plot results
-    output_path = "../results/figures/autoencoder_regularization_effect.png"
+    output_path = "../results/figures/autoencoder_regularization_effect.pdf"
     
     # Create plot
     plot_autoencoder_regularization_effect(df, output_path)

### 2025-05-27 parameter_sweep/data_analysis_scripts/guided_lambda_analysis.py

@@ -18,50 +18,6 @@ def load_sweep_results(filepath: str) -> pd.DataFrame:
     
     return df
 
-# def plot_guided_lambda_effect(df: pd.DataFrame, output_path: Optional[str] = None):
-    # """Create a plot showing guided_lambda vs QRC test accuracy."""
-    # # Set up aesthetics
-    # sns.set_style("whitegrid")
-    
-    # # Sort by guided_lambda value
-    # df_sorted = df.sort_values(by='guided_lambda')
-    
-    # # Convert accuracy to percentage
-    # y_values = df_sorted['QRC_final_test_acc'].values * 100
-    # x_values = df_sorted['guided_lambda'].values
-    
-    # # Create a figure with appropriate size
-    # plt.figure(figsize=(10, 6))
-    
-    # # Plot line connecting points
-    # plt.plot(x_values, y_values, '-', linewidth=2, color='#1f77b4')
-    
-    # # Plot individual points
-    # plt.scatter(x_values, y_values, s=80, color='#1f77b4', zorder=5)
-    
-    # # Add value labels
-    # for i, (x, y) in enumerate(zip(x_values, y_values)):
-    #     plt.annotate(f"{y:.1f}%", (x, y), xytext=(0, 10), 
-    #                 textcoords="offset points", ha='center')
-    
-    # # Labels and title
-    # plt.xlabel('Guided Lambda', fontsize=12)
-    # plt.ylabel('QRC Test Accuracy (%)', fontsize=12)
-    # plt.title("Effect of Guided Lambda on QRC Accuracy", fontsize=14)
-    
-    # # Add grid for better readability
-    # plt.grid(True, linestyle='--', alpha=0.7)
-    
-    # plt.tight_layout()
-    
-    # # Save figure if output path is provided
-    # if output_path:
-    #     os.makedirs(os.path.dirname(output_path), exist_ok=True)
-    #     plt.savefig(output_path, dpi=300, bbox_inches='tight')
-    #     print(f"Figure saved to {output_path}")
-    
-    # plt.show()
-
 def plot_guided_lambda_effect(df: pd.DataFrame, output_path: Optional[str] = None):
     """Create a plot showing guided_lambda vs QRC test accuracy."""
     # Set up aesthetics
@@ -94,7 +50,7 @@ def plot_guided_lambda_effect(df: pd.DataFrame, output_path: Optional[str] = Non
     # Labels and title
[…]

### 2025-05-27 parameter_sweep/data_analysis_scripts/n_qubit_reduction_method_analysis.py

@@ -4,7 +4,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import argparse
 
-def plot_qrc_performance_by_method(results_file, output_dir="../results/figures/generated"):
+def plot_qrc_performance_by_method(results_file, output_dir="../../results/figures/generated"):
     """
     Plot QRC performance (train and test accuracy) for different reduction methods
     across varying number of qubits.
@@ -96,8 +96,8 @@ def plot_qrc_performance_by_method(results_file, output_dir="../results/figures/
     plt.tight_layout()
     
     # Save the figure
-    output_file = os.path.join(output_dir, 'qrc_performance_by_reduction_method.png')
-    plt.savefig(output_file, dpi=300, bbox_inches='tight')
+    output_file = os.path.join(output_dir, 'qrc_performance_by_reduction_method.pdf')
+    plt.savefig(output_file, format='pdf', bbox_inches='tight')
     print(f"Figure saved to {output_file}")
     
     # Show the plot
@@ -183,8 +183,8 @@ def main():
     if args.results is None:
         # Try to find the most recent results file in the default location
         default_dir = os.path.join(os.path.dirname(os.path.dirname(os.path.abspath(__file__))), 
-                                  "results", "generated", "n_qubits")
-        
+                                  "..", "results", "generated", "n_qubits")
+
         if os.path.exists(default_dir):
             folders = [f for f in os.listdir(default_dir) if "encoding_dimensions" in f]
             if folders:

### 2025-05-27 parameter_sweep/data_analysis_scripts/quantum_update_frequency_analysis.py

@@ -58,7 +58,7 @@ def plot_quantum_update_frequency_effect(df: pd.DataFrame, output_path: Optional
     # Labels and title
     plt.xlabel('Quantum Update Frequency', fontsize=12)
     plt.ylabel('QRC Accuracy (%)', fontsize=12)
-    plt.title("Effect of Quantum Update Frequency on QRC Performance", fontsize=14)
+    #plt.title("Effect of Quantum Update Frequency on QRC Performance", fontsize=14)
     
     # Add legend
     plt.legend(loc='lower right')
@@ -79,7 +79,7 @@ def plot_quantum_update_frequency_effect(df: pd.DataFrame, output_path: Optional
     # Save figure if output path is provided
     if output_path:
         os.makedirs(os.path.dirname(output_path), exist_ok=True)
-        plt.savefig(output_path, dpi=300, bbox_inches='tight')
+        plt.savefig(output_path, format='pdf', bbox_inches='tight')
         print(f"Figure saved to {output_path}")
     
     plt.show()
@@ -140,8 +140,8 @@ def main():
                       default="/home/nbatista/GIC-quAI-QRC/results/generated/guided_autoencoder_update_frequency/guided_autoencoder_update_frequency_20250520_160624_20250520_160624/all_results.csv",
                       help='Path to the results CSV file')
     parser.add_argument('--output', type=str, 
-                      default="/home/nbatista/GIC-quAI-QRC/results/figures/generated/quantum_update_frequency_effect.png",
-                      help='Path to save the output figure')
+                      default="/home/nbatista/GIC-quAI-QRC/results/figures/generated/quantum_update_frequency_effect.pdf",
+                      help='Path to save the output figure (PDF)')
     
     args = parser.parse_args()
     

### 2025-05-27 parameter_sweep/data_analysis_scripts/time_steps_evolution_time_analysis.py

@@ -95,7 +95,7 @@ def create_heatmap(df: pd.DataFrame, param_grid: dict, output_path: Optional[str
     # Set axis labels and title
     plt.xlabel('Time Steps', fontsize=12)
     plt.ylabel('Evolution Time', fontsize=12)
-    plt.title('Effect of Time Steps and Evolution Time on QRC Test Accuracy', fontsize=14)
+    #plt.title('Effect of Time Steps and Evolution Time on QRC Test Accuracy', fontsize=14)
     
     # Ensure all labels are visible
     plt.tight_layout()
@@ -103,7 +103,7 @@ def create_heatmap(df: pd.DataFrame, param_grid: dict, output_path: Optional[str
     # Save figure if output path is provided
     if output_path:
         os.makedirs(os.path.dirname(output_path), exist_ok=True)
-        plt.savefig(output_path, dpi=300, bbox_inches='tight')
+        plt.savefig(output_path, format='pdf', bbox_inches='tight')
         print(f"Figure saved to {output_path}")
     
     plt.show()
@@ -176,8 +176,8 @@ def main():
     parser.add_argument('--param-grid', type=str, default="/home/nbatista/GIC-quAI-QRC/results/generated/time_steps/evolution_time/time_steps_evolution_time_20250522_012322/param_grid.json",
                       help='Path to the parameter grid JSON file')
     parser.add_argument('--output', type=str, 
-                      default="/home/nbatista/GIC-quAI-QRC/results/figures/generated/time_steps_evolution_time_heatmap.png",
-                      help='Path to save the output figure')
+                      default="/home/nbatista/GIC-quAI-QRC/results/figures/generated/time_steps_evolution_time_heatmap.pdf",
+                      help='Path to save the output figure (PDF)')
     
     args = parser.parse_args()
     

### 2025-05-27 src/data_processing/data_processing.py

@@ -217,7 +217,7 @@ class DatasetLoader:
         
         return train_dataset, test_dataset
 
-    def load_mnist_dataset(self, data_dir: str = './data',
+    def load_mnist_dataset(self, data_dir: str = './data/datasets',
                           mnist_type: str = 'mnist',
                           target_size: Tuple[int, int] = (28, 28),
                           num_examples: Optional[int] = None,

### 2025-05-27 src/embeddings_visualization.py

@@ -1,0 +1,238 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D  # Import for 3D plotting
+from sklearn.manifold import TSNE
+import umap
+from sklearn.metrics import silhouette_score
+import seaborn as sns
+import os
+from typing import Dict, List, Tuple, Any, Optional
+
+from src.data_processing.data_processing import load_dataset, select_random_samples
+from src.feature_reduction.feature_reduction import (
+    apply_pca, apply_autoencoder, apply_guided_autoencoder,
+    scale_to_detuning_range
+)
+from src.quantum_layer.qrc_layer import DetuningLayer
+
+N_FEATURES = 12  # Number of features for the quantum layer
+
+def generate_quantum_embeddings(
+    data_train: Dict[str, Any],
+    quantum_layer: DetuningLayer,
+    method_name: str,
+    encoding_dim: int = N_FEATURES,
+    n_examples: int = 100,
+    guided_lambda: float = 0.3,
+    epochs: int = 30
+) -> Tuple[np.ndarray, np.ndarray]:
+    """Generate quantum embeddings using the specified reduction method"""
+    print(f"\n== Processing {method_name} ==")
+    
+    # Select a subset of samples
+    selected_indices, selected_features, selected_targets = select_random_samples(
+        data_train, num_samples=n_examples, seed=42
+    )
+    
+    # Apply dimensionality reduction
+    if method_name == "pca":
+        xs, _, spectral = apply_pca(
+            data_train, 
+            dim_pca=encoding_dim,
+            selected_indices=selected_indices,
+            selected_features=selected_features
+        )
+    elif method_name == "autoencoder":
+        xs, _, spectral = apply_autoencoder(
+            data_train, 
+            encoding_dim=encoding_dim,
+            epochs=epochs,
+            selected_indices=selected_indices,
+            selected_features=selected_features
+        )
+    elif method_name == "guided_autoencoder":
+        xs, _, spectral, _ = apply_guided_autoencoder(
+            data_train,
+            quantum_layer=quantum_layer,
+            encoding_dim=encoding_dim,
+            epochs=epochs,
+            guided_lambda=guided_lambda,
+            selected_indices=selected_indices,
+            selected_features=selected_features,
+            selected_targets=selected_targets
+        )
+    
+    # Scale to detuning range
[…]

### 2025-05-27 src/feature_reduction/feature_reduction.py

@@ -77,8 +77,7 @@ def apply_autoencoder(data: Dict[str, Any],
                     verbose: bool = True,
                     use_batch_norm: bool = True,
                     dropout: float = 0.1,
-                    weight_decay: float = 1e-5,
-                    autoencoder_regularization: Optional[float] = None,
+                    autoencoder_regularization: Optional[float] = 1e-5,
                     selected_indices: Optional[np.ndarray] = None,
                     selected_features: Optional[np.ndarray] = None) -> Tuple[np.ndarray, Autoencoder, float]:
     """
@@ -147,7 +146,6 @@ def apply_autoencoder(data: Dict[str, Any],
         verbose=verbose, 
         use_batch_norm=use_batch_norm, 
         dropout=dropout, 
-        #weight_decay=weight_decay,
         autoencoder_regularization=autoencoder_regularization
     )
     
@@ -316,8 +314,7 @@ def apply_guided_autoencoder(data: Dict[str, Any],
                             verbose: bool = True,
                             use_batch_norm: bool = True,
                             dropout: float = 0.1,
-                            weight_decay: float = 1e-5,
-                            autoencoder_regularization: Optional[float] = None,
+                            autoencoder_regularization: Optional[float] = 1e-5,
                             selected_indices: Optional[np.ndarray] = None,
                             selected_features: Optional[np.ndarray] = None,
                             selected_targets: Optional[np.ndarray] = None,

### 2025-05-27 parameter_sweep/run_experiments.py
confusion matrix metrics stored

@@ -103,8 +103,6 @@ EXPERIMENT_PROFILES = {
     },
     
     
-
-    
     # RUNNING ON CRAY-4
     "time_steps_evolution_time": {
         "description": "Sweep over quantum evolution time and steps",
@@ -139,7 +137,19 @@ EXPERIMENT_PROFILES = {
             "reduction_method": ["guided_autoencoder"],
             "dim_reduction": [12],
         }
+    },
+    
+    # MOST IMPORTANT FOR THE PAPER
+    "general_test": {
+        "description": "Run a general test with fixed parameters",
+        "seed": [42, 43, 44],
+        "dataset_type": ["binary_mnist", "generated_polyp_dataset", "cvc_clinic_db_patches"],
+        "num_examples": [2000],
+        "num_test_examples": [400],
+        "reduction_method": ["guided_autoencoder", "autoencoder", "pca"],
+        "dim_reduction": [12]
     }
+    
 }
 
 def run_experiment_profile(profile_name: str, 

### 2025-05-27 src/classification_models/training.py

@@ -1,3 +1,4 @@
+from calendar import c
 import numpy as np
 import torch
 import torch.nn as nn
@@ -121,17 +122,50 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
         
         # Evaluate
         model.eval() # Set the model to evaluation mode
+        confusion_matrix_train = {
+            "true_positive": 0,
+            "true_negative": 0,
+            "false_positive": 0,
+            "false_negative": 0
+        }
+        confusion_matrix_test = confusion_matrix_train.copy() 
+
         with torch.no_grad():
             # Training accuracy
             train_outputs = model(x_train_tensor)
             _, train_pred = torch.max(train_outputs, 1)
             train_targets = torch.argmax(y_train_tensor, dim=1)
             train_acc = (train_pred == train_targets).sum().item() / train_targets.size(0)
+            # Update confusion matrix for training
+            for i in range(len(train_pred)):
+                if train_pred[i] == train_targets[i]:
+                    if train_pred[i] == 1:
+                        confusion_matrix_train["true_positive"] += 1
+                    else:
+                        confusion_matrix_train["true_negative"] += 1
+                else:
+                    if train_pred[i] == 1:
+                        confusion_matrix_train["false_positive"] += 1
+                    else:
+                        confusion_matrix_train["false_negative"] += 1            
             
             # Test accuracy
             test_outputs = model(x_test_tensor)
             _, test_pred = torch.max(test_outputs, 1)
             test_acc = (test_pred == y_test_tensor).sum().item() / y_test_tensor.size(0)
+            # Update confusion matrix for testing
+            for i in range(len(test_pred)):
+                if test_pred[i] == y_test_tensor[i]:
+                    if test_pred[i] == 1:
+                        confusion_matrix_test["true_positive"] += 1
+                    else:
+                        confusion_matrix_test["true_negative"] += 1
+                else:
+                    if test_pred[i] == 1:
+                        confusion_matrix_test["false_positive"] += 1
+                    else:
+                        confusion_matrix_test["false_negative"] += 1
+                        
         
         losses.append(epoch_loss / len(train_loader))
         accs_train.append(train_acc)
@@ -141,4 +175,4 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
         if verbose:
[…]

### 2025-05-27 src/main.py

@@ -326,7 +326,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
     # Train different models
     results = {}
     
-    loss_lin, accs_train_lin, accs_test_lin, model_lin = train(
+    loss_lin, accs_train_lin, accs_test_lin, model_lin, confusion_matrix_train, confusion_matrix_test = train(
         x_train=xs, 
         y_train=ys, 
         x_test=test_features, 
@@ -338,7 +338,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         verbose=not args.no_progress,
         nonlinear=False
     )
-    results["linear"] = (loss_lin, accs_train_lin, accs_test_lin, model_lin)
+    results["linear"] = (loss_lin, accs_train_lin, accs_test_lin, model_lin, confusion_matrix_train, confusion_matrix_test)
     
     print(model_lin)
     
@@ -351,7 +351,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         
         """)
     
-    loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc = train(
+    loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc, confusion_matrix_train, confusion_matrix_test = train(
         embeddings, ys, test_embeddings, test_targets, 
         regularization=args.classifier_regularization,  # Use classifier regularization
         nepochs=args.nepochs, 
@@ -360,7 +360,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         verbose=not args.no_progress,
         nonlinear=False
     )
-    results["QRC"] = (loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc)
+    results["QRC"] = (loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc, confusion_matrix_train, confusion_matrix_test)
     
     print(model_qrc)
     
@@ -372,7 +372,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         =========================================
         
         """)
-    loss_nn, accs_train_nn, accs_test_nn, model_nn = train(
+    loss_nn, accs_train_nn, accs_test_nn, model_nn, confusion_matrix_train, confusion_matrix_test = train(
         xs, ys, test_features, test_targets, 
[…]

### 2025-05-27 src/utils/statistics_tracking.py

@@ -60,7 +60,7 @@ def save_classifier_loss_plot(results_dict: Dict[str, Tuple[List[float], List[fl
     """
     plt.figure(figsize=(10, 6))
     
-    for name, (losses, _, _, _) in results_dict.items():
+    for name, (losses, _, _, _, _, _) in results_dict.items():
         plt.plot(losses, label=f"{name} loss")
     
     plt.xlabel("Epoch")
@@ -91,7 +91,7 @@ def save_classifier_accuracy_plot(results_dict: Dict[str, Tuple[List[float], Lis
     """
     plt.figure(figsize=(10, 6))
     
-    for name, (_, accs_train, accs_test, _) in results_dict.items():
+    for name, (_, accs_train, accs_test, _, _, _) in results_dict.items():
         plt.plot(accs_train, label=f"{name} train")
         plt.plot(accs_test, label=f"{name} test")
     
@@ -168,6 +168,22 @@ def save_guided_autoencoder_losses(losses: Dict[str, List[float]], output_dir: s
         
         logging.info(f"Saved surrogate model loss plot to {surrogate_plot_path}")
 
+def get_f1_score(confusion_matrix: Dict[str, int]) -> float:
+    tp = confusion_matrix['true_positive']
+    tn = confusion_matrix['true_negative']
+    fp = confusion_matrix['false_positive']
+    fn = confusion_matrix['false_negative']
+    
+    precision = tp / (tp + fp) if (tp + fp) > 0 else 0.0
+    recall = tp / (tp + fn) if (tp + fn) > 0 else 0.0
+    
+    if precision + recall == 0:
+        return 0.0
+    
+    f1_score = 2 * (precision * recall) / (precision + recall)
+    
+    return f1_score
+
 def save_loss_logs(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]], 
                   output_dir: str) -> None:
     """
@@ -182,7 +198,7 @@ def save_loss_logs(results_dict: Dict[str, Tuple[List[float], List[float], List[
     """
     log_data = {}
     
-    for name, (losses, accs_train, accs_test, _) in results_dict.items():
+    for name, (losses, accs_train, accs_test, _, _, _) in results_dict.items():
         log_data[name] = {
             'losses': losses,
             'train_accuracy': accs_train,
@@ -234,10 +250,14 @@ def extract_metrics(results_dict: Dict[str, Tuple[List[float], List[float], List
     }
     
     # Extract final metrics from each model
[…]

### 2025-05-27 src/utils/visualization.py

@@ -14,7 +14,7 @@ def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float]
     
     # Plot accuracies
     plt.subplot(1, 2, 1)
-    for name, (_, accs_train, accs_test, _) in results_dict.items():
+    for name, (_, accs_train, accs_test, _, _, _) in results_dict.items():
         plt.plot(accs_train, label=f"{name} train")
         plt.plot(accs_test, label=f"{name} test")
     
@@ -25,7 +25,7 @@ def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float]
     
     # Plot losses
     plt.subplot(1, 2, 2)
-    for name, (losses, _, _, _) in results_dict.items():
+    for name, (losses, _, _, _, _, _) in results_dict.items():
         plt.plot(losses, label=name)
     
     plt.xlabel("Epoch")
@@ -47,6 +47,10 @@ def print_results(results_dict: Dict[str, Tuple[List[float], List[float], List[f
     """
     print("\nResults Summary:")
     print("-" * 50)
-    for name, (_, _, accs_test, _) in results_dict.items():
+    for name, (_, _, accs_test, _, _, confusion_matrix_test) in results_dict.items():
         print(f"{name} test accuracy = {accs_test[-1]*100:.2f}%")
+
+        print(f"Confusion Matrix for {name} (Test Set):")
+        print(confusion_matrix_test)
+        
     print("-" * 50)

### 2025-05-29 parameter_sweep/data_analysis_scripts/guided_lambda_analysis.py
better lambda analysis and visualization

@@ -2,6 +2,8 @@ import os
 import pandas as pd
 import matplotlib.pyplot as plt
 import seaborn as sns
+import numpy as np
+import argparse
 from typing import Optional
 
 def load_sweep_results(filepath: str) -> pd.DataFrame:
@@ -18,44 +20,118 @@ def load_sweep_results(filepath: str) -> pd.DataFrame:
     
     return df
 
+def transform_lambda_to_mu(lambda_val):
+    """Transform lambda to mu using mu = sqrt(lambda/(1-lambda))."""
+    # Handle edge cases
+    # if lambda_val >= 0.999999:  # Treat values very close to 1 as a large finite number
+        # return 1000.0  # Use: a large value instead of infinity
+    if lambda_val == 0.999:
+        return 15
+    if lambda_val == 0.9999:
+        return 30
+    if lambda_val == 0.99999:
+        return 50
+    if lambda_val == 0.999999:
+        return 75
+    elif lambda_val >= 1.0:
+        return 100
+    
+    elif lambda_val <= 0.0:
+        return 0.0
+    else:
+        return np.sqrt(lambda_val / (1 - lambda_val))
+
 def plot_guided_lambda_effect(df: pd.DataFrame, output_path: Optional[str] = None):
-    """Create a plot showing guided_lambda vs QRC test accuracy."""
+    """Create a plot showing guided_lambda vs QRC test accuracy with three separate views."""
     # Set up aesthetics
     sns.set_style("whitegrid")
     
     # Sort by guided_lambda value
     df_sorted = df.sort_values(by='guided_lambda')
     
-    # Convert accuracy to percentage
-    y_values = df_sorted['QRC_final_test_acc'].values * 100
-    x_values = df_sorted['guided_lambda'].values
+    # Create a figure with 3 subplots horizontally arranged
+    fig, axes = plt.subplots(1, 3, figsize=(18, 6))
     
-    # Create a figure with appropriate size
-    plt.figure(figsize=(10, 6))
+    # Filter data for each range
+    df_small = df_sorted[(df_sorted['guided_lambda'] > 0) & (df_sorted['guided_lambda'] <= 0.1)]
+    df_mid = df_sorted[(df_sorted['guided_lambda'] >= 0.1) & (df_sorted['guided_lambda'] <= 0.9)]
[…]

### 2025-05-29 src/embeddings_visualization.py

@@ -6,6 +6,7 @@ import umap
 from sklearn.metrics import silhouette_score
 import seaborn as sns
 import os
+import json
 from typing import Dict, List, Tuple, Any, Optional
 
 from src.data_processing.data_processing import load_dataset, select_random_samples
@@ -177,6 +178,31 @@ def calculate_metrics(
     plt.savefig(f"{output_dir}/quantum_embeddings_metrics.png", dpi=300)
     plt.show()
 
+def save_embeddings_to_json(
+    results: Dict[str, Dict[str, np.ndarray]],
+    output_dir: str
+):
+    """Save embeddings and their labels to a JSON file"""
+    os.makedirs(output_dir, exist_ok=True)
+    
+    # Prepare data for JSON serialization (convert numpy arrays to lists)
+    json_data = {}
+    for method_name, data in results.items():
+        embeddings = data["embeddings"].tolist()  # Convert to Python list
+        labels = data["labels"].tolist()  # Convert to Python list
+        
+        json_data[method_name] = {
+            "embeddings": embeddings,
+            "labels": labels
+        }
+    
+    # Save to JSON file
+    json_path = os.path.join(output_dir, "quantum_embeddings.json")
+    with open(json_path, 'w') as f:
+        json.dump(json_data, f)
+    
+    print(f"Saved embeddings to {json_path}")
+
 def main():
     global N_FEATURES 
 
@@ -217,6 +243,12 @@ def main():
             "embeddings": embeddings,
             "labels": labels
         }
+    
+    # Save embeddings to JSON
+    save_embeddings_to_json(
+        results,
+        output_dir="/home/nbatista/GIC-quAI-QRC/results/embeddings"
+    )
     
     # Visualize embeddings
     visualize_embeddings(

### 2025-05-29 parameter_sweep/parameter_sweep.py
quantum surrogate loss function tracking

@@ -15,7 +15,7 @@ sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), '..', 's
 
 from main import main
 from utils.config_manager import ConfigManager
-from utils.statistics_tracking import save_all_statistics  # Add this import
+from utils.statistics_tracking import save_all_statistics, extract_metrics  # Import the functions directly
 
 class ParameterSweep:
     """
@@ -177,30 +177,34 @@ class ParameterSweep:
             exp_dir = os.path.join(sweep_dir, exp_id)
             os.makedirs(exp_dir, exist_ok=True)
             
-            # Run main function with args - now unpacking both return values
+            # Run main function with args
             results, guided_losses = main(args, results_dir=exp_dir)
             
-            # Save statistics in the experiment directory - no need to check __main__
-            #save_all_statistics(results, guided_losses, exp_dir)
+            # Save all statistics using the same function as main.py
+            save_all_statistics(results, guided_losses, exp_dir, args)
             
-            # Extract metrics
-            metrics = self._extract_metrics(results)
+            # Extract metrics using the common function
+            metrics = extract_metrics(results)
             
-            # Add configuration parameters to metrics
+            # Add configuration parameters to metrics for the DataFrame
             for key, value in config.items():
                 # Skip large configurations
                 if key not in ['autoencoder_hidden_dims', 'target_size']:
                     metrics[key] = value
-                    
-            # Save configuration
-            with open(os.path.join(exp_dir, "config.json"), "w") as f:
-                config_serializable = {k: str(v) if isinstance(v, np.ndarray) else v 
-                                     for k, v in config.items()}
-                json.dump(config_serializable, f, indent=4)
+            
+            # Add guided_autoencoder metrics if available
+            if guided_losses is not None:
+                metrics["guided_autoencoder_final_total_loss"] = float(guided_losses["total_loss"][-1])
+                metrics["guided_autoencoder_final_recon_loss"] = float(guided_losses["recon_loss"][-1])
+                metrics["guided_autoencoder_final_class_loss"] = float(guided_losses["class_loss"][-1])
                 
-            # Save metrics
-            with open(os.path.join(exp_dir, "metrics.json"), "w") as f:
-                json.dump(metrics, f, indent=4)
+                # Add surrogate metrics if available
[…]

### 2025-05-29 parameter_sweep/run_experiments.py

@@ -142,12 +142,28 @@ EXPERIMENT_PROFILES = {
     # MOST IMPORTANT FOR THE PAPER
     "general_test": {
         "description": "Run a general test with fixed parameters",
-        "seed": [42, 43, 44],
-        "dataset_type": ["binary_mnist", "generated_polyp_dataset", "cvc_clinic_db_patches"],
-        "num_examples": [2000],
-        "num_test_examples": [400],
-        "reduction_method": ["guided_autoencoder", "autoencoder", "pca"],
-        "dim_reduction": [12]
+        "param_grid": {
+            "seed": [42, 43, 44, 45, 46],
+            "dataset_type": ["binary_mnist", "generated_polyp_dataset", "cvc_clinic_db_patches"],
+            "num_examples": [2000],
+            "num_test_examples": [400],
+            "reduction_method": ["guided_autoencoder", "autoencoder", "pca"],
+            "dim_reduction": [12]
+        }
+    },
+    
+    # MOST IMPORTANT FOR THE PAPER
+    "test": {
+        "description": "asdf",
+        "param_grid": {
+            "seed": [42],
+            "dataset_type": ["generated_polyp_dataset"],
+            "num_examples": [5],
+            "num_test_examples": [2],
+            "quantum_update_frequency": [25],
+            "reduction_method": ["guided_autoencoder"],
+            "dim_reduction": [4]
+        }
     }
     
 }

### 2025-05-29 src/classification_models/training.py

@@ -6,6 +6,7 @@ import torch.optim as optim
 from torch.utils.data import DataLoader, TensorDataset
 from tqdm import tqdm
 from typing import Tuple, List, Dict, Any, Optional, Union
+from sklearn.metrics import confusion_matrix
 
 from src.classification_models.models import LinearClassifier, NeuralNetwork
 
@@ -137,35 +138,23 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
             train_targets = torch.argmax(y_train_tensor, dim=1)
             train_acc = (train_pred == train_targets).sum().item() / train_targets.size(0)
             # Update confusion matrix for training
-            for i in range(len(train_pred)):
-                if train_pred[i] == train_targets[i]:
-                    if train_pred[i] == 1:
-                        confusion_matrix_train["true_positive"] += 1
-                    else:
-                        confusion_matrix_train["true_negative"] += 1
-                else:
-                    if train_pred[i] == 1:
-                        confusion_matrix_train["false_positive"] += 1
-                    else:
-                        confusion_matrix_train["false_negative"] += 1            
+            confusion_matrix_train = confusion_matrix(
+                train_targets.numpy(), 
+                train_pred.numpy(), 
+                labels=range(output_dim)
+            )          
             
             # Test accuracy
             test_outputs = model(x_test_tensor)
             _, test_pred = torch.max(test_outputs, 1)
             test_acc = (test_pred == y_test_tensor).sum().item() / y_test_tensor.size(0)
             # Update confusion matrix for testing
-            for i in range(len(test_pred)):
-                if test_pred[i] == y_test_tensor[i]:
-                    if test_pred[i] == 1:
-                        confusion_matrix_test["true_positive"] += 1
-                    else:
-                        confusion_matrix_test["true_negative"] += 1
-                else:
-                    if test_pred[i] == 1:
-                        confusion_matrix_test["false_positive"] += 1
-                    else:
-                        confusion_matrix_test["false_negative"] += 1
-                        
+            confusion_matrix_test = confusion_matrix(
+                y_test_tensor.numpy(), 
+                test_pred.numpy(), 
+                labels=range(output_dim)
+            )
+            
         
         losses.append(epoch_loss / len(train_loader))
         accs_train.append(train_acc)
[…]

### 2025-05-29 src/feature_reduction/autoencoder/guided_autoencoder.py

@@ -224,7 +224,7 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
                              device: str,
                              batch_size_qe: int = 50,
                              detuning_max: float = 6.0,
-                             verbose: bool = True) -> QuantumSurrogate:
+                             verbose: bool = True) -> Tuple[QuantumSurrogate, float]:
     """
     Update quantum embeddings and train surrogate model.
     
@@ -249,8 +249,8 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
         
     Returns
     -------
-    QuantumSurrogate
-        Trained surrogate model
+    Tuple[QuantumSurrogate, float]
+        Trained surrogate model and final surrogate loss
     """
     model.clear_cache()
     n_samples = X.shape[0]
@@ -311,9 +311,11 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
         if verbose:
             tqdm.write("Training surrogate model with current quantum embeddings...")
         
+        surrogate_loss = None
+        
         if model.surrogate is None:
             # Create and train new surrogate
-            surrogate = create_and_train_surrogate(
+            surrogate, loss_history = create_and_train_surrogate(
                 quantum_layer=quantum_layer,
                 encodings=all_encodings_tensor,
                 quantum_embeddings=all_quantum_embs_tensor,
@@ -321,9 +323,10 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
                 verbose=verbose,
                 n_shots=n_shots
             )
+            surrogate_loss = loss_history[-1] if loss_history else None
         else:
             # Update existing surrogate
-            surrogate = train_surrogate(
+            surrogate, loss_history = train_surrogate(
                 surrogate_model=model.surrogate,
                 quantum_layer=quantum_layer,
                 input_data=all_encodings_tensor,
@@ -335,12 +338,13 @@ def update_quantum_embeddings(model: GuidedAutoencoder,
                 verbose=verbose,
                 n_shots=n_shots
             )
+            surrogate_loss = loss_history[-1] if loss_history else None
         
         # Set surrogate in model
         model.initialize_surrogate(surrogate)
         
     model.autoencoder.train()
-    return surrogate
+    return surrogate, surrogate_loss
 
 def train_guided_epoch(model: GuidedAutoencoder,
                       dataloader: torch.utils.data.DataLoader,
[…]

### 2025-05-29 src/feature_reduction/autoencoder/quantum_surrogate.py

@@ -174,7 +174,7 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
                    learning_rate: float = 0.001,
                    device: str = 'cpu',
                    verbose: bool = True,
-                   n_shots: int = 1000) -> QuantumSurrogate:
+                   n_shots: int = 1000) -> Tuple[QuantumSurrogate, List[float]]:
     """
     Train the surrogate model to mimic the quantum layer.
     
@@ -203,8 +203,8 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
         
     Returns
     -------
-    QuantumSurrogate
-        Trained surrogate model
+    Tuple[QuantumSurrogate, List[float]]
+        Trained surrogate model and loss history
     """
     # Ensure model is on the correct device and in training mode
     surrogate_model = surrogate_model.to(device)
@@ -213,6 +213,9 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
     # Set up optimizer and loss function
     optimizer = optim.Adam(surrogate_model.parameters(), lr=learning_rate)
     criterion = nn.MSELoss()
+    
+    # Track loss history
+    loss_history = []
     
     # Convert input data to numpy for quantum layer if needed
     input_numpy = None
@@ -239,7 +242,6 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
         quantum_embeddings = torch.tensor(quantum_embs.T, dtype=torch.float32)
     
     # Make sure quantum embeddings are on the correct device
-    # quantum_embeddings = torch.tensor(quantum_embeddings, dtype=torch.float32).to(device)
     if isinstance(quantum_embeddings, torch.Tensor):
         quantum_embeddings = quantum_embeddings.clone().detach().to(device).requires_grad_(True)
     else:
@@ -249,9 +251,10 @@ def train_surrogate(surrogate_model: QuantumSurrogate,
     if isinstance(input_data, np.ndarray):
         input_tensor = torch.tensor(input_data.T, dtype=torch.float32).to(device)
     else:
-        input_tensor = input_data.detach().cpu()#.to(device)  # Detach to ensure we don't track unnecessary gradients
+        input_tensor = input_data.detach().cpu()  # Detach to ensure we don't track unnecessary gradients
         
     quantum_embeddings_cpu = quantum_embeddings.cpu() if hasattr(quantum_embeddings, 'is_cuda') else quantum_embeddings
+    
     # Determine batch size based on dataset size
[…]

### 2025-05-29 src/main.py

@@ -331,7 +331,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         y_train=ys, 
         x_test=test_features, 
         y_test=test_targets, 
-        regularization=args.classifier_regularization,  # Use classifier regularization 
+        regularization=args.classifier_regularization,  
         nepochs=args.nepochs, 
         batchsize=args.batchsize, 
         learning_rate=args.learning_rate,
@@ -353,7 +353,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
     
     loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc, confusion_matrix_train, confusion_matrix_test = train(
         embeddings, ys, test_embeddings, test_targets, 
-        regularization=args.classifier_regularization,  # Use classifier regularization
+        regularization=args.classifier_regularization, 
         nepochs=args.nepochs, 
         batchsize=args.batchsize, 
         learning_rate=args.learning_rate,
@@ -374,7 +374,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         """)
     loss_nn, accs_train_nn, accs_test_nn, model_nn, confusion_matrix_train, confusion_matrix_test = train(
         xs, ys, test_features, test_targets, 
-        regularization=args.classifier_regularization,  # Use classifier regularization 
+        regularization=args.classifier_regularization,
         nepochs=args.nepochs, 
         batchsize=args.batchsize, 
         learning_rate=args.learning_rate,

### 2025-05-29 src/utils/cli_utils.py

@@ -8,14 +8,14 @@ from src.utils.config_manager import ConfigManager
 AVAILABLE_GEOMETRIES = ["chain"] 
 AVAILABLE_READOUT_TYPES = ["Z", "ZZ", "all"]
 AVAILABLE_REDUCTION_METHODS = ["pca", "autoencoder", "guided_autoencoder"]
-AVAILABLE_DATASET_TYPES = ["mnist", "binary_mnist", "fashion_mnist", "image_folder"]
+#AVAILABLE_DATASET_TYPES = ["mnist", "binary_mnist", "fashion_mnist", "image_folder"]
 
 def parse_args() -> argparse.Namespace:
     """Parse command line arguments"""
     parser = argparse.ArgumentParser(description="Quantum Reservoir Computing for Image Classification")
     
     # Dataset parameters
-    parser.add_argument("--dataset-type", type=str, default=None, choices=AVAILABLE_DATASET_TYPES,
+    parser.add_argument("--dataset-type", type=str, default=None,
                        help="Type of dataset to use")
     parser.add_argument("--data-dir", type=str, default=None, 
                        help="Path to dataset directory containing class subfolders")

### 2025-05-29 src/utils/statistics_tracking.py

@@ -116,7 +116,7 @@ def save_guided_autoencoder_losses(losses: Dict[str, List[float]], output_dir: s
     Parameters
     ----------
     losses : Dict[str, List[float]]
-        Dictionary with keys 'total_loss', 'recon_loss', 'class_loss' mapping to loss histories
+        Dictionary with keys 'total_loss', 'recon_loss', 'class_loss', 'surrogate_loss' mapping to loss histories
     output_dir : str
         Directory to save the plot
     """
@@ -166,7 +166,16 @@ def save_guided_autoencoder_losses(losses: Dict[str, List[float]], output_dir: s
         plt.savefig(surrogate_plot_path, dpi=300, bbox_inches='tight')
         plt.close()
         
+        # Also save surrogate loss to a separate JSON file for easier access
+        surrogate_json_path = os.path.join(output_dir, "surrogate_loss.json")
+        with open(surrogate_json_path, 'w') as f:
+            json.dump({
+                "epochs": surrogate_epochs,
+                "loss_values": surrogate_values
+            }, f, indent=4)
+        
         logging.info(f"Saved surrogate model loss plot to {surrogate_plot_path}")
+        logging.info(f"Saved surrogate model loss data to {surrogate_json_path}")
 
 def get_f1_score(confusion_matrix: Dict[str, int]) -> float:
     tp = confusion_matrix['true_positive']
@@ -230,6 +239,7 @@ def save_guided_autoencoder_logs(losses: Dict[str, List[float]], output_dir: str
     
     logging.info(f"Saved guided autoencoder metrics to {log_path}")
 
+# Rename the internal function to be exported
 def extract_metrics(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]]) -> Dict[str, Any]:
     """
     Extract key metrics from results dictionary.
@@ -365,7 +375,7 @@ def save_all_statistics(results_dict: Dict[str, Tuple[List[float], List[float], 
         save_guided_autoencoder_losses(guided_losses, output_dir)
         save_guided_autoencoder_logs(guided_losses, output_dir)
     
-    # Extract and save metrics in JSON format (similar to parameter_sweep)
+    # Extract and save metrics in JSON format (using the common function)
     metrics = extract_metrics(results_dict)
     
     # Add guided autoencoder metrics if available

### 2025-05-29 src/utils/visualization.py

@@ -47,10 +47,11 @@ def print_results(results_dict: Dict[str, Tuple[List[float], List[float], List[f
     """
     print("\nResults Summary:")
     print("-" * 50)
-    for name, (_, _, accs_test, _, _, confusion_matrix_test) in results_dict.items():
+    for name, (_, _, accs_test, _, confusion_matrix_train, confusion_matrix_test) in results_dict.items():
         print(f"{name} test accuracy = {accs_test[-1]*100:.2f}%")
 
         print(f"Confusion Matrix for {name} (Test Set):")
         print(confusion_matrix_test)
+
         
     print("-" * 50)

### 2025-05-29 src/classification_models/training.py
f1 score

@@ -6,7 +6,7 @@ import torch.optim as optim
 from torch.utils.data import DataLoader, TensorDataset
 from tqdm import tqdm
 from typing import Tuple, List, Dict, Any, Optional, Union
-from sklearn.metrics import confusion_matrix
+from sklearn.metrics import confusion_matrix, f1_score
 
 from src.classification_models.models import LinearClassifier, NeuralNetwork
 
@@ -123,14 +123,6 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
         
         # Evaluate
         model.eval() # Set the model to evaluation mode
-        confusion_matrix_train = {
-            "true_positive": 0,
-            "true_negative": 0,
-            "false_positive": 0,
-            "false_negative": 0
-        }
-        confusion_matrix_test = confusion_matrix_train.copy() 
-
         with torch.no_grad():
             # Training accuracy
             train_outputs = model(x_train_tensor)
@@ -154,7 +146,6 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
                 test_pred.numpy(), 
                 labels=range(output_dim)
             )
-            
         
         losses.append(epoch_loss / len(train_loader))
         accs_train.append(train_acc)
@@ -164,4 +155,24 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
         if verbose:
             tqdm.write(f"Epoch {epoch+1}/{nepochs} - Loss: {losses[-1]:.4f} - Train Acc: {train_acc:.4f} - Test Acc: {test_acc:.4f}")
     
-    return losses, accs_train, accs_test, model, confusion_matrix_train, confusion_matrix_test
+    # Calculate final F1 scores
+    if output_dim == 2:
+        average_type = 'binary'
+    else:
+        average_type = 'weighted'
+
+    f1_train = f1_score(
+        train_targets.numpy(),
+        train_pred.numpy(),
+        labels=range(output_dim),
+        average=average_type
+    )
+    f1_test = f1_score(
+        y_test_tensor.numpy(),
+        test_pred.numpy(),
+        labels=range(output_dim),
+        average=average_type
+    )
+    
+    
+    return losses, accs_train, accs_test, model, confusion_matrix_train, confusion_matrix_test, f1_train, f1_test

### 2025-05-29 src/main.py

@@ -326,7 +326,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
     # Train different models
     results = {}
     
-    loss_lin, accs_train_lin, accs_test_lin, model_lin, confusion_matrix_train, confusion_matrix_test = train(
+    loss_lin, accs_train_lin, accs_test_lin, model_lin, confusion_matrix_train, confusion_matrix_test, f1_train, f1_test = train(
         x_train=xs, 
         y_train=ys, 
         x_test=test_features, 
@@ -338,7 +338,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         verbose=not args.no_progress,
         nonlinear=False
     )
-    results["linear"] = (loss_lin, accs_train_lin, accs_test_lin, model_lin, confusion_matrix_train, confusion_matrix_test)
+    results["linear"] = (loss_lin, accs_train_lin, accs_test_lin, model_lin, confusion_matrix_train, confusion_matrix_test, f1_train, f1_test)
     
     print(model_lin)
     
@@ -351,7 +351,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         
         """)
     
-    loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc, confusion_matrix_train, confusion_matrix_test = train(
+    loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc, confusion_matrix_train, confusion_matrix_test, f1_train, f1_test = train(
         embeddings, ys, test_embeddings, test_targets, 
         regularization=args.classifier_regularization, 
         nepochs=args.nepochs, 
@@ -360,7 +360,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         verbose=not args.no_progress,
         nonlinear=False
     )
-    results["QRC"] = (loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc, confusion_matrix_train, confusion_matrix_test)
+    results["QRC"] = (loss_qrc, accs_train_qrc, accs_test_qrc, model_qrc, confusion_matrix_train, confusion_matrix_test, f1_train, f1_test)
     
     print(model_qrc)
     
@@ -372,7 +372,7 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
         =========================================
         
         """)
[…]

### 2025-05-29 src/utils/statistics_tracking.py

@@ -60,7 +60,7 @@ def save_classifier_loss_plot(results_dict: Dict[str, Tuple[List[float], List[fl
     """
     plt.figure(figsize=(10, 6))
     
-    for name, (losses, _, _, _, _, _) in results_dict.items():
+    for name, (losses, _, _, _, _, _, _, _) in results_dict.items():
         plt.plot(losses, label=f"{name} loss")
     
     plt.xlabel("Epoch")
@@ -91,7 +91,7 @@ def save_classifier_accuracy_plot(results_dict: Dict[str, Tuple[List[float], Lis
     """
     plt.figure(figsize=(10, 6))
     
-    for name, (_, accs_train, accs_test, _, _, _) in results_dict.items():
+    for name, (_, accs_train, accs_test, _, _, _, _, _) in results_dict.items():
         plt.plot(accs_train, label=f"{name} train")
         plt.plot(accs_test, label=f"{name} test")
     
@@ -177,22 +177,6 @@ def save_guided_autoencoder_losses(losses: Dict[str, List[float]], output_dir: s
         logging.info(f"Saved surrogate model loss plot to {surrogate_plot_path}")
         logging.info(f"Saved surrogate model loss data to {surrogate_json_path}")
 
-def get_f1_score(confusion_matrix: Dict[str, int]) -> float:
-    tp = confusion_matrix['true_positive']
-    tn = confusion_matrix['true_negative']
-    fp = confusion_matrix['false_positive']
-    fn = confusion_matrix['false_negative']
-    
-    precision = tp / (tp + fp) if (tp + fp) > 0 else 0.0
-    recall = tp / (tp + fn) if (tp + fn) > 0 else 0.0
-    
-    if precision + recall == 0:
-        return 0.0
-    
-    f1_score = 2 * (precision * recall) / (precision + recall)
-    
-    return f1_score
-
 def save_loss_logs(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]], 
                   output_dir: str) -> None:
     """
@@ -207,7 +191,7 @@ def save_loss_logs(results_dict: Dict[str, Tuple[List[float], List[float], List[
     """
     log_data = {}
     
-    for name, (losses, accs_train, accs_test, _, _, _) in results_dict.items():
+    for name, (losses, accs_train, accs_test, _, _, _, _, _) in results_dict.items():
         log_data[name] = {
             'losses': losses,
             'train_accuracy': accs_train,
[…]

### 2025-05-29 src/utils/visualization.py

@@ -14,7 +14,7 @@ def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float]
     
     # Plot accuracies
     plt.subplot(1, 2, 1)
-    for name, (_, accs_train, accs_test, _, _, _) in results_dict.items():
+    for name, (_, accs_train, accs_test, _, _, _, _, _) in results_dict.items():
         plt.plot(accs_train, label=f"{name} train")
         plt.plot(accs_test, label=f"{name} test")
     
@@ -25,7 +25,7 @@ def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float]
     
     # Plot losses
     plt.subplot(1, 2, 2)
-    for name, (losses, _, _, _, _, _) in results_dict.items():
+    for name, (losses, _, _, _, _, _, _, _) in results_dict.items():
         plt.plot(losses, label=name)
     
     plt.xlabel("Epoch")
@@ -47,8 +47,9 @@ def print_results(results_dict: Dict[str, Tuple[List[float], List[float], List[f
     """
     print("\nResults Summary:")
     print("-" * 50)
-    for name, (_, _, accs_test, _, confusion_matrix_train, confusion_matrix_test) in results_dict.items():
+    for name, (_, _, accs_test, _, _, confusion_matrix_test, _, f1_test) in results_dict.items():
         print(f"{name} test accuracy = {accs_test[-1]*100:.2f}%")
+        print(f"{name} test F1 score = {f1_test:.4f}")
 
         print(f"Confusion Matrix for {name} (Test Set):")
         print(confusion_matrix_test)

### 2025-05-29 parameter_sweep/run_experiments.py
new profile

@@ -140,7 +140,7 @@ EXPERIMENT_PROFILES = {
     },
     
     # MOST IMPORTANT FOR THE PAPER
-    "general_test": {
+    "ae_pca_qgars_5_runs": {
         "description": "Run a general test with fixed parameters",
         "param_grid": {
             "seed": [42, 43, 44, 45, 46],
@@ -152,7 +152,6 @@ EXPERIMENT_PROFILES = {
         }
     },
     
-    # MOST IMPORTANT FOR THE PAPER
     "test": {
         "description": "asdf",
         "param_grid": {

### 2025-05-29 paper/main.tex
Updates from Overleaf

@@
 \usepackage{inputenc}
+\usepackage{bm}
+%\usepackage{amsmath}
 
@@
 
-% \and
-% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-% \textit{name of organization (of Aff.)}\\
-% City, Country \\
-% email address or ORCID}
+% % \and
+% % \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+% % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+% % \textit{name of organization (of Aff.)}\\
+% % City, Country \\
+% % email address or ORCID}
 % }
@@
 architecture for medical image classification based on neutral-atom quantum processors.
-This approach is designed to address the challenges of
-medical imaging, with a particular focus on tasks such
-as polyp detection and classification.
-By integrating an autoencoder guided by a quantum 
-reservoir, the pipeline learns compact and discriminative 
-representations of image data that are also well-suited for quantum reservoir 
-computing. To overcome the non-differentiability of quantum 
-measurements, we circumvent this `gradient barrier' 
-by incorporating a differentiable surrogate model that simulates the 
-behaviour of the quantum layer, enabling end-to-end backpropagation. 
-The guided training process jointly optimizes for both image 
-reconstruction and classification accuracy, ensuring that the latent 
-representations are both meaningful and effective for quantum processing. 
-In our implementation, image data is encoded as atom detuning parameters 
-in a Rydberg Hamiltonian, and quantum embeddings are obtained through expectation 
-values. These embeddings are then passed to a linear classifier, 
-enabling faster training and inference compared to deep classical networks.
-Our experiments show that this method outperforms traditional 
-approaches using PCA or unguided autoencoders. We also conduct 
-ablation studies to evaluate the impact of quantum and training 
-parameters, demonstrating the robustness and flexibility of the 
-proposed pipeline for real-world medical imaging applications, 
-even in the NISQ era.
+This approach is designed to address the challenges of medical imaging, with a particular focus on tasks such as polyp detection and classification.
+By integrating an autoencoder guided by a quantum reservoir, the pipeline learns compact and discriminative representations of image data that are also well-suited for quantum reservoir computing.
+To overcome the non-differentiability of quantum measurements, we circumvent this `gradient barrier' by incorporating a differentiable surrogate model that simulates the behaviour of the quantum layer, enabling end-to-end backpropagation.
+The guided training process jointly optimizes for both image reconstruction and classification accuracy, ensuring that the latent representations are both meaningful and effective for quantum processing.
+In our implementation, image data is encoded as atom detuning parameters in a Rydberg Hamiltonian, and quantum embeddings are obtained through expectation values.
+These embeddings are then passed to a linear classifier, enabling faster training and inference compared to deep classical networks.
+Our experiments show that this method outperforms traditional approaches using PCA or unguided autoencoders.
+We also conduct ablation studies to evaluate the impact of quantum and training parameters, demonstrating the robustness and flexibility of the proposed pipeline for real-world medical imaging applications, even in the NISQ era.
+
 \end{abstract}
@@
 \section{Methodology}
-\subsection{System Architecture Overview}
+\subsection{Proposed System Architecture}
 Our proposed pipeline is a Quantum-Guided Autoencoder with a 
 Reservoir Surrogate (QGARS). It is a hybrid architecture that 
-synergizes classical autoencoding techniques with quantum 
-reservoir processing to achieve efficient learning of features
-that work well with the reservoir computing layer whose embeddings
-will be used to perferm a classification task.
+synergizes classical autoencoding techniques with quantum reservoir processing to achieve efficient learning of features.
+These are expected to better match with the reservoir computing layer, whose embeddings are used to perform a classification task.
+
 The system is designed to overcome the inherent non-differentiability of quantum operations, which traditionally impede gradient-based training methods.
@@
             \item \textbf{Encoder:}
-            The encoder network \(\varepsilon_\omega(x)\)
+            The encoder network \(\varepsilon_\omega(\mathbf{x})\)
             transforms high-dimensional input data \(\mathbf{x} \in \mathbb{R}^{n}\) into 
@@
             \item \textbf{Decoder:}
-            The decoder network \(\mathcal{D}_\rho(z)\)
-            Reconstructs the input data from the latent representation, producing
-            \(\mathbf{\hat{x}} \in \mathbb{R}\).
+            The decoder network \(\mathcal{D}_\rho(\mathbf{z})\)
+            reconstructs the input data from the latent representation, producing
+            \(\mathbf{\hat{x}} \in \mathbb{R}^{n}\).
         \end{itemize}
@@
             \item \textbf{Quantum Evolution:}
-            Evolves the system under a Hamiltonian \(H(\Delta)\), resulting in a quantum 
-            state \(|\psi(\Delta)\rangle\).
+            Evolves the system under a Hamiltonian \(H(\mathbf{\Delta)}\), resulting in a quantum 
+            state \(|\psi(\mathbf{\Delta})\rangle\).
             \item \textbf{Measurement:}
@@
     \item \textbf{Surrogate Model:}
-    A differentiable classical neural network \( f_{\text{sur}}(\mathbf{z}; \theta_{\text{sur}}) \)
-    where \(\theta_{\text{sur}}\), which are it's parameters, are tuned to approximate the 
+    A differentiable classical neural network \( f_{\text{sur}}(\mathbf{z}; \bm{\theta}_{\text{sur}}) \),
+    where \(\bm{\theta}_{\text{sur}}\) are parameters tuned to approximate the 
     mapping from \(\mathbf{z}\) to \(\mathbf{q}\), effectively emulating
@@
 
-\begin{figure*}[!ht]
+\begin{figure*}[!t]
     \centering
@@
     \label{fig:qgars_pipeline}
-    \caption{Overview of the Quantum-Guided Autoencoder Reservoir Computing System.}
+    \caption{
+    Overview of the QGARS pipeline.
+(a) Classical autoencoder: $\varepsilon_{\omega}(\mathbf{x})$ is the encoder network that maps input $\mathbf{x}$ to latent code $\mathbf{z}$, and $\mathcal{D}_{\rho}(\mathbf{z})$ is the decoder network reconstructing $\hat{\mathbf{x}}$.
+(b) Rydberg atom quantum reservoir: the latent code $\mathbf{z}$ is mapped to detuning parameters $\Delta_i$, the system evolves under the Hamiltonian in Eq.~\eqref{eq:rydberg_hamiltonian}, and measurements of observables $\langle Z_i\rangle,\langle Z_iZ_j\rangle,\langle Z_iZ_jZ_k\rangle$ yield quantum embeddings.
+(c) Surrogate model: a feedforward neural network $f_{\mathrm{sur}}(\mathbf{z};\theta_{\mathrm{sur}})$ trained every $N$ epochs to approximate the quantum embeddings.
+(d) Linear classifier: maps surrogate outputs $\mathbf{q}'$ to class predictions, closing the loop with total loss $\mathcal{L}=(1-\lambda)\mathcal{L}_R+\lambda\mathcal{L}_C$.
+    }
 \end{figure*}
@@
 \begin{itemize}
-    \item \textbf{The Decoder Network \(\mathcal{D}_\rho(z)\):}
-    that tries to reconstruct the original input data.
-    \item \textbf{The Quantum Reservoir:}
-    that extracts highly dimensional quantum embeddings and
+    \item \textbf{The Decoder Network \(\mathcal{D}_\rho(z)\),}
+    which tries to reconstruct the original input data.
+    \item \textbf{The Quantum Reservoir,}
+    which extracts highly dimensional quantum embeddings and
     passes them to a linear classifier to perform a classification
@@
 \subsubsection{Surrogate Model Architecture}
-In our pipeline, the surrogate model \( f_{\text{sur}}(\mathbf{z}; \theta_{\text{sur}}) \) 
-is implemented as a feedforward neural network 
-designed to learn the mapping 
-from the autoencoder's latent space to the quantum 
-embeddings produced by the quantum reservoir. 
+In our pipeline, the surrogate model \( f_{\text{sur}}(\mathbf{z}; \theta_{\text{sur}}) \) is implemented as a feedforward neural network designed to learn the mapping from the autoencoder's latent space to the quantum embeddings produced by the quantum reservoir.
 \begin{equation}
@@
 \subsection{Rydberg Hamiltonian and Quantum Dynamics}
+Although it can seem that the surrogate model is enough to cover the job of the quantum reservoir, that is far from the truth.
+while a fixed classical reservoir can nonlinearly mix inputs, its capacity is polynomially bounded.
+In contrast, our neutral-atom reservoir leverages many-body quantum dynamics to encode data in a space whose dimension grows exponentially with atom number, unlocking patterns that classical reservoirs would need unreasonably complex architectures to approximate.
+This means that the quantum reservoir is a crucial part of our pipeline and that the surrogate is just an approximation to aid in the autoencoder training process.
+
 The quantum reservoir in our architecture is implemented
@@
   \includegraphics[width=0.32\textwidth]{qgars_qrc_tsne.pdf}
-  \caption{t-SNE projections of the 4 700-D QRC embeddings for (left) PCA + QRC, (center) AE + QRC, (right) QGARS + QRC.}
+  \caption{3-dimensional t-SNE projections of the 4768-dimensional QRC embeddings for (left) PCA + QRC, (center) AE + QRC, (right) QGARS + QRC}
   \label{fig:embeddings_tsne}
@@
   \centering
-  \includegraphics[width=1\linewidth]{paper/images/results/generated_polyp_dataset/guided_lambda_log_effect.pdf}
+  \includegraphics[width=1\linewidth]{paper/images/results/generated_polyp_dataset/guided_lambda_log_mu_effect.pdf}
   \caption{Accuracy as a function of the reconstruction/classification trade-off parameter $\lambda$.}
@@
     AE + Linear       &  0.0 ± 0.0     &  0.0 ± 0.0   &  0.0 ± 0.0    \\
+    PCA + NN          &  0.0 ± 0.0     &  0.0 ± 0.0   &  0.0 ± 0.0    \\
     AE + NN           &  0.0 ± 0.0     &  0.0 ± 0.0   &  0.0 ± 0.0    \\
@@
 
+% ADICIONAR COMENTÁRIO SOBRE COMO É MELHOR ESTUDAR MAIS O PARÂMETRO DE LAMBDA
+
 %============================================

### 2025-05-29 parameter_sweep/data_analysis_scripts/classical_qrc_losses_plotter.py
better guided lambda analysis

@@ -1,0 +1,276 @@
+import json
+import os
+import argparse
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+from typing import Dict, Any, Optional, List, Tuple
+from matplotlib.ticker import MaxNLocator
+
+def load_metrics(filepath: str) -> Dict[str, Any]:
+    """Load metrics from a JSON file."""
+    try:
+        with open(filepath, 'r') as f:
+            return json.load(f)
+    except Exception as e:
+        print(f"Error loading metrics file: {e}")
+        return {}
+
+def plot_losses(metrics: Dict[str, Any], 
+                output_dir: str, 
+                filename_prefix: str = "classifier", 
+                log_scale: bool = True,
+                smooth_window: int = 0,
+                dpi: int = 300) -> Tuple[str, str]:
+    """
+    Plot the loss evolution for linear, QRC, and NN classifiers.
+    
+    Args:
+        metrics: Dictionary containing classifier metrics
+        output_dir: Directory to save the plots
+        filename_prefix: Prefix for the output filenames
+        log_scale: Whether to use log scale for y-axis
+        smooth_window: Window size for moving average smoothing (0 for no smoothing)
+        dpi: DPI for PNG output
+        
+    Returns:
+        Tuple of (pdf_path, png_path) where the plots were saved
+    """
+    # Create output directory if it doesn't exist
+    os.makedirs(output_dir, exist_ok=True)
+    
+    # Set the style for a professional look
+    sns.set_style("whitegrid")
+    plt.rcParams['font.family'] = 'sans-serif'
+    plt.rcParams['font.sans-serif'] = ['Arial', 'DejaVu Sans', 'Liberation Sans']
+    
+    # Create figure and axis with appropriate size
+    fig, ax = plt.subplots(figsize=(10, 6))
+    
+    # Define colors and line styles for different classifiers
+    styles = {
+        "linear": {"color": "#1f77b4", "linestyle": "-", "marker": "o", "markersize": 3, "markevery": 10},
+        "QRC": {"color": "#d62728", "linestyle": "-", "marker": "s", "markersize": 3, "markevery": 10},
+        "NN": {"color": "#2ca02c", "linestyle": "-", "marker": "^", "markersize": 3, "markevery": 10}
+    }
+    
+    # Dictionary to store min loss values and epochs for annotation
+    min_loss_data = {}
+    
+    # Apply smoothing function if window size > 0
+    def smooth(y, window_size):
[…]

### 2025-05-29 parameter_sweep/data_analysis_scripts/guided_lambda_analysis.py

@@ -187,8 +187,26 @@ def plot_guided_mu_effect(df: pd.DataFrame, output_path: Optional[str] = None):
             y_offset = 10
             x_offset = 0
             
+            # Special case for lambda = 0.4
+            if abs(lambda_val - 0.4) < 0.001:  # Use a small tolerance for floating point precision
+                x_offset = -20  # Move label to the left
+                y_offset = 0    # Keep at same height
+            if abs(lambda_val - 0.999999) < 0.001:
+                x_offset = -2
+                y_offset = 0
+            if abs(lambda_val - 0.9999) < 0.001:
+                x_offset = -2
+                y_offset = 0
+                
+            if abs(lambda_val - 1) < 0.001:
+                x_offset = -3
+                y_offset = 0
+            if abs(lambda_val - 0.2) < 0.001:
+                x_offset = -4
+                y_offset = 0
+                
             # For points very close to y-axis, offset to the right
-            if x < 0.05:  # Adjust threshold as needed
+            elif x < 0.05:  # Adjust threshold as needed
                 x_offset = 15
                 y_offset = 0
             
@@ -282,7 +300,7 @@ def plot_guided_mu_effect(df: pd.DataFrame, output_path: Optional[str] = None):
         plt.xlim(0.001, 100)  # Upper limit to 10^2
     
     # Set y-axis limits to 0-100%
-    plt.ylim(70, 95)
+    plt.ylim(70, 90)
     
     # Add secondary x-axis with lambda values
     ax1 = plt.gca()

### 2025-05-31 paper/images/diagrams/autoencoder_architecture.tex
draft
Updates from Overleaf

--- 2025-05-27 paper/autoencoder_architecture.tex
@@
-\begin{figure}[!t]
+% FIND WHITE NODES
+
+\begin{figure}[!tb]
     \centering
@@
             \draw[ fill={rgb,255:red,246; green,245; blue,244} ] (4.75,11.25) rectangle (6,10);
-            \node[font=\LARGE] at (5,12.25) {x};
+            \node[font=\LARGE] at (5,12.25) {$\mathbf{x}$};
             \draw[->, >=Stealth] (6,11) -- (8,11);
-            \draw(8,12.25) rectangle  node {\Huge $f_{\theta}(x)$} (10.5,9.75);
+            \draw(8,12.25) rectangle  node {\Huge $f_{\bm\theta}(\bm{x})$} (10.5,9.75);
             \node[font=\large] at (9.5,11.25) {};
@@
             \draw[short] (10.5,9.75) -- (13.75,10.5);
-            \draw(13.75,11.75) rectangle  node {\LARGE z} (14.5,10);
-            \draw(17.75,12.25) rectangle  node {\Huge $g_{\phi}(z)$} (20.25,9.75);
+            \draw(13.75,11.75) rectangle  node {\LARGE $\bm{z}$} (14.5,10);
+            \draw(17.75,12.25) rectangle  node {\Huge $g_{\bm\phi}(\bm{z})$} (20.25,9.75);
             \draw[short] (17.75,12.25) -- (14.5,11.5);
@@
             \draw[ fill={rgb,255:red,154; green,153; blue,150} ] (22.25,11.25) rectangle (23.5,10);
-            \node[font=\LARGE] at (23,12.25) {$\hat{x}$};
+            \node[font=\LARGE] at (23,12.25) {$\bm{\hat{x}}$};
             \draw[<->, >=Stealth, dashed] (5.25,11.75) .. controls (7.5,16.25) and (21.75,15.75) .. (23,11.75)node[pos=0.5, fill=white]{$\mathcal{L}$};
-            \node[font=\huge] at (17.75,51.5) {};
+            %\node[font=\huge] at (17.75,51.5) {};
             \node[font=\Large] at (14,12.25) {Latent};
@@
 \caption{
-    Autoencoder architecture. The encoder network $f_{\theta}(x)$ 
-    compresses the input $x$ into a latent representation $z$, while the decoder 
-    network $g_{\phi}(z)$ reconstructs the input from the latent representation, 
+    Autoencoder architecture. The encoder network $f_{\bm\theta}(\bm{x})$ 
+    compresses the input $\bm{x}$ into a latent representation $\bm{z}$, while the decoder 
+    network $g_{\bm\phi}(\bm{z})$ reconstructs the input from the latent representation, 
     producing $\hat{x}$. The reconstruction loss $\mathcal{L}$ measures the 
-    compresses the input $x$ into a latent representation $z$, while the decoder 
-    network $g_{\phi}(z)$ reconstructs the input from the latent representation, 
-    producing $\hat{x}$.
+    compresses the input $\bm{x}$ into a latent representation $\bm{z}$, while the decoder 
+    network $g_{\bm\phi}(\bm{z})$ reconstructs the input from the latent representation, 
+    producing $\hat{\bm{x}}$.
 }

### 2025-05-31 paper/main.tex
Updates from Overleaf

--- 2025-05-27 paper/main.tex
@@
 \usepackage{inputenc}
+\usepackage{bm}
+%\usepackage{amsmath}
+
+\newcommand{\ket}[1]{\left\lvert #1 \right\rangle}
+\newcommand{\ketbra}[2]{\left\lvert #1 \right\rangle \left\langle #2 \right\rvert}
 
@@
 \begin{abstract}
-We introduce a quantum-classical hybrid 
-architecture for medical image classification 
-based on neutral atom quantum processors. 
-This approach is designed to address the challenges of 
-medical imaging, with a particular focus on tasks such 
-as polyp detection and classification.
-By integrating an autoencoder guided by a quantum 
-reservoir, the pipeline learns compact and discriminative 
-representations of image data that are also well-suited for quantum reservoir 
-computing. To overcome the non-differentiability of quantum 
-measurements, we circumvent this `gradient barrier' 
-by incorporating a differentiable surrogate model that simulates the 
-behaviour of the quantum layer, enabling end-to-end backpropagation. 
-The guided training process jointly optimizes for both image 
-reconstruction and classification accuracy, ensuring that the latent 
-representations are both meaningful and effective for quantum processing. 
-In our implementation, image data is encoded as atom detuning parameters 
-in a Rydberg Hamiltonian, and quantum embeddings are obtained through expectation 
-values. These embeddings are then passed to a linear classifier, 
-enabling faster training and inference compared to deep classical networks.
-Our experiments show that this method outperforms traditional 
-approaches using PCA or unguided autoencoders. We also conduct 
-ablation studies to evaluate the impact of quantum and training 
-parameters, demonstrating the robustness and flexibility of the 
-proposed pipeline for real-world medical imaging applications, 
-even in the NISQ era.
+% We introduce a quantum-classical hybrid
+% architecture for medical image classification based on neutral-atom quantum processors.
+This approach is designed to address the challenges of medical imaging, with a particular focus on tasks such as polyp detection and classification.
+By integrating an autoencoder guided by a quantum reservoir, the pipeline learns compact and discriminative representations of image data that are also well-suited for quantum reservoir computing.
+To overcome the non-differentiability of quantum measurements, we circumvent this `gradient barrier' by incorporating a differentiable surrogate model that simulates the behaviour of the quantum layer, enabling end-to-end backpropagation.
+The guided training process jointly optimizes for both image reconstruction and classification accuracy, ensuring that the latent representations are both meaningful and effective for quantum processing.
+In our implementation, image data is encoded as atom detuning parameters in a Rydberg Hamiltonian, and quantum embeddings are obtained through expectation values.
+These embeddings are then passed to a linear classifier, enabling faster training and inference compared to deep classical networks.
+Our experiments show that this method outperforms traditional approaches using PCA or unguided autoencoders.
+We also performed ablation studies to evaluate the impact of quantum and training parameters, demonstrating the robustness and flexibility of the proposed pipeline for real-world medical imaging applications, even in the NISQ era.
+
+%We introduce a hybrid quantum-classical system using neutral-atom quantum processors for medical image classification, focusing on tasks like polyp detection.
+Our approach integrates an autoencoder guided by a quantum reservoir.
+
+We introduce a hybrid quantum-classical pipeline, based on neutral-atom reservoir computing, for medical image classification, focusing on the binary classification task of polyp detection.
+To deal effectively with the high dimensionality, we integrate a guided auto-encoder.
+This pipeline learns compact and discriminative representations of image data that are also well-suited for quantum reservoir computing.
+A key challenge in such systems is the non-differentiable nature of quantum measurements, which creates a `gradient barrier' for standard training.
+We overcome this barrier by incorporating a differentiable surrogate model that emulates the quantum layer, enabling end-to-end backpropagation through the entire system.
+This guided training process is jointly optimized for classification accuracy and for faithful image recovery from the auto-encoder.
+The learned latent representations are encoded as pulse detuning parameters within a Rydberg Hamiltonian, and quantum embeddings are subsequently obtained through expectation values.
+These embeddings are then passed to a linear classifier.
+Our simulations show that this method outperforms some traditional approaches that use PCA or unguided autoencoders.
+We also conduct ablation studies to assess the impact of various quantum and training parameters, demonstrating the robustness and flexibility of our proposed pipeline for real-world medical imaging applications, even in the current NISQ era.
+
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Reservoir Computing, Quantum-Guided Autoencoding,
-Neutral Atoms, Autoencoder, Dimensionality Reduction, 
-Quantum Machine Learning, Hybrid Quantum-Classical Algorithms, 
-Medical Image Classification, Quantum Surrogate Models
+Reservoir Computing, Quantum-Guided Autoencoding, Neutral Atoms, Autoencoder, Dimensionality Reduction, Quantum Machine Learning, Hybrid Quantum-Classical Algorithms, Medical Image Classification, Quantum Surrogate Models
 \end{IEEEkeywords}
 
-%============================================
-% INTRODUCTION
-%============================================
 \section{Introduction}
 
-\subsection{Background and Motivation}
-Advances in medical imaging have significantly improved 
-disease diagnosis and treatment planning. For conditions 
-like colorectal cancer, early detection of polyps through 
-colonoscopy image analysis is critical for reducing mortality~\cite{estevaGuideDeepLearning2019}. 
-Deep learning techniques, especially autoencoders, are widely 
-used to extract compressed, informative features from 
-high-dimensional images for classification and segmentation 
-tasks~\cite{bengioLearningDeepArchitectures}. However, classical neural networks may struggle 
-to capture intricate correlations in complex medical data.
+%\subsection{Background and Motivation}
+Advances in medical imaging have significantly improved disease diagnosis and treatment planning.
+For conditions like colorectal cancer, early detection of polyps through colonoscopy image analysis is critical for reducing mortality~\cite{estevaGuideDeepLearning2019a}.
+Deep learning techniques, especially autoencoders, are widely used to extract compressed, informative features from high-dimensional images for classification~\cite{bengioLearningDeepArchitecturesa}.
+However, classical neural networks may struggle to capture intricate correlations in complex medical data~\cite{meiFrameworkProcessingLargescale2024}.
 
-Quantum computing offers novel opportunities 
-for machine learning, particularly through quantum reservoir 
-computing (QRC), where a physical quantum system processes 
-classical inputs into high-dimensional nonlinear embeddings~\cite{tanakaRecentAdvancesPhysical2019,fujiiHarnessingDisorderedQuantum2017}. Recent works show that analog quantum systems, 
-such as neutral-atom platforms, can serve as untrained 
-reservoirs with rich dynamics for temporal and pattern 
-recognition tasks \cite{domingoOptimalQuantumReservoir2022,kornjavcaLargescaleQuantumReservoir2024}. 
-In hybrid approaches, a classical encoder compresses 
-image data, and a quantum reservoir expands 
-the encoded features into a higher-dimensional space, 
-potentially boosting classification performance.
+Quantum computing offers novel opportunities for machine learning, particularly through quantum reservoir computing (QRC), where a physical quantum system processes classical inputs into high-dimensional nonlinear embeddings~\cite{tanakaRecentAdvancesPhysical2019,fujiiHarnessingDisorderedQuantum2017}.
+Recent works show that analog quantum systems, such as neutral-atom platforms, can serve as untrained reservoirs with rich dynamics for temporal and pattern recognition tasks~\cite{domingoOptimalQuantumReservoir2022,kornjavcaLargescaleQuantumReservoir2024}.
 
-A major challenge in such hybrid quantum-classical 
-models is the non-differentiability of quantum measurements, 
-which obstructs gradient-based optimization. Additionally, 
-tuning quantum parameters can suffer from barren plateaus, where 
-gradients vanish in high-dimensional Hilbert spaces \cite{mccleanBarrenPlateausQuantum2018}. 
-To address this, we introduce a classical neural surrogate 
-that emulates the quantum reservoir's input-output behavior. 
-This surrogate enables end-to-end training via backpropagation, 
-while the quantum system remains fixed and non-trainable.
+In hybrid approaches, a classical encoder compresses image data, and a quantum reservoir expands the encoded features into a higher-dimensional Hilbert space, potentially boosting classification performance, as these expanded features might capture high-order correlations and nonlinearities that classical methods cannot tractably represent.
 
-\subsection{Contributions of This Work}
-We propose a quantum-guided autoencoder architecture that 
-integrates a classical image encoder with a neutral-atom
-quantum reservoir.
+A major challenge in such hybrid quantum-classical models is the non-differentiability of quantum measurements, which obstructs gradient-based optimization.
+Additionally, tuning quantum parameters can suffer from barren plateaus, where gradients vanish in high-dimensional Hilbert spaces~\cite{mccleanBarrenPlateausQuantum2018}.
+To address this, we introduce a classical neural surrogate that emulates the quantum reservoir's input-output behavior.
+This surrogate enables end-to-end training via backpropagation, while the quantum system remains fixed and non-trainable and is only used to train the surrogate model and in downstream classifiers.
 
-A classical surrogate network of the reservoir itself enables 
-gradient flow through the whole model during training.
+Critically, the physical reservoir’s exponentially large state‐space (and associated many-body correlations) cannot be classically emulated, hence the surrogate is merely a differentiable proxy during training, not a replacement of the real quantum dynamics at inference.
 
-Our results illustrate the viability of QRC for real-world medical 
-tasks and offer a scalable path to hybrid quantum-classical learning,
-even in the noisy intermediate-scale quantum (NISQ) era.
+%\subsection{Contributions of This Work}
+We propose a quantum-guided autoencoder architecture that integrates a classical image encoder with a neutral-atom quantum reservoir.
+
+A classical surrogate network of the reservoir itself enables gradient flow through the whole model during training.
+
+Our results illustrate the viability of QRC for real-world medical tasks and offer a scalable path to hybrid quantum-classical learning, even in the noisy intermediate-scale quantum (NISQ) era.
 
@@
 \section{Background and Related Work}
-
+This section provides an overview of the foundational concepts and prior research relevant to this work.
 \subsection{Principles of Reservoir Computing}
-Reservoir computing is a computational framework 
-derived from recurrent neural networks (RNNs). It 
-involves a fixed, high-dimensional dynamical system—the 
-reservoir—that projects input data into a rich feature 
-space. Only the output layer is trained, simplifying 
-the learning process and reducing computational 
-overhead. This approach is particularly effective 
-for time-series prediction and pattern recognition tasks.
+Reservoir computing is a computational framework derived from recurrent neural networks (RNNs). It involves a fixed, high-dimensional dynamical system (the 
+reservoir) that projects input data into a rich feature space.
+Only the output layer is trained, simplifying the learning process and reducing computational overhead.
+This approach is particularly effective for time-series prediction and pattern recognition tasks.
 
@@
 \( y(t) \in \mathbb{R}^k \) the output. The reservoir dynamics
-and output are given by:
-\begin{equation}
-    x(t) = f(W_{in} u(t) + W_{res} x(t-1))
-\end{equation}
-\begin{equation}
-    y(t) = W_{out} x(t)
-\end{equation}
-
+and output are given by
+\begin{align}
+    x(t) &= f(W_{in} u(t) + W_{res} x(t-1)) \\
+    y(t) &= W_{out} x(t),
+\end{align}
+%
 %Where ff is a nonlinear activation function, WinWin​ and WW are fixed input and reservoir weight matrices, and WoutWout​ is the trained output weight matrix.
-
-Where \( f \) is a nonlinear activation function,
+%
+where \( f \) is a nonlinear activation function,
 \( W_{in} \) and \( W_{res} \) are fixed input and reservoir weight matrices, 
 and \( W_{out} \) is the trained output weight matrix.
+A diagram of a typical reservoir computing architecture is shown in Fig.~\ref{fig:reservoir_architecture}.
 
-A diagram of a typical Reservoir Computing architecture is shown in `Fig.-\ref{fig:reservoir_architecture}'.
+\input{images/diagrams/reservoir_architecture.tex}
 
-\input{reservoir_architecture.tex}
+\subsection{Quantum Reservoir Computing with neutral atoms}
 
-\subsection{Quantum Reservoir Computing}
+%\subsubsection{Main Differences}
 Quantum Reservoir Computing (QRC) extends the reservoir 
-computing paradigm into the quantum domain. By leveraging quantum systems' 
-inherent properties, such as superposition and entanglement, QRC aims 
-to enhance computational capabilities. Implementations using quantum 
-oscillators have shown promise in solving complex learning tasks, 
-offering advantages over classical counterparts. Notably, large-scale 
-experiments utilizing neutral-atom analog quantum computers have demonstrated 
-the scalability and effectiveness of QRC in various machine learning applications~\cite{kornjavcaLargescaleQuantumReservoir2024}
+computing paradigm into the quantum domain. 
+%By leveraging quantum systems' inherent properties, such as superposition and entanglement, 
+QRC aims to enhance computational capabilities by accessing larger state spaces and different types of non-linearities. 
+%Implementations using quantum oscillators have shown promise in solving complex learning tasks, offering advantages over classical counterparts. 
+Notably, large-scale experiments utilizing neutral-atom analog quantum computers have demonstrated the scalability and effectiveness of QRC in various machine learning applications~\cite{kornjavcaLargescaleQuantumReservoir2024}
 
@@
 \begin{equation}
-    |\psi(t+1)\rangle = U |\psi(t)\rangle = e^{-iH\Delta t} |\psi(t)\rangle,
+    \ket{\psi(t+\Delta t)}
+    = U \let{\psi(t)}
+    = e^{-iH\Delta t} \ket{\psi(t)},
 \end{equation}
-where \( U \) is the unitary evolution operator. Measurements of observables \( \hat{O} \) yield outputs:
+where \( U \) is the unitary evolution operator and \(\Delta t\) is the difference between consecutive timestamps. Measurements of observables \( \hat{O} \) yield the outputs
 \begin{equation}
@@
 
-\subsection{Reservoir Computing with Neutral Atoms}
-As Quantum Reservoir Computing (QRC) leverages the complex 
-dynamics of quantum systems to process information, 
-extending the Classical Reservoir Computing 
-paradigm into the quantum domain. Neutral atom platforms, 
-particularly those utilizing Rydberg states, 
-have emerged as promising candidates for implementing 
-QRC due to their scalability and controllable interactions.
+%\subsubsection{Quantum Reservoir Computing with Neutral Atoms}
+Neutral atom platforms, particularly those utilizing Rydberg states, have emerged as promising candidates for implementing QRC due to their scalability and controllable interactions.
 
-In the work by M. Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}, 
-a large-scale, gradient-free QRC algorithm was 
-developed and experimentally implemented on a 
-neutral-atom analog quantum computer. 
-This system achieved competitive performance across 
-various machine learning tasks, 
-including classification and time-series prediction, 
-demonstrating effective learning with increasing system 
-sizes up to 108 qubits.
+In the work by M.
+Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}, a large-scale, gradient-free QRC algorithm was developed and experimentally implemented on a neutral-atom analog quantum computer.
+This system achieved competitive performance across various machine learning tasks, including classification and time-series prediction, demonstrating effective learning with increasing system sizes up to $108$ qubits.
 
-The dynamics of such neutral atom systems can be 
-described by the Rydberg Hamiltonian, which captures 
-the essential physics of laser-driven interactions 
-among atoms in Rydberg states. Following Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024}
-the Hamiltonian for a system of neutral-atoms is given by:
+% The dynamics of such neutral atom systems can be described by the Rydberg Hamiltonian, which captures 
+% the essential physics of laser-driven interactions among atoms in Rydberg states.
+Following Kornjača et al.~\cite{kornjavcaLargescaleQuantumReservoir2024} the Hamiltonian for a system of neutral-atoms is given by:
 
-\begin{equation}
-    \begin{aligned}
-        H(t) &= \dfrac{\Omega(t)}{2}\sum_j \left(\ket{g_j}\bra{r_j}+\ket{r_j}\bra{g_j}\right) \\
-             &{{\quad}} + \sum_{j<k}V_{jk}n_jn_k - \sum_j \left[\Delta_{\mathrm{g}}(t) + \alpha_j\Delta_{\mathrm{l}}(t)\right] n_j,
-    \end{aligned}
+The dynamics of such laser-driven neutral atoms, working in the Rydberg state regime, is given by the Hamiltonian~\cite{kornjavcaLargescaleQuantumReservoir2024}
+
+\begin{align}
+    H(t)
+    &=
+    \frac{\Omega(t)}{2}
+    \sum_j
+    \left(
+        \ketbra{g_j}{r_j}
+        + \ketbra{r_j}{g_j}
+    \right) \nonumber \\
+    & \quad + \sum_{j<k}
+    V_{jk} n_j n_k
+    - \sum_j
+    \left[
+        \Delta_{\mathrm{g}}(t)
+        + \alpha_j \Delta_{\mathrm{l}}(t)
+    \right] n_j,
 \label{eq:rydberg_hamiltonian}
-\end{equation}
-where \( \Omega(t) \) is the global Rabi drive amplitude between a 
-ground state \( |g_j\rangle \) and a highly-excited Rydberg state 
-\( |r_j\rangle \) of an atom,
-\( n_j = |r_j\rangle \langle r_j| \), 
-\( V_{jk} = C/\|r_j - r_k\|^6 \) describes the van der Waals interactions
+\end{align}
+where $\Omega(t)$ is the global Rabi drive amplitude between a 
+ground state $\ket{g_j}$ and a highly-excited Rydberg state 
+$\ket{r_j}$ of an atom $j$,
+$n_j = \ketbra{r_j}{r_j}$, and 
+\( V_{jk} = C_6/\|r_j - r_k\|^6 \) describes the van der Waals interactions
 between atoms, and the detuning is split into a global term 
@@
 much information as possible. This process enhances 
-computational efficiency, mitigates the curse of 
-dimensionality, and facilitates data visualization
+computational efficiency and facilitates data visualization.
 
@@
 
-Given an input \( x \in \mathbb{R}^d \), the encoder maps 
-\( x \) to a latent representation \( z \in \mathbb{R}^k \) 
+Given an input \( \bm x \in \mathbb{R}^d \), the encoder maps 
+\( \bm x \) to a latent representation \( \bm z \in \mathbb{R}^k \) 
 (where \( k < d \)):
@@
 \begin{equation}
-    z = f_\theta(x)
+    \bm z = \varepsilon_{\bm\omega}(\bm x).
 \end{equation}
-
-The decoder then reconstructs the input:
+%
+The decoder then reconstructs the input $\bm{\hat{x}} \in \mathbb{R}^d$:
 \begin{equation}
-    \hat{x} = g_\phi(z)
+    \bm{\hat{x}} = \mathcal{D}_{\bm \rho} (\bm z),
 \end{equation}
+where we follow the notation from~\cite{belisGuidedQuantumCompression2024}.
 The network is trained to minimize the reconstruction loss:
 \begin{equation}
-    \mathcal{L}(x, \hat{x}) = \| x - \hat{x} \|^2
+    \mathcal{L}(\bm{x}, \bm{\hat{x}}) = \| \bm{x} - \bm{\hat{x}} \|^2.
 \end{equation}
+%
+This architecture of an autoencoder is illustrated in Fig. 2 a). %~\ref{fig:qgars}~(panel (a)).
 
-This typical architecture of an autoencoder is illustrated in 
-`Fig.-\ref{fig:autoencoder_architecture}'.
-
-\input{autoencoder_architecture.tex}
+%\input{images/diagrams/autoencoder_architecture.tex}
 
@@
 like image compression, denoising, and anomaly 
-detection~\cite{hintonReducingDimensionalityData2006, estevaGuideDeepLearning2019}.
+detection~\cite{hintonReducingDimensionalityData2006, estevaGuideDeepLearning2019a}.
 
 \subsubsection{Quantum-Guided Autoencoding}
-Quantum-Guided Autoencoders integrate quantum 
-computing principles into the autoencoder framework 
-to leverage quantum advantages in processing and 
-representing data. These models aim to perform 
-dimensionality reduction and classification within a 
-single architecture, enhancing performance on complex 
-datasets.
+Quantum-Guided Autoencoders integrate quantum computing principles into the autoencoder framework to leverage quantum advantages in processing and representing data.
+These models aim to perform dimensionality reduction and classification within a single architecture, enhancing performance on complex datasets.
 
-In the Quantum-Guided Autoencoder (QGAE) model, a 
-classical encoder first reduced the dimensionality
-of the input data. The compressed data is then processed 
-by a parametrized quantum circuit, which acts as the decoder
-and classifier. The quantum circuit transforms the 
-input state \( |\psi_{\text{in}}\rangle \) into an
-output state \( |\psi_{\text{out}}\rangle \) using a unitary operation
+In the Quantum-Guided Autoencoder (QGA) model, a classical encoder first reduced the dimensionality of the input data.
+The compressed data is then processed by a parametrized quantum circuit, which acts as the decoder and classifier.
+The quantum circuit transforms the input state \( |\psi_{\text{in}}\rangle \) into an output state \( |\psi_{\text{out}}\rangle \) using a unitary operation:
 \begin{equation}
-    |\psi_{\text{out}}\rangle = U(\theta) |\psi_{\text{in}}\rangle
+    |\psi_{\text{out}}\rangle = U(\bm\theta) |\psi_{\text{in}}\rangle.
 \end{equation}
-Measurements on \( |\psi_{\text{out}}\rangle \) yield the final
-classification result. The parameters \( \theta \) are
-optimized to minimize a loss function that combines
-reconstruction error and classification accuracy.
+Measurements on \( |\psi_{\text{out}}\rangle \) yield the final classification result.
+The parameters \( \bm\theta \) are optimized to minimize a loss function that combines reconstruction error and classification accuracy.
 
-This approach has demonstrated improved performance 
-over traditional methods in tasks such as identifying 
-the Higgs boson in particle collision data, 
-showcasing the potential of quantum-guided models in 
-handling high-dimensional, complex datasets~\cite{belisGuidedQuantumCompression2024}.
+This approach has demonstrated improved performance over traditional methods in tasks such as identifying the Higgs boson in particle collision data, showcasing the potential of quantum-guided models in handling high-dimensional, complex datasets~\cite{belisGuidedQuantumCompression2024}.
 
@@
 \section{Methodology}
-\subsection{System Architecture Overview}
+\subsection{Proposed System Architecture}
+Our proposed pipeline is a Quantum-Guided Autoencoder with a 
+Reservoir Surrogate (QGARS). It is a hybrid architecture that 
+synergizes classical autoencoding techniques with quantum reservoir processing to achieve efficient learning of features.
+These are expected to better match with the reservoir computing layer, whose embeddings are used to perform a classification task.
 
-Our proposed pipeline integrates classical and quantum 
-components for dimensionality reduction and classification. 
-A classical autoencoder compresses high-dimensional image 
-data into a lower-dimensional latent space. This latent 
-representation is encoded into a Rydberg atom chain, 
-serving as input to a quantum reservoir. The reservoir's 
-dynamics, governed by the Rydberg Hamiltonian, produce 
-quantum embeddings through measurements of quantum observables.
-The overall architecture is illustrated in `Fig.-\ref{fig:qgars_pipeline}'.
+The system is designed to overcome the inherent non-differentiability of quantum operations, which traditionally impede gradient-based training methods.
+The overall architecture is illustrated in Fig. 2 and the \textbf{architectural components} are the following:
+\begin{itemize}
+    \item \textbf{Classical Autoencoder:}
+        \begin{itemize}
+            \item \textbf{Encoder:}
+            transforms high-dimensional input data \(\bm{x} \) into 
+            a lower-dimensional latent representation \(\bm{z}\).
+            \item \textbf{Decoder:}
+            The decoder network \(\mathcal{D}_\rho(\bm{z})\) reconstructs the input data from the latent representation, producing \(\bm{\hat{x}}\).
+        \end{itemize}
 
-To facilitate end-to-end training, we employ a differentiable 
-surrogate model that approximates the quantum reservoir's 
-behavior, enabling gradient flow from the classification 
-loss back to the autoencoder. The surrogate's outputs feed 
-into a linear classifier trained to map quantum embeddings to class labels.
+    \item \textbf{Quantum Reservoir Layer:}
+        \begin{itemize}
+            \item \textbf{Parameter Mapping:}
+            \(g(\bm{z})\) maps the latent representation \(\bm{z}\) to the local detuning frequencies
+            \(\Delta_{\mathrm{i}}\) of the Rydberg Atoms.
+            \item \textbf{Quantum Evolution:}
+            Evolves the system under a Hamiltonian \(H(\bm{\Delta)}\), resulting in a quantum 
+            state \(|\psi(\bm{\Delta})\rangle\).
+            \item \textbf{Measurement:}
+            Performs measurements on the quantum state to obtain embeddings \(\bm{q} \in \mathbb{R}^k\), which serve as training targets for the Surrogate Model or as 
+            inputs for the classification task later on.
+        \end{itemize}
 
-The autoencoder is trained to minimize a combined loss 
-function that interpolates between reconstruction loss and 
-classification loss, promoting representations that are both 
-informative for input reconstruction and effective for quantum 
-reservoir computing.
+    \item \textbf{Surrogate Model:}
+    A differentiable classical neural network \( f_{\text{sur}}(\bm{z}; \bm{\theta}_{\text{sur}}) \),
+    where \(\bm{\theta}_{\text{sur}}\) are parameters tuned to approximate the 
+    mapping from \(\bm{z}\) to \(\bm{q}\), effectively emulating
+    the quantum layer's behaviour.
 
-Once trained, the encoder compresses data into the latent 
-space, which is then processed by the actual quantum reservoir. 
-The resulting quantum embeddings are used to train a linear 
-classifier for final predictions.
+    \item \textbf{Linear Classifier:}
+    A classical neural network that maps the quantum embeddings \(\bm{q}\) to 
+    predicted class labels \(\hat{y}\).
 
-\begin{figure*}[!h]
+\end{itemize}
+The following sections give a more detailed explanation of each component.
+
+\begin{figure*}[!t]
     \centering
     \resizebox{0.9\textwidth}{!}{
-        \includegraphics{qgars_pipeline_cropped.pdf}
+        \includegraphics{images/diagrams/qgars_pipeline_cropped.pdf}
     }
-    \label{fig:qgars_pipeline}
-    \caption{Overview of the Quantum-Guided Autoencoder Reservoir Computing System.}
+    \label{fig:qgars}
+    \caption{
+    Overview of the QGARS pipeline.
+(a) Classical autoencoder: $\varepsilon_{\bm\omega}(\bm{x})$ is the encoder network that maps input $\bm{x}$ to latent code $\bm{z}$, and $\mathcal{D}_{\bm\rho}(\bm{z})$ is the decoder network reconstructing $\hat{\bm{x}}$.
+(b) Rydberg atom quantum reservoir: the latent code $\bm{z}$ is mapped to detuning parameters $\Delta_i$, the system evolves under the Hamiltonian in Eq.~\eqref{eq:rydberg_hamiltonian}, and measurements of observables $\langle Z_i\rangle,\langle Z_iZ_j\rangle,\langle Z_iZ_jZ_k\rangle$ yield quantum embeddings.
+(c) Surrogate model: a feedforward neural network $f_{\mathrm{sur}}(\bm{z};\bm\theta_{\mathrm{sur}})$ trained every $N$ epochs to approximate the quantum embeddings.
+(d) Linear classifier: maps surrogate outputs $\bm{q}'$ to class predictions, closing the loop with total loss $\mathcal{L}=(1-\lambda)\mathcal{L}_R+\lambda\mathcal{L}_C$.
+    }
 \end{figure*}
 
-\subsection{Quantum Guided Autoencoder}
-The Quantum Guided Autoencoder (QGA) combines classical 
-preprocessing with quantum processing to leverage 
-the strengths of both paradigms. The classical encoder 
-reduces the dimensionality of the input data, facilitating 
-efficient quantum processing. The quantum circuit 
-then processes the compressed data, aiming to reconstruct 
-the original input and perform classification simultaneously.
+\subsection{Guided Autoencoder}
+The Guided Autoencoder combines classical with quantum processing to leverage the strengths of both paradigms.
+
+Ideally, the latent space representation created by the encoder \(\varepsilon_\omega(\bm{x})\)
+would be passed through two sections of the model:
+\begin{itemize}
+    \item \textbf{The Decoder Network \(\mathcal{D}_\rho(\bm{z})\),}
+    which tries to reconstruct the original input data.
+    \item \textbf{The Quantum Reservoir,}
+    which extracts highly dimensional quantum embeddings and passes them to a linear classifier to perform a classification task and map the embeddings to the image labels.
+\end{itemize}
+
+However, some problems emerge when using a real quantum reservoir directly during the autoencoder training stage, which will be discussed in a later section.
 
@@
 \begin{itemize}
-    \item \( \mathcal{L}_R = \| x - \hat{x} \|^2 \) is the reconstruction loss, measuring the mean squared error between the input \( x \) and its reconstruction \( \hat{x} \).
-    \item \( \mathcal{L}_C = -\sum_i y_i \log(\hat{y}_i) \) is the classification loss, computed as the cross-entropy between the true labels \( y_i \) and the predicted probabilities \( \hat{y}_i \).
+    \item \( \mathcal{L}_R = \| \bm{x} - \hat{\bm{x}} \|^2 \) is the reconstruction loss, measuring the mean squared error between the input \( \bm{x} \) and its reconstruction \( \bm{\hat{x}} \).
+    \item \( \mathcal{L}_C = -\sum_i y_i \log(\hat{y}_i) \) is the classification loss, computed as the cross-entropy between the true labels \( y_i \) and the predicted probabilities \( \hat{y}_i \), which are obtained through a linear mapping of the
+    approximated embeddings \(\bm{q}'\).
     \item \( 0 < \lambda < 1 \) is a hyperparameter that controls the trade-off between reconstruction and classification objectives.
 \end{itemize}    
-
-%\subsubsection{Balancing Reconstruction and Classification}
-
 \subsection{The Gradient Barrier Problem}
-Hybrid quantum-classical models, such as the Quantum Guided Autoencoder 
-(QGAE), aim to leverage quantum computational advantages within 
-classical machine learning frameworks. However, a significant 
-challenge arises due to the non-differentiable nature of quantum 
-operations, commonly referred to as the `gradient barrier'.
+Hybrid quantum-classical models, such as the Quantum-Guided Autoencoder (QGAE), aim to leverage quantum computational advantages within classical machine learning frameworks.
+However, a significant challenge arises due to the non-differentiable nature of quantum
+operations, we refer to this as the `gradient barrier'.
 
@@
 training of hybrid models. The surrogate model 
-solution, proposed in the next section, addresses 
+solution, proposed in the next section~\ref{surrogate_modeling_for_quantum_layers}, addresses 
 this problem.
 
-\subsection{Surrogate Modeling for Quantum Layers}
+\subsection{Surrogate Modeling for Quantum Layers} \label{surrogate_modeling_for_quantum_layers}
 To circumvent the gradient barrier, surrogate models 
@@
 
-Recent studies have demonstrated the efficacy of 
-surrogate models in mitigating the effects of barren plateaus 
-and facilitating the training of hybrid quantum-classical models. 
-These approaches allow for the practical implementation of 
-quantum-enhanced machine learning models by leveraging classical 
-optimization techniques while still capturing quantum computational 
-advantages~\cite{xieQuantumSurrogateDrivenImage2025b}
+Recent studies have demonstrated the efficacy of surrogate models in mitigating the effects of barren plateaus and facilitating the training of hybrid quantum-classical models.
+These approaches leverage
+%for the practical implementation of quantum-enhanced machine learning models by leveraging 
+classical optimization techniques while still capturing quantum computational advantages~\cite{xieQuantumSurrogateDrivenImage2025a}.
 
-\subsubsection{Architecture}
-In our pipeline, the surrogate model is implemented as a 
-feedforward neural network designed to learn the mapping 
-from the autoencoder's latent space to the quantum 
-embeddings produced by the quantum reservoir.
+\subsubsection{Surrogate Model Architecture}
+In our pipeline, the surrogate model \( f_{\text{sur}}(\bm{z}; \bm\theta_{\text{sur}}) \) is implemented as a feedforward neural network designed to learn the mapping $\bm q´$ from the autoencoder's latent space to the quantum embeddings $\bm q$ produced by the quantum reservoir:
+\begin{equation}
+    f_{\text{sur}}(\bm{z}; \bm\theta_{\text{sur}}) = \bm{q}' \approx \bm{q} =  \langle \psi(g(\bm{z}))| \hat{O} |\psi(g(\bm{z}))\rangle
+\end{equation}
+where \(\hat{O}\) represents the observables whose expectation values are to be determined.
 
-Its architecture comprises multiple fully connected 
-layers with nonlinear activation functions, facilitating 
-the approximation of the complex, non-linear 
-transformations inherent in quantum dynamics.
+Its architecture comprises multiple fully connected layers with nonlinear activation functions, facilitating the approximation of the complex, non-linear transformations inherent in quantum dynamics.
 
-This approach draws inspiration from recent studies 
-demonstrating the effectiveness of surrogate models 
-in approximating quantum circuit behaviors, thereby 
-facilitating efficient training of hybrid 
-models.~\cite{xieQuantumSurrogateDrivenImage2025b,schreiberClassicalSurrogatesQuantum2023}
+%This approach draws inspiration from recent studies demonstrating the effectiveness of surrogate models in approximating quantum circuit behaviors, thereby facilitating efficient training of hybrid models~\cite{xieQuantumSurrogateDrivenImage2025a,schreiberClassicalSurrogatesQuantum2023}.
 
@@
     \item Data Input: 
-    Feed a batch of latent data \( z \) from the autoencoder 
+    Feed a batch of latent data \( \bm{z} \) from the autoencoder 
     into the real quantum reservoir, recording the resulting quantum 
@@
     \item Surrogate Training:
-    Train the surrogate model using the recorded quantum
-    embeddings as targets, minimizing the loss function:
+    Train the surrogate model  by minimizing the loss function:
     \begin{equation}
-        \mathcal{L}_{\text{surrogate}} = \| \hat{y} - y \|^2,
+        \mathcal{L}_{\text{sur}} = \| \bm{q}' - \bm{q} \|^2,
     \end{equation}
-    where \( \hat{y} \) are the surrogate model's predictions
-    and \( y \) are the actual quantum embeddings.
 
@@
     Periodically update the surrogate model during training
-    to insure it accurately reflects the quantum reservoir's
-    behavior. This can be done after a fixed number of epochs
-    or when the surrogate's performance on a validation set
-    reaches a certain threshold.
+    to ensure it accurately reflects the quantum reservoir's
+    behavior. 
+    %This can be done after a fixed number of epochs or when the surrogate's performance on a validation set reaches a certain threshold.
 \end{itemize}
@@
 \subsubsection{Gradient Flow Through Surrogate Models}
-By integrating the surrogate model into the training pipeline, 
-we establish a continuous computational graph from the 
-output layer back to the input layer, enabling the use of 
-gradient-based optimization techniques. The surrogate model 
-serves as a differentiable proxy for the quantum reservoir, 
-allowing the classification loss to influence the 
-autoencoder's parameters effectively.
+By integrating the surrogate model into the training pipeline, we establish a continuous computational graph from the output layer back to the input layer, enabling the use of gradient-based optimization techniques.
+The surrogate model serves as a differentiable proxy for the quantum reservoir, allowing the classification loss to influence the autoencoder's parameters effectively.
 
-As a results, we not only achieve end-to-end training of the
-hybrid model but also makes it computationally efficient,
-as we reduce the need for frequent quantum measurements.
+%As a result, we not only achieve end-to-end training of the hybrid model but also makes it computationally efficient, as we reduce the need for frequent quantum measurements.
 
 \subsection{Rydberg Hamiltonian and Quantum Dynamics}
-The quantum reservoir in our architecture is implemented
-using a neutral-atom system, specifically utilizing
-Rydberg atoms to create a programmable quantum reservoir.
-The dynamics of the system are governed by the Rydberg Hamiltonian,
-which describes the interactions between atoms in Rydberg states~\ref{eq:rydberg_hamiltonian}.
+%Although it can seem that the surrogate model is enough to cover the job of the quantum reservoir, that is far from the truth. 
+While a fixed classical reservoir can nonlinearly mix inputs, its capacity is polynomially bounded.
+In contrast, our neutral-atom reservoir leverages many-body quantum dynamics to encode data in a space whose dimension grows exponentially with atom number, unlocking patterns that classical reservoirs would need unreasonably complex architectures to approximate.
+This means that the quantum reservoir is a crucial part of our pipeline and that the surrogate is just an approximation to aid in the autoencoder training process.
+
+%The quantum reservoir in our architecture is implemented using a neutral-atom system, specifically utilizing Rydberg atoms to create a programmable quantum reservoir.
+The dynamics of the system are governed by the Rydberg Hamiltonian, which describes the interactions between atoms in Rydberg states~\ref{eq:rydberg_hamiltonian}.
 
@@
 
-\subsubsection{Quantum Readout Methods}
+\subsubsection{Quantum Readout Methods and Embedding Dimensionality}
 The quantum reservoir's output is obtained by probing, 
@@
 after the systems evolution up to a certain time \( t \).
-
 The readout is performed by measuring specific observables,
 such as $\langle Z_i \rangle$, $\langle Z_i \, Z_j\rangle$, and $\langle Z_i \, Z_j \, Z_k \rangle$, which measure respectively one-, two- and three-qubit correlations.
-% \begin{itemize}
-%     \item \textbf{Single-atom Measurements}: 
-%     \( \langle Z_i \rangle \) for each atom \( i \), where \( Z_i \) is the Pauli-Z operator.
-%     \item \textbf{Two-atom Correlations}: 
-%     \( \langle Z_i Z_j \rangle \) for pairs of atoms \( (i, j) \), capturing correlations between their states.
-%     \item \textbf{Three-atom Correlations}:
-%     \( \langle Z_i Z_j Z_k \rangle \) for triplets of atoms \( (i, j, k) \), providing higher-order correlations.
-% \end{itemize}
+
+This means that the dimensionality of the QRC embeddings generated in our pipeline is a direct consequence of the number of qubits ($N$) employed in the neutral-atom chain, the specific quantum mechanical observables measured, and the number of discrete time steps ($T$) over which these measurements are collected every \(\Delta t\).
+
+If we measure all of the $3$ proposed observables, the total count of the values measured at a single time step $O_N$ is the sum of the counts for each type of observable:
+\begin{equation}
+    \begin{aligned}
+    O_N &= N + \binom{N}{2} + \binom{N}{3}.
+    \end{aligned}
+    \label{eq:embeddings_per_timestep}
+\end{equation}
+This set of $O_N$ measurements is repeated at $T$ successive time steps during the evolution of the quantum reservoir.
+Consequently, the total dimensionality, $\lvert \bm q \rvert$, of the final QRC embedding vector is:
+\begin{equation}
+    \lvert \bm q \rvert = T O_N.
+    \label{eq:total_embedding_dimension}
+\end{equation}
+This formulation shows that the embedding dimension scales polynomially with the number of qubits $N$ and the number of timesteps $T$ as $\mathcal{O}(TN^3)$ for the considered observables.
 
@@
 \subsection{Datasets}
-We evaluate our proposed architecture on three primary datasets:
+We evaluate our proposed architecture on three datasets:
 \begin{enumerate}
-    \item \textbf{Synthetic Polyp Dataset:} 
+    \item \textbf{Synthetic Polyps:} 
     A synthetic dataset of polyp images generated to simulate 
-    realistic medical imaging scenarios.
+    realistic medical imaging scenes.
 
-    \item \textbf{CVC-ClinicDB Dataset:} 
+    \item \textbf{CVC-ClinicDB:} 
     Real image patches extracted from the CVC-ClinicDB 
     dataset, a well-known benchmark for polyp detection. 
-    This dataset provides a diverse set of polyp images 
-    with varying characteristics, allowing for robust 
-    evaluation of classification performance.
+    %This dataset provides a diverse set of polyp images with varying characteristics, allowing for robust evaluation of classification performance.
 
-    \item \textbf{MNIST Dataset:} 
+    \item \textbf{MNIST:} 
     A reduced version of the MNIST dataset, 
     containing only the digits 0 and 1, suitable 
-    for binary classification tasks. This dataset 
-    serves as a controlled environment to assess 
-    the model's performance on simpler classification tasks.
+    for binary classification tasks.
+    %This dataset serves as a controlled environment to assess the model's performance on simpler classification tasks.
 \end{enumerate}
@@
 \subsubsection{Principal Component Analysis (PCA)}
-PCA was employed as a baseline linear 
-dimensionality reduction technique. 
+%PCA was employed as a baseline linear dimensionality reduction technique.
 We extracted the top-k principal components 
@@
 \subsubsection{Autoencoder}
-We implemented a neural network-based autoencoder with the following architecture:
+We implemented a simple neural network-based autoencoder with a encoder network that compresses the input to a lower dimensional latent space and a decoder network that reconstructs the original input from the latent representation.
+To improve training stability and prevent overfitting, we also added batch normalization and dropout layers.
+
+The autoencoder was trained to minimize reconstruction loss using the Adam optimizer with learning rates between $0.001$--$0.01$ and weight decay for regularization.
+
+\subsubsection{Quantum-Guided Autoencoder}
+We introduced a quantum-guided autoencoder that incorporates feedback from the quantum reservoir during training. This approach combines:
 
 \begin{itemize}
-    \item An encoder network that compresses
-    the input to a lower-dimensional latent space.
-
-    \item A decoder network that reconstructs 
-    the original input from the latent representation
-
-    \item Batch normalization and dropout layers 
-    to improve training stability and prevent overfitting.
+    \item A standard autoencoder architecture.
+    \item A linear classifier that maps quantum embeddings to class labels.
+    \item A classification loss computed on quantum embeddings.
+    \item A weighting parameter $\lambda$ to control the trade-off between reconstruction and classification objectives~\ref{eq:composed_loss_function}.
+    \item A differentiable surrogate model that approximates the quantum reservoir's behavior, enabling end-to-end backpropagation through the entire pipeline.
+The frequency of quantum updates is controlled by a hyperparameter, allowing for flexible integration of quantum dynamics into the training process.
 \end{itemize}
 
-The autoencoder was trained to minimize reconstruction 
-loss using the Adam optimizer with learning rates 
-between $0.001$--$0.01$ and weight decay for regularization.
-
-\subsubsection{Quantum-Guided Autoencoder}
-We introduced a quantum-guided autoencoder 
-that incorporates feedback from the quantum 
-reservoir during training. This approach combines:
+\subsection{Quantum Reservoir Computing Layer}
+Our quantum reservoir computing implementation uses a Rydberg Atom simulator with the following parameterizable components of an atomic chain:
 
 \begin{itemize}
-    \item A standard autoencoder architecture
-    \item A linear classifier that maps quantum embeddings to class labels.
-    \item A classification loss computed on quantum embeddings
-    \item A weighting parameter $\lambda$ to control the trade-off between 
-    reconstruction and classification objectives~\ref{eq:composed_loss_function}.
-    \item A differentiable surrogate model that approximates the quantum reservoir's behavior,
-    enabling end-to-end backpropagation through the entire pipeline. The frequency 
-    of quantum updates is controlled by a hyperparameter, allowing for
-    flexible integration of quantum dynamics into the training process.
-\end{itemize}
-
-\subsubsection{Quantum Reservoir Computing Layer}
-Our quantum reservoir computing implementation uses 
-a Rydberg atom simulator with the following parameterizable
-components:
-
-\begin{itemize}
-    \item \textbf{Atom Chain Length:} 
-    The number of atoms in the chain, which can be adjusted 
-    to control the reservoir's capacity and complexity.
+    \item \textbf{Atom Chain Length (number of qubits):} 
+    The number of atoms in the chain.
 
     \item \textbf{Rabi Frequency:} 
-    The global Rabi drive amplitude, which influences the 
-    strength of interactions between atoms.
+    The global Rabi drive amplitude, which influences the strength of interactions between atoms.
 
-    \item \textbf{Detuning Parameters:} 
-    Site-dependent detuning parameters that encode the input data.
+    \item \textbf{Lattice Spacing:} 
+    The empty space between two consecutive atoms.
 
@@
 
-    \item \textbf{Encoding Scale:}
-    The scale of the input data encoding, which can be adjusted
-    to optimize the quantum reservoir's response to the input data.
+    \item \textbf{Time steps:}
+    The number of time steps where the reservoir will be probed.
 
-    \item \textbf{Time steps:}
-    The number of time steps for which the quantum reservoir evolves,
-    allowing for the capture of temporal dynamics in the data.
+    \item \textbf{Evolution Time:}
+    The time we let the quantum reservoir evolve under the Hamiltonian~\ref{eq:rydberg_hamiltonian}.
 \end{itemize}
 
-\subsubsection{Classical Network Architectures}
-After the traning process of the quantum-guided autoencoder,
+Unless explicitly stated, the results that we show in the next section~\ref{results_and_discussion} were obtained with the following hyperparameter values: $12$ qubits, $10 \mu m$ between atoms, \(\langle Z_i\rangle\), \(\langle Z_iZ_j\rangle,\) and \(\langle Z_iZ_jZ_k\rangle\) observables, Rabi Frequency $\Omega = \ \pi$, $T = 16$ time steps and evolved over $4.0 \mu s$ (which means that the system was probed every $\frac{4.0}{T} = \Delta t = 0.25 \mu s$.
+
+We can use Eq.~\ref{eq:total_embedding_dimension} to determine the dimension of our quantum embeddings \(\lvert \bm q \rvert \) when using the mentioned parameters:
+
+\begin{equation}
+    \lvert \bm q \rvert = \frac{16 \times 12^3 + 5 \times 16 \times 12}{6} = 4768
+    \label{eq:experimental_embedding_dimension}
+\end{equation}
+
+\subsection{Linear Classifier for the QRC Embeddings} \label{linear_classifier_for_the_qrc_embeddings}
+After the traning process of the QGARS model,
 the following quantum embeddings are passed to a linear classifier.
-The classifier is a simple feedforward neural network with the following architecture:
+The classifier is a simple feedforward neural network without hidden layers with the following architecture:
 \begin{itemize}
-    \item A single hidden layer with a ReLU activation function.
+    \item An input layer with the size of the embeddings.
+    \item An output layer with two neurons.
     \item Cross-entropy loss for training, optimized using the Adam optimizer.
 \end{itemize}
+In total, this linear classifier has \(9538\) parameters when the reservoir contains $12$ qubits.
 
 \subsection{Comparison Methods}
-We benchmarked our model against some
-classical methods, including:
+We benchmarked our model against the following classical methods:
 \begin{itemize}
-    \item \textbf{PCA + Linear Classifier:} 
-    A linear classifier trained on the PCA-reduced features.
+    \item \textbf{PCA/Autoencoder + Linear Classifier:} 
+    A linear classifier trained on the PCA-reduced features or the latent representations from the autoencoder.
+Essentially, this model is the same as the one used to classify the quantum embeddings described in section~\ref{linear_classifier_for_the_qrc_embeddings}, but it only receives the classically extracted features as input.
 
-    \item \textbf{Autoencoder + Linear Classifier:} 
-    A linear classifier trained on the latent representations from the autoencoder.
-
-    \item \textbf{PCA + Neural Network:}
-    A fully connected neural network trained on the PCA-reduced features.
-
-    \item \textbf{Autoencoder + Neural Network:}
-    A fully connected neural network trained on the latent representations from the autoencoder.
+    \item \textbf{PCA/Autoencoder + Neural Network:}
+    A fully connected 4-layer neural network with two $100$-neuron hidden layers, trained on the features from the same reduction methods as the linear classifier.
+This neural network has $11602$ parameters when training on $12$ dimensional inputs.
 \end{itemize}
 
-\subsection{Performance Metrics}
-We evaluated the performance of our model using the following metrics:
-Classification Accuracy, F1 Score, Confusion Matrix.
+We also benchmarked our QGARS encoding strategy
+against the performance of the Linear Mapping of
+QRC embeddings extracted from the latent representation 
+of data using other feature reduction methods, such as
+PCA and the Classical Autoencoder.
+
+% \subsection{Performance Metrics}
+% We evaluated the performance of our model using the following metrics:
+% Classification Accuracy, F1 Score, Confusion Matrix.
 
@@
     \item \textbf{Guided Lambda Parameter (\( \lambda \))}: 
-    This parameter controls the trade-off between reconstruction and classification objectives in the loss function~\ref{eq:composed_loss_function}.
-We tested values ranging from 0.1 to 0.9 in increments of 0.1.
+    To analyse the effect that the trade-off between reconstruction and classification objectives in the loss function~\ref{eq:composed_loss_function} affects
+    the QRC
 
     \item \textbf{Quantum Update Frequency:} 
-    The frequency at which the quantum reservoir is updated during training. We experimented with update frequencies of 1, 5, and 10 epochs.
+    The frequency at which the quantum reservoir is updated during training. We experimented with update frequencies of 5, 7, 10 and 25 epochs.
 
-    \item \textbf{Quantum Parameters:} 
-    Including the number of atoms in the chain, Rabi frequency, and detuning parameters. We varied these parameters to assess their impact on classification performance.
-    \item \textbf{Batch Size:}
-    \item \textbf{Learning Rate:}
+    % \item \textbf{Quantum Parameters:} 
+    % Rabi frequency, and detuning parameters. We varied these parameters to assess their impact on classification performance.
+
+    \item \textbf{Number of Features:}
+    The number of features is the same as the number of qubits (atoms), 
+    as each feature encoded in each atom's local detuning parameter.
+
 \end{itemize}
+The results from these experiments are discussed in the next section~\ref{results_and_discussion}.
 
@@
 %============================================
-\section{Results and Discussion}
-\subsection{Classification Performance Comparison}
+\section{Results and Discussion} \label{results_and_discussion}
+
+\subsection{Dimensionality Reduction Method Comparison}
+We first compare classification accuracy of the Linear Classifier on the QRC embeddings across the different dimensionality reduction methods PCA, vanilla autoencoder, and our QGARS pipeline on each dataset.
+Table~\ref{tab:qrc_reduction_accuracy} summarizes mean accuracy and standard deviation over five runs, using 12 features/qubits.
+These results show that the QGARS pipeline is able to outperform traditional feature reduction methods commonly used as preprocessing steps on QRC pipelines.
+
+\begin{table}[!b]
+  \caption{Classification accuracy (\%) on QRC with different dimensionality reduction methods (2000 training images and 400 testing images)}
+  \label{tab:qrc_reduction_accuracy}
+  \centering
+  \begin{tabular}{lccc}
+    \hline
+    Method         & Synthetic  & CVC-ClinicDB & MNIST(0/1) \\
+    \hline
+    PCA + QRC      &  00 ± 0.0  &  84.95 ± 2.44    &  97.70 ± 0.89 \\
+    AE + QRC       &  00 ± 0.0  &  73.75 ± 1.91    &  87.05 ± 1.85 \\
+    \textbf{QGARS + QRC}    &  \textbf{00 ± 0.0} &  \textbf{88.90 ± 1.47} &  \textbf{00 ± 0.0} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=\linewidth]{images/results/generated_polyp_dataset/n_qubit_qrc_performance_by_reduction_method.pdf} 
+    \caption{Accuracy of the different dimensionality reduction methods for QRC as a function of the number of qubits with the Synthetic Polyp Dataset.
+(green) QGARS, (blue) PCA and (red) Autoencoder.
+The dashed lines represent the training accuracy, while the whole lines are the test accuracies.}
+    \label{fig:n_qubits_accuracy}
+\end{figure}
+
+\begin{figure*}[!ht]
+  \centering
+  \includegraphics[width=0.32\textwidth]{images/results/generated_polyp_dataset/quantum_embeddings_2d_tsne_PCA.pdf}
+  \includegraphics[width=0.32\textwidth]{images/results/generated_polyp_dataset/quantum_embeddings_2d_tsne_autoencoder.pdf}
+  \includegraphics[width=0.32\textwidth]{images/results/generated_polyp_dataset/quantum_embeddings_2d_tsne_guided_autoencoder.pdf}
+  \caption{2-dimensional t-SNE projections of the 4768-dimensional QRC embeddings for (left) PCA + QRC, (center) AE + QRC, (right) QGARS + QRC}
+  \label{fig:embeddings_tsne}
+\end{figure*}
+
+We also evaluate the effect of the number of qubits $N$ on the performance with the different dimensionality reduction methods as shown in Fig.~\ref{fig:n_qubits_accuracy} in general, it's easier to train the model as we increase $N$, the slight performance drop as we go from $N = 10$ to $N = 12$ might just be a byproduct due to the higher dimensionality of $\bm q$, we expect that with a larger amount of input images, the accuracy would keep improving.
+We hope to better understand this effect with future experiments.
+
+In order to have a visual demonstration of how well the quantum reservoir embeddings separate classes, we project the high-dimensional embeddings down to $3$ dimensions via t-distributed Stochastic Neighbour Embedding (t-SNE).
+Fig.~\ref{fig:embeddings_tsne} shows side-by-side plots for PCA → QRC, AE → QRC, and QGARS → QRC pipelines.
+The QGARS embeddings exhibit the tightest and most distinct clusters, indicating superior discriminative structure.
+
 \subsection{Ablation Studies}
-\subsubsection{Impact of Guided Lambda Parameter}
-\subsubsection{Effect of Quantum Update Frequency}
-\subsubsection{Influence of Quantum Parameters}
-\subsection{Dimensionality Reduction Comparison}
-\subsection{Surrogate Model Fidelity Analysis}
-\subsection{Generalization to Unseen Data}
+
+%\subsubsection{Impact of Guided Lambda Parameter}
+We verified that varying the guidance weighting $\lambda$ in Eq.~\ref{eq:composed_loss_function} has significant impact on the performance of our accuracy.
+To assess its influence, Fig.~\ref{fig:lambda_sweep} plots test accuracy vs.\ $\lambda$.
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=1\linewidth]{images/results/generated_polyp_dataset/guided_lambda_log_mu_effect.pdf}
+  \caption{Accuracy as a function of the reconstruction/classification trade-off parameter $\lambda$.}
+  \label{fig:lambda_sweep}
+\end{figure}
+
+\noindent
+We find optimal performance around $\lambda=0.5$; too little guidance ($\lambda<0.1$) approaches a point where it is equivalent to a traditional AE, so ir yields under-trained classifiers.
+
+A strange effect that we verified was that even using the value \(\lambda = 1\), the performance of the classifier is still competitive.
+Although we still need to study this phenomenon, it suggests that the classical part of the autoencoder might not be as important as the surrogate component when producing a latent space that is appropriate for QRC classification.
+
+In future work, we plan to to conduct more ablation studies, mainly on the effect that the frequency with which we update the surrogate model and the influence of the quantum parameters in the reservoir layer.
+%\subsubsection{Effect of Quantum Update Frequency}
+
+%\subsubsection{Influence of Quantum Parameters}
+
+% \subsection{Surrogate Model Fidelity Analysis}
+% We assess how well the surrogate approximates the quantum reservoir by 
+% computing the MSE between true quantum embeddings and surrogate outputs over the test set (Fig.~\ref{fig:surrogate_fidelity}). Low MSE indicates high fidelity.
+
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics[width=\linewidth]{surrogate_mse.pdf}
+%   \caption{Surrogate vs.\ quantum embedding MSE over training epochs.}
+%   \label{fig:surrogate_fidelity}
+% \end{figure}
+
+\subsection{Comparison with Classical Methods}
+To benchmark the benefit of our QGARS pipeline, 
+we compare against purely classical feature-reduction 
+approaches followed by standard classifiers. 
+Table~\ref{tab:classical_comp} reports test accuracy on each dataset 
+for the different methods.
+
+\begin{table}[tb]
+  \caption{Classification accuracy (\%) for classical baselines (2000 training images and 400 testing images)}
+  \label{tab:classical_comp}
+  \centering
+  \begin{tabular}{lccc}
+    \hline
+    Method            & Synthetic & CVC-ClinicDB & MNIST (0/1) \\
+    \hline
+    PCA + Linear      &  82.0 ± 0.79    &  71.05 ± 1.35   &  99.90 ± 0.12    \\
+    AE + Linear       &  80.3 ± 3.01    &  68.25 ± 2.31   &  98.80 ± 0.29    \\
+    PCA + NN          &  85.85 ± 1.24   &  91.65 ± 1.38   &  99.80 ± 0.19    \\
+    AE + NN           &  83.75 ± 2.21   &  90.50 ± 1.72   &  99.90 ± 0.12    \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\noindent
+In contrast, the QGARS pipeline achieves a maximum of $\mathbf{0.0\%}$ and $\mathbf{90.50\%}$, $\mathbf{0.0\%}$ on the same datasets (see Table~\ref{tab:qrc_reduction_accuracy}).
+
+\subsubsection{Training Loss Dynamics}
+
+Fig.~\ref{fig:loss_curves} shows the training loss curves for the classical baselines and QGARS on the Synthetic Polyp dataset.
+We plot the reconstruction loss for PCA/AE methods combined with classification loss for the AE+NN and QGARS pipelines.
+
+\noindent
+Although the loss function of the QGARS pipeline shows signs of instability, which can be attributed to the high dimensionality of the quantum embeddings, it reaches a lower loss in comparison to the classical counterparts.
 
@@
 \subsection{Current Limitations}
+While QGARS demonstrates strong performance, several limitations remain:
+\begin{itemize}
+    \item \textbf{Classical overhead:}  
+    Frequent surrogate retraining and batch quantum simulations introduce non-trivial classical compute overhead, potentially offsetting some of the quantum advantage.
+
+    \item \textbf{Data and Task Specificity:}
+    The current architecture has been evaluated primarily on specific medical imaging tasks. Its generalization to other domains or types of data remains to be thoroughly investigated.
+
+\end{itemize}
+
+\begin{figure}[!tb]
+  \centering
+  \includegraphics[width=\linewidth]{images/results/generated_polyp_dataset/classifier_losses.pdf}
+  \caption{Training loss vs. epoch for PCA+Linear, AE+Linear, AE+NN, and QGARS with \( \lambda=0.7 \) on the Synthetic Polyp dataset.}
+  \label{fig:loss_curves}
+\end{figure}
+
 \subsection{Potential Extensions}
-\subsection{Hardware Implementation Considerations}
+Several avenues can further enhance QGARS:
+\begin{itemize}
+    \item \textbf{Different Hamiltonians:}
+    Explore alternative interaction Hamiltonians (such as the one from the Heisenberg XXZ model) to better tailor reservoir dynamics, as experimented by A.
+D.
+Lorenzis et al.~\cite{lorenzisHarnessingQuantumExtreme2025}.
+
+    \item \textbf{Convolutional Autoencoders:}
+    Replace the fully connected autoencoder with a convolutional AE~\cite{lerchConvolutionalAutoencodersSpatiallyinformed2022} to learn spatially localized features.
+That would not only allow the model to better manage input size but also provide a richer structure for the reservoir.
+
+    \item \textbf{Reservoir Encoding Schemes:}
+    Investigate alternative methods of encoding data into the atomic register such as \textit{position encoding} or \textit{global pulse encoding}.
+We could also encode the features in $2$-dimensional register and experiment with the different possible lattice configurations.
+~\cite{kornjavcaLargescaleQuantumReservoir2024}.
+
+    \item \textbf{Surrogate Model Complexity:}
+    As the surrogate model tries to replicate a highly complex system such as the quantum reservoir, it naturally has an extremely large amount of parameters.
+To keep the surrogate parameter count in check, we aim to explore alternative surrogate model architectures such as Physics-Informed Neural Networks (PINNs)~\cite{raissiPhysicsInformedDeep2017}, which incorporate known physical laws into the learning process, potentially improving the fidelity of the surrogate model.
+
+    \item \textbf{Experiment on Real Quantum Hardware:}
+    Although the experiments we conducted on simulators showed promising results, to validate the usefulness of our pipeline in the NISQ era, we still need to run experiences on real QPUs with more qubits.
+
+\end{itemize}
 
@@
 \section{Conclusion}
+
+We have presented Quantum-Guided Autoencoder with Reservoir Surrogate (QGARS), a hybrid quantum-classical architecture that integrates a classical autoencoder with a neutral-atom quantum reservoir and a differentiable surrogate model to enable end‐to‐end training.
+By jointly optimizing reconstruction and classification losses, QGARS learns compact and discriminative features tailored for quantum reservoir computing.
+Our experiments on synthetic polyp, CVC-ClinicDB, and binary MNIST datasets demonstrate that QGARS outperforms classical PCA‐ and autoencoder‐based baselines that are commonly used for Quantum Reservoir Computing Pipelines.
+Ablation studies highlight the importance of guidance weight \(\lambda\), surrogate update frequency, and reservoir parameters.
+Although challenges remain—particularly regarding high‐dimensional embeddings, surrogate fidelity, and classical overhead, our proposed extensions and hardware considerations chart a clear path toward robust, scalable quantum‐enhanced medical imaging pipelines in the NISQ era.
 

### 2025-05-31 archive/embeddings_visualization.py
results for the paper

@@ -1,0 +1,898 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D  # Import for 3D plotting
+from sklearn.manifold import TSNE
+import umap
+from sklearn.metrics import silhouette_score, davies_bouldin_score, calinski_harabasz_score
+import seaborn as sns
+import os
+import json
+from typing import Dict, List, Tuple, Any, Optional
+from scipy.spatial.distance import pdist, squareform
+import plotly.graph_objects as go  # For interactive 3D plots
+
+from src.data_processing.data_processing import load_dataset, select_random_samples
+from src.feature_reduction.feature_reduction import (
+    apply_pca, apply_autoencoder, apply_guided_autoencoder,
+    scale_to_detuning_range
+)
+from src.quantum_layer.qrc_layer import DetuningLayer
+
+N_FEATURES = 12  # Number of features for the quantum layer
+
+def generate_quantum_embeddings(
+    data_train: Dict[str, Any],
+    quantum_layer: DetuningLayer,
+    method_name: str,
+    encoding_dim: int = N_FEATURES,
+    n_examples: int = 100,
+    guided_lambda: float = 0.3,
+    epochs: int = 30
+) -> Tuple[np.ndarray, np.ndarray]:
+    """Generate quantum embeddings using the specified reduction method"""
+    print(f"\n== Processing {method_name} ==")
+    
+    # Select a subset of samples
+    selected_indices, selected_features, selected_targets = select_random_samples(
+        data_train, num_samples=n_examples, seed=42
+    )
+    
+    # Apply dimensionality reduction
+    if method_name == "pca":
+        xs, _, spectral = apply_pca(
+            data_train, 
+            dim_pca=encoding_dim,
+            selected_indices=selected_indices,
+            selected_features=selected_features
+        )
+    elif method_name == "autoencoder":
+        xs, _, spectral = apply_autoencoder(
+            data_train, 
+            encoding_dim=encoding_dim,
+            epochs=epochs,
+            selected_indices=selected_indices,
+            selected_features=selected_features
+        )
+    elif method_name == "guided_autoencoder":
+        xs, _, spectral, _ = apply_guided_autoencoder(
+            data_train,
+            quantum_layer=quantum_layer,
+            encoding_dim=encoding_dim,
+            epochs=epochs,
+            guided_lambda=guided_lambda,
+            selected_indices=selected_indices,
+            selected_features=selected_features,
[…]

### 2025-05-31 data/data_generation/data_augmentation.py

@@ -56,7 +56,7 @@ def augment_dataset(input_dir, output_dir, target_size, img_ext=".jpg"):
     os.makedirs(output_dir, exist_ok=True)
     
     # Get all input image paths
-    image_files = [f for f in os.listdir(input_dir) if f.endswith(img_ext)]
+    image_files = [f for f in os.listdir(input_dir) if (f.endswith(img_ext) or f.endswith('png'))]
     current_size = len(image_files)
     
     if current_size >= target_size:

### 2025-05-31 parameter_sweep/data_analysis_scripts/n_qubit_reduction_method_analysis.py

@@ -103,74 +103,6 @@ def plot_qrc_performance_by_method(results_file, output_dir="../../results/figur
     # Show the plot
     plt.show()
 
-def analyze_performance_drops(results_file):
-    """
-    Analyze why performance sometimes drops when adding more qubits/features.
-    
-    Parameters
-    ----------
-    results_file : str
-        Path to the CSV file containing the results
-    """
-    print("\nAnalyzing Performance Fluctuations with Increasing Qubit Count")
-    print("=" * 70)
-    
-    # Read the CSV file
-    df = pd.read_csv(results_file)
-    
-    # Filter only successful runs
-    df = df[df['status'] == 'success']
-    
-    # For each reduction method, analyze performance patterns
-    methods = ['pca', 'autoencoder', 'guided_autoencoder']
-    
-    for method in methods:
-        # Filter for this reduction method
-        method_df = df[df['reduction_method'] == method]
-        
-        if method_df.empty:
-            continue
-            
-        # Sort by dimension
-        method_df = method_df.sort_values('dim_reduction')
-        
-        print(f"\nMethod: {method.upper()}")
-        print("-" * 30)
-        
-        # Print dimension vs performance
-        print(f"{'Qubits':<8} | {'Test Acc':<10} | {'Train Acc':<10} | {'Gap':<10} | {'Observation'}")
-        print("-" * 70)
-        
-        prev_acc = None
-        for _, row in method_df.iterrows():
-            qubits = row['dim_reduction']
-            test_acc = row['QRC_final_test_acc']
-            train_acc = row['QRC_final_train_acc']
-            gap = train_acc - test_acc
-            
-            observation = ""
-            if prev_acc is not None:
-                if test_acc < prev_acc:
-                    observation = "DECREASED - Likely overfitting or noise"
-                    if gap > 0.2:
-                        observation += " (large train-test gap suggests overfitting)"
-                elif test_acc > prev_acc + 0.05:
-                    observation = "IMPROVED significantly"
-                else:
-                    observation = "Small change"
-            
-            print(f"{qubits:<8} | {test_acc:<10.3f} | {train_acc:<10.3f} | {gap:<10.3f} | {observation}")
-            prev_acc = test_acc
-            
-    print("\nExplanations for Performance Drops:")
-    print("-" * 70)
-    print("1. Overfitting: As dimensionality increases, models may fit to noise rather than signal")
[…]

### 2025-05-31 parameter_sweep/run_experiments.py

@@ -143,12 +143,12 @@ EXPERIMENT_PROFILES = {
     "ae_pca_qgars_5_runs": {
         "description": "Run a general test with fixed parameters",
         "param_grid": {
-            "seed": [42, 43, 44, 45, 46],
-            "dataset_type": ["binary_mnist", "generated_polyp_dataset", "cvc_clinic_db_patches"],
+            "dataset_type": ["cvc_clinic_db_patches"],
             "num_examples": [2000],
             "num_test_examples": [400],
             "reduction_method": ["guided_autoencoder", "autoencoder", "pca"],
-            "dim_reduction": [12]
+            "dim_reduction": [12],
+            "seed": [42, 43, 44, 45, 46]
         }
     },
     

### 2025-05-31 src/embeddings_visualization.py

@@ -1,270 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from mpl_toolkits.mplot3d import Axes3D  # Import for 3D plotting
-from sklearn.manifold import TSNE
-import umap
-from sklearn.metrics import silhouette_score
-import seaborn as sns
-import os
-import json
-from typing import Dict, List, Tuple, Any, Optional
-
-from src.data_processing.data_processing import load_dataset, select_random_samples
-from src.feature_reduction.feature_reduction import (
-    apply_pca, apply_autoencoder, apply_guided_autoencoder,
-    scale_to_detuning_range
-)
-from src.quantum_layer.qrc_layer import DetuningLayer
-
-N_FEATURES = 12  # Number of features for the quantum layer
-
-def generate_quantum_embeddings(
-    data_train: Dict[str, Any],
-    quantum_layer: DetuningLayer,
-    method_name: str,
-    encoding_dim: int = N_FEATURES,
-    n_examples: int = 100,
-    guided_lambda: float = 0.3,
-    epochs: int = 30
-) -> Tuple[np.ndarray, np.ndarray]:
-    """Generate quantum embeddings using the specified reduction method"""
-    print(f"\n== Processing {method_name} ==")
-    
-    # Select a subset of samples
-    selected_indices, selected_features, selected_targets = select_random_samples(
-        data_train, num_samples=n_examples, seed=42
-    )
-    
-    # Apply dimensionality reduction
-    if method_name == "pca":
-        xs, _, spectral = apply_pca(
-            data_train, 
-            dim_pca=encoding_dim,
-            selected_indices=selected_indices,
-            selected_features=selected_features
-        )
-    elif method_name == "autoencoder":
-        xs, _, spectral = apply_autoencoder(
-            data_train, 
-            encoding_dim=encoding_dim,
-            epochs=epochs,
-            selected_indices=selected_indices,
-            selected_features=selected_features
-        )
-    elif method_name == "guided_autoencoder":
-        xs, _, spectral, _ = apply_guided_autoencoder(
-            data_train,
-            quantum_layer=quantum_layer,
-            encoding_dim=encoding_dim,
-            epochs=epochs,
-            guided_lambda=guided_lambda,
-            selected_indices=selected_indices,
-            selected_features=selected_features,
-            selected_targets=selected_targets
-        )
-    
-    # Scale to detuning range
[…]

### 2025-05-31 src/classification_models/training.py
cleaning

@@ -7,14 +7,19 @@ from torch.utils.data import DataLoader, TensorDataset
 from tqdm import tqdm
 from typing import Tuple, List, Dict, Any, Optional, Union
 from sklearn.metrics import confusion_matrix, f1_score
+from src.globals import ResultsDict
 
+from archive.benchmarking.parameter_testing import RESULTS_DIR
 from src.classification_models.models import LinearClassifier, NeuralNetwork
+
 
 def train(x_train: np.ndarray, y_train: np.ndarray, 
           x_test: np.ndarray, y_test: np.ndarray, 
           regularization: float = 0.0, nepochs: int = 100, 
           batchsize: int = 100, learning_rate: float = 0.01, 
-          verbose: bool = True, nonlinear: bool = False) -> Tuple[List[float], List[float], List[float], nn.Module]:
+          verbose: bool = True, nonlinear: bool = False
+          ) -> Tuple[List[float], List[float], List[float], nn.Module, 
+           np.ndarray, np.ndarray, float, float]:
     """
     Train a model on the given data.
     
@@ -43,11 +48,16 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
     
     Returns
     -------
-    Tuple[List[float], List[float], List[float], nn.Module]
+    Tuple[List[float], List[float], List[float], nn.Module, 
+           np.ndarray, np.ndarray, float, float]
         - losses: Training losses per epoch
         - accs_train: Training accuracies per epoch
         - accs_test: Test accuracies per epoch
         - model: Trained model
+        - confusion_matrix_train: Confusion matrix for training set
+        - confusion_matrix_test: Confusion matrix for test set
+        - f1_train: F1 score for training set
+        - f1_test: F1 score for test set
     """
     # Convert numpy arrays to PyTorch tensors
     x_train_tensor = torch.FloatTensor(x_train.T)  # Transpose to match PyTorch's expected shape
@@ -98,6 +108,13 @@ def train(x_train: np.ndarray, y_train: np.ndarray,
     losses = []
     accs_train = []
     accs_test = []
+    train_targets: Optional[torch.Tensor] = None
+    train_pred: Optional[torch.Tensor] = None
+    test_pred: Optional[torch.Tensor] = None
+    confusion_matrix_train: Optional[np.ndarray] = None
+    confusion_matrix_test: Optional[np.ndarray] = None
+    f1_train: Optional[float] = None
+    f1_test: Optional[float] = None
[…]

### 2025-05-31 src/globals.py

@@ -1,0 +1,27 @@
+from pathlib import Path 
+from typing import Dict, List, Tuple, Any
+import numpy as np
+import torch.nn as nn
+
+# Define the root directory of the project
+ROOT_DIR = Path(__file__).resolve().parent.parent
+
+# Define the path to the datasets directory
+DEFAULT_DATA_DIR = ROOT_DIR / 'data' / 'datasets'
+
+# Define the path to the default results directory
+DEFAULT_RESULTS_DIR = ROOT_DIR / 'results' / 'default_results'
+    
+# Define results dictionary typing
+ResultsDict = Dict[str, 
+                   Tuple[
+                        List[float],    # Training losses per epoch
+                        List[float],    # Training accuracies per epoch
+                        List[float],    # Test accuracies per epoch
+                        nn.Module,      # Trained model
+                        np.ndarray,     # Confusion matrix for training set
+                        np.ndarray,     # Confusion matrix for test set
+                        float,          # F1 score for training set
+                        float           # F1 score for test set
+                        ]
+                   ]

### 2025-05-31 src/main.py

@@ -8,8 +8,9 @@ import torch
 import matplotlib.pyplot as plt
 from typing import Dict, Any, Tuple, Optional, List
 import time
+from pathlib import Path
 
-sys.path.append(os.path.dirname(os.path.abspath(__file__)))
+#sys.path.append(os.path.dirname(os.path.abspath(__file__)))
 
 # Import custom modules
 from src.feature_reduction.autoencoder.autoencoder import Autoencoder
@@ -27,10 +28,12 @@ from src.quantum_layer.qrc_layer import DetuningLayer
 from src.classification_models.training import train
 from src.utils.visualization import plot_training_results, print_results
 
-from utils.cli_utils import get_args
+from src.utils.cli_utils import get_args
 import argparse
 
-def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> Tuple[Dict[str, Tuple[np.ndarray, np.ndarray, np.ndarray, torch.nn.Module]], Optional[Dict[str, List[float]]]]:
+from src.globals import ROOT_DIR, DEFAULT_DATA_DIR, DEFAULT_RESULTS_DIR
+
+def main(args: Optional[argparse.Namespace] = None, results_dir: Path = DEFAULT_RESULTS_DIR) -> Tuple[Dict[str, Tuple[np.ndarray, np.ndarray, np.ndarray, torch.nn.Module]], Optional[Dict[str, List[float]]]]:
     """
     Main function to run the quantum reservoir computing pipeline.
     
@@ -66,21 +69,28 @@ def main(args: Optional[argparse.Namespace] = None, results_dir: str = None) -> 
           """)
     
 
-    DATA_DIR = args.data_dir if args.data_dir else os.path.join(os.path.dirname(os.path.abspath(__file__)), "..", "data", "datasets")
+    #DATA_DIR = args.data_dir if args.data_dir else os.path.join(os.path.dirname(os.path.abspath(__file__)), "..", "data", "datasets")
+    # first, declare the DATA_DIR variable as a string
+    data_dir: Path
+    if args.data_dir is None:
+        data_dir = DEFAULT_DATA_DIR
+    else:
+        data_dir = Path(args.data_dir)
 
     # Load dataset based on the specified dataset type
     if args.dataset_type in ["mnist", "fashion_mnist", "binary_mnist"]:
         # Define the path to MNIST dataset
         data_train, data_test = load_dataset(
             args.dataset_type,
-            data_dir=DATA_DIR,
+            data_dir=data_dir,
             target_size=tuple(args.target_size),
[…]

### 2025-05-31 src/utils/config_manager.py

@@ -4,7 +4,9 @@ import argparse
 import numpy as np
 from typing import Dict, Any, Optional
 
-DEFAULT_RESULTS_DIR = os.path.join(os.path.dirname(os.path.dirname(__file__)), "..", "results")
+from src.globals import DEFAULT_RESULTS_DIR
+
+#DEFAULT_RESULTS_DIR = os.path.join(os.path.dirname(os.path.dirname(__file__)), "..", "results")
 
 class ConfigManager:
     """

### 2025-05-31 src/utils/statistics_tracking.py

@@ -6,7 +6,10 @@ import json
 import datetime
 import logging
 
-def ensure_directory_exists(directory: str) -> None:
+from pathlib import Path
+from src.globals import ResultsDict
+
+def ensure_directory_exists(directory: Path) -> None:
     """
     Ensure the specified directory exists, creating it if necessary.
     
@@ -15,10 +18,13 @@ def ensure_directory_exists(directory: str) -> None:
     directory : str
         Directory path to check/create
     """
-    if not os.path.exists(directory):
-        os.makedirs(directory, exist_ok=True)
+    # if not os.path.exists(directory):
+    #     os.makedirs(directory, exist_ok=True)
+    if not directory.exists():
+        directory.mkdir(parents=True, exist_ok=True)
 
-def setup_stats_directory(base_dir: str = "run_stats") -> str:
+
+def setup_stats_directory(base_dir: Path) -> Path:
     """
     Set up a directory for saving run statistics.
     
@@ -29,12 +35,15 @@ def setup_stats_directory(base_dir: str = "run_stats") -> str:
         
     Returns
     -------
-    str
-        Path to the created directory
+    Path
+        Path to the created statistics directory
+
     """
     # Create a timestamped directory
     timestamp = datetime.datetime.now().strftime("%Y%m%d_%H%M%S")
-    stats_dir = os.path.join(os.path.dirname(os.path.abspath(__file__)), base_dir, timestamp)
+    #stats_dir = os.path.join(os.path.dirname(os.path.abspath(__file__)), base_dir, timestamp)
+    stats_dir = Path(base_dir) / timestamp
+    
     ensure_directory_exists(stats_dir)
     
     # Set up logging
@@ -46,7 +55,7 @@ def setup_stats_directory(base_dir: str = "run_stats") -> str:
     
     return stats_dir
 
-def save_classifier_loss_plot(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]], 
+def save_classifier_loss_plot(results_dict: ResultsDict, 
                              output_dir: str) -> None:
     """
     Plot and save training losses for each classifier.
@@ -77,7 +86,7 @@ def save_classifier_loss_plot(results_dict: Dict[str, Tuple[List[float], List[fl
     
     logging.info(f"Saved classifier loss plot to {plot_path}")
 
-def save_classifier_accuracy_plot(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]], 
[…]

### 2025-05-31 src/utils/visualization.py

@@ -1,7 +1,11 @@
 import matplotlib.pyplot as plt
 from typing import Dict, Tuple, List, Any
+import numpy as np
+import torch.nn as nn
+from src.globals import ResultsDict
 
-def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]]) -> None:
+def plot_training_results(
+    results_dict: ResultsDict) -> None:
     """
     Plot training results for multiple models.
     
@@ -36,7 +40,7 @@ def plot_training_results(results_dict: Dict[str, Tuple[List[float], List[float]
     plt.tight_layout()
     plt.show()
 
-def print_results(results_dict: Dict[str, Tuple[List[float], List[float], List[float], Any]]) -> None:
+def print_results(results_dict: ResultsDict) -> None:
     """
     Print the final test accuracies for each model.
     

### 2025-06-01 paper/main.tex
Updates from Overleaf

@@
     \ket{\psi(t+\Delta t)}
-    = U \let{\psi(t)}
+    = U \ket{\psi(t)}
     = e^{-iH\Delta t} \ket{\psi(t)},
@@
 Our proposed pipeline is a Quantum-Guided Autoencoder with a 
-Reservoir Surrogate (QGARS). It is a hybrid architecture that 
+Reservoir Surrogate (QGARS). It is an hybrid architecture that 
 synergizes classical autoencoding techniques with quantum reservoir processing to achieve efficient learning of features.
@@
             \item \textbf{Measurement:}
-            Performs measurements on the quantum state to obtain embeddings \(\bm{q} \in \mathbb{R}^k\), which serve as training targets for the Surrogate Model or as 
+            Performs measurements on the quantum state to obtain embeddings \(\bm{q} \in \mathbb{R}^k\), which serve as training targets for the surrogate model or as 
             inputs for the classification task later on.
@@
     A classical neural network that maps the quantum embeddings \(\bm{q}\) to 
-    predicted class labels \(\hat{y}\).
+    predicted class labels \(\bm{\hat{y}}\).
 
@@
 
-If we measure all of the $3$ proposed observables, the total count of the values measured at a single time step $O_N$ is the sum of the counts for each type of observable:
+If we measure all of the three proposed observables, the total count of the values measured at a single time step $O_N$ is the sum of the counts for each type of observable:
 \begin{equation}
@@
     \item A classification loss computed on quantum embeddings.
-    \item A weighting parameter $\lambda$ to control the trade-off between reconstruction and classification objectives~\ref{eq:composed_loss_function}.
+    \item A weighting parameter $\lambda$ to control the trade-off between reconstruction and classification objectives~\eqref{eq:composed_loss_function}.
     \item A differentiable surrogate model that approximates the quantum reservoir's behavior, enabling end-to-end backpropagation through the entire pipeline.
@@
     \item \textbf{Evolution Time:}
-    The time we let the quantum reservoir evolve under the Hamiltonian~\ref{eq:rydberg_hamiltonian}.
+    The time we let the quantum reservoir evolve under the Hamiltonian~\eqref{eq:rydberg_hamiltonian}.
 \end{itemize}
@@
 
-We can use Eq.~\ref{eq:total_embedding_dimension} to determine the dimension of our quantum embeddings \(\lvert \bm q \rvert \) when using the mentioned parameters:
+We can use~\eqref{eq:total_embedding_dimension} to determine the dimension of our quantum embeddings \(\lvert \bm q \rvert \) when using the mentioned parameters:
 
@@
 \subsection{Comparison Methods}
-We benchmarked our model against the following classical methods:
+We benchmarked the prediction accuracy of our model against the following classical methods:
 \begin{itemize}
@@
 of data using other feature reduction methods, such as
-PCA and the Classical Autoencoder.
+PCA and the classical autoencoder.
+
+The accuracy results were measured with 2000 training images and 400 testing images over 5 different seeds.
 
@@
     \item \textbf{Guided Lambda Parameter (\( \lambda \))}: 
-    To analyse the effect that the trade-off between reconstruction and classification objectives in the loss function~\ref{eq:composed_loss_function} affects
+    To analyse the effect that the trade-off between reconstruction and classification objectives in the loss function~\eqref{eq:composed_loss_function} affects
     the QRC
@@
 \subsection{Dimensionality Reduction Method Comparison}
-We first compare classification accuracy of the Linear Classifier on the QRC embeddings across the different dimensionality reduction methods PCA, vanilla autoencoder, and our QGARS pipeline on each dataset.
+We first compare classification accuracy of the linear classifier on the QRC embeddings across the different dimensionality reduction methods PCA, vanilla autoencoder, and our QGARS pipeline on each dataset.
 Table~\ref{tab:qrc_reduction_accuracy} summarizes mean accuracy and standard deviation over five runs, using 12 features/qubits.
@@
 \begin{table}[!b]
-  \caption{Classification accuracy (\%) on QRC with different dimensionality reduction methods (2000 training images and 400 testing images)}
+  \caption{Mean ± standard deviation of the classification accuracy (\%) QRC with different dimensionality reduction methods}
+  \centering
   \label{tab:qrc_reduction_accuracy}
-  \centering
   \begin{tabular}{lccc}
@@
     \hline
-    PCA + QRC      &  00 ± 0.0  &  84.95 ± 2.44    &  97.70 ± 0.89 \\
-    AE + QRC       &  00 ± 0.0  &  73.75 ± 1.91    &  87.05 ± 1.85 \\
-    \textbf{QGARS + QRC}    &  \textbf{00 ± 0.0} &  \textbf{88.90 ± 1.47} &  \textbf{00 ± 0.0} \\
+    PCA + QRC      &  79.20 ± 2.53  &  84.95 ± 2.31   &  97.70 ± 0.97 \\
+    AE + QRC       &  77.50 ± 2.70  & 73.75 ± 1.90   &  87.05 ± 2.09 \\
+    \textbf{QGARS + QRC}    &  \textbf{89.45 ± 2.06} &  \textbf{88.90 ± 1.65} &  \textbf{97.00 ± 2.37} \\
     \hline
@@
     \includegraphics[width=\linewidth]{images/results/generated_polyp_dataset/n_qubit_qrc_performance_by_reduction_method.pdf} 
-    \caption{Accuracy of the different dimensionality reduction methods for QRC as a function of the number of qubits with the Synthetic Polyp Dataset.
+    \caption{Accuracy of the different dimensionality reduction methods for QRC as a function of the number of qubits with the synthetic polyp dataset.
 (green) QGARS, (blue) PCA and (red) Autoencoder.
@@
   \includegraphics[width=0.32\textwidth]{images/results/generated_polyp_dataset/quantum_embeddings_2d_tsne_guided_autoencoder.pdf}
-  \caption{2-dimensional t-SNE projections of the 4768-dimensional QRC embeddings for (left) PCA + QRC, (center) AE + QRC, (right) QGARS + QRC}
+  \caption{two-dimensional t-SNE projections of the 4768-dimensional QRC embeddings for (left) PCA + QRC, (center) AE + QRC, (right) QGARS + QRC}
   \label{fig:embeddings_tsne}
@@
 
-In order to have a visual demonstration of how well the quantum reservoir embeddings separate classes, we project the high-dimensional embeddings down to $3$ dimensions via t-distributed Stochastic Neighbour Embedding (t-SNE).
+In order to have a visual demonstration of how well the quantum reservoir embeddings separate classes, we project the high-dimensional embeddings down to three dimensions via t-distributed Stochastic Neighbour Embedding (t-SNE).
 Fig.~\ref{fig:embeddings_tsne} shows side-by-side plots for PCA → QRC, AE → QRC, and QGARS → QRC pipelines.
@@
 %\subsubsection{Impact of Guided Lambda Parameter}
-We verified that varying the guidance weighting $\lambda$ in Eq.~\ref{eq:composed_loss_function} has significant impact on the performance of our accuracy.
+We verified that varying the guidance weighting $\lambda$ in~\eqref{eq:composed_loss_function} has significant impact on the performance of our accuracy.
 To assess its influence, Fig.~\ref{fig:lambda_sweep} plots test accuracy vs.\ $\lambda$.
@@
     \hline
-    PCA + Linear      &  82.0 ± 0.79    &  71.05 ± 1.35   &  99.90 ± 0.12    \\
-    AE + Linear       &  80.3 ± 3.01    &  68.25 ± 2.31   &  98.80 ± 0.29    \\
-    PCA + NN          &  85.85 ± 1.24   &  91.65 ± 1.38   &  99.80 ± 0.19    \\
-    AE + NN           &  83.75 ± 2.21   &  90.50 ± 1.72   &  99.90 ± 0.12    \\
+    PCA + Linear      &  82.00 ± 0.88    &  71.05 ± 1.51   &  99.90 ± 0.14   \\
+    AE + Linear       &  80.30 ± 3.37    &  68.25 ± 2.58   &  98.80 ± 0.33   \\
+    PCA + NN          &  85.85 ± 1.39   &  91.65 ± 1.55   &  99.80 ± 0.21    \\
+    AE + NN           &  83.75 ± 2.47    &  90.50 ± 1.93   &  99.90 ± 0.14    \\
     \hline
@@
 \noindent
-In contrast, the QGARS pipeline achieves a maximum of $\mathbf{0.0\%}$ and $\mathbf{90.50\%}$, $\mathbf{0.0\%}$ on the same datasets (see Table~\ref{tab:qrc_reduction_accuracy}).
+In contrast, the QGARS pipeline achieves a maximum of $\mathbf{92.0\%}$ and $\mathbf{90.5\%}$, $\mathbf{99.0\%}$ on the same datasets (see Table~\ref{tab:qrc_reduction_accuracy}).
 
@@
     \item \textbf{Experiment on Real Quantum Hardware:}
-    Although the experiments we conducted on simulators showed promising results, to validate the usefulness of our pipeline in the NISQ era, we still need to run experiences on real QPUs with more qubits.
+    Although the experiments we conducted on simulators showed promising results, to validate the usefulness of our pipeline in the NISQ era, we still need to run experiences on real quantum processing units (QPUs) with more qubits.
 
@@
 By jointly optimizing reconstruction and classification losses, QGARS learns compact and discriminative features tailored for quantum reservoir computing.
-Our experiments on synthetic polyp, CVC-ClinicDB, and binary MNIST datasets demonstrate that QGARS outperforms classical PCA‐ and autoencoder‐based baselines that are commonly used for Quantum Reservoir Computing Pipelines.
+Our experiments on synthetic polyp, CVC-ClinicDB, and binary MNIST datasets demonstrate that QGARS outperforms classical PCA‐ and autoencoder‐based baselines that are commonly used for quantum reservoir computing pipelines.
 Ablation studies highlight the importance of guidance weight \(\lambda\), surrogate update frequency, and reservoir parameters.
